### 2022-11-20 analysis.Rmd
analysis

---
title: "PHQ-4-R: Adaptation of the 4-item screening for Anxiety and Depression for a subclinical sensitivity"
subtitle: "Analysis"
author: "Dominique Makowski et al."
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: paper
    code_folding: hide
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(
  digits = 3,
  mc.cores = 4,
  brms.algorithm = "sampling",
  brms.backend = "cmdstanr",
  dplyr.summarise.inform = FALSE,
  knitr.kable.NA = ""
)

fig.width <- see::golden_ratio(7)
fig.height <- 7

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.path = "./figures/",
  fig.width = fig.width,
  fig.height = fig.height
)
```

See [**tutorial**](https://bookdown.org/bean_jerry/using_r_for_social_work_research/item-response-theory.html) on Item Response Theory (IRT).

# Data

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggside)
library(easystats)
library(patchwork)
library(mirt)

df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |> 
  select(starts_with("Item_PHQ4")) |> 
  datawizard::remove_empty_rows()

names(df) <- str_remove_all(names(df), "Item_PHQ4_")
```

# Depression

```{r message=FALSE, warning=FALSE}
# s <- 'Anxiety = 1,2
#       Depression = 3,4
#       COV = Anxiety*Depression'
# mod1 <- mirt(df, model = mirt.model(s), itemtype = 'graded', SE = TRUE)

data <- select(df, contains("Depression"))

mod1 <- mirt(data, model = 1, itemtype = 'graded', SE = TRUE)

# M2(mod1, type = "C2", calcNULL = FALSE)

itemfit(mod1)

# IRT parameters
IRT_parms <- coef(mod1, IRTpars = TRUE, simplify = TRUE)
IRT_parms$items

# Factor
summary(mod1)

# Plots
plot(mod1, type = 'trace', theta_lim = c(-3, 3))
plot(mod1, type = 'infotrace', theta_lim = c(-3, 3))
plot(mod1, type = 'score', theta_lim = c(-3, 3))
plot(mod1, type = 'infoSE', theta_lim = c(-3, 3))
plot(mod1, type = 'rxx', theta_lim = c(-3, 3))
```

## Category Characteristic Curves (CRC) and Item Information Curves

Typically, an optimally informative item will have a large location and broad category coverage (as indicated by location parameters) over theta.

```{r message=FALSE, warning=FALSE}
crc <- function(mod, data, x = c(-1.25, 2.25)) {
  Theta <- matrix(seq(-3, 3, by = .01))
  rez <- data.frame()
  for(i in 1:2) {
    dat <- as.data.frame(probtrace(extract.item(mod1, i), Theta))
    dat$Theta <- Theta[,1]
    dat$Information <- normalize(iteminfo(extract.item(mod1, i), Theta))
    dat <- pivot_longer(dat, -one_of(c("Theta", "Information")), names_to = "Answer", values_to = "Probability")
    dat$Item <- names(data)[i]
    rez <- rbind(rez, dat)
  }

  rez <- rez |> 
    mutate(
      Answer = case_when(
      Answer == "P.1" ~ "Not at all",
      Answer == "P.2" ~ "Once or twice",
      Answer == "P.3" ~ "Several days",
      Answer == "P.4" ~ "More than half the days",
      TRUE ~ "Nearly every day"
      )) |> 
    mutate(
      Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")))

  p <- rez |> 
    ggplot(aes(x = Theta, y = Probability)) +
    geom_line(aes(y = Information), linetype = "dotted", color = "grey") + 
    geom_line(aes(color = Answer)) + 
    geom_vline(xintercept = x, linetype = "dashed") +
    facet_grid(~Item) + 
    theme_modern() 
  list(p = p, rez = rez)
}

out <- crc(mod1, data, x = c(-1.25, 2.25))
out$p + labs(x = expression(Depression~(theta))) 
```

## Normalized scoring

```{r message=FALSE, warning=FALSE}
item4 <- out$rez |> 
  filter(Item == "Depression_4") |> 
  filter(Theta >= -1.5 & Theta <= 2.25) |> 
  mutate(Theta = normalize(Theta))

scores <- item4 |> 
  group_by(Answer) |> 
  filter(Probability == max(Probability)) |> 
  ungroup()

item4 |> 
  ggplot(aes(x = Theta, y = Probability, color = Answer)) +
  geom_line() + 
  theme_modern() 
```

# Anxiety

```{r message=FALSE, warning=FALSE}
data <- select(df, contains("Anxiety"))

m_anxiety <- mirt(data, model = 1, itemtype = 'graded', SE = TRUE)

# M2(m_anxiety, type = "C2", calcNULL = FALSE)

itemfit(m_anxiety)

# IRT parameters
IRT_parms <- coef(m_anxiety, IRTpars = TRUE, simplify = TRUE)
IRT_parms$items

# Factor
summary(m_anxiety)

# Plots
plot(m_anxiety, type = 'score', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'infoSE', theta_lim = c(-3, 3))
plot(m_anxiety, type = 'rxx', theta_lim = c(-3, 3))
```

## Category Characteristic Curves (CRC) and Item Information Curves

```{r message=FALSE, warning=FALSE}
out <- crc(m_anxiety, data, x = c(-1.25, 2.25))
out$p + labs(x = expression(Anxiety~(theta))) 
```

### 2023-01-07 analysis.Rmd
manuscript

@@
 
+https://aidenloe.github.io/introToIRT.html
+
 # Data
+
+The data has been collected during the study of . The dataset is available in open-access at ...
 
@@
 
-df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |> 
+df <- read.csv("../IllusionGameReliability/data/preprocessed_questionnaires.csv") |> 
+# df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |> 
   select(starts_with("Item_PHQ4")) |> 
@@
 
-# Depression
+# Distribution
+
+```{r distribution, message=FALSE, warning=FALSE, fig.width=fig.height}
+data <- df |> 
+  pivot_longer(everything(), names_to = "Item", values_to = "Answer") |> 
+  group_by(Item, Answer) |> 
+  summarise(n = n() / nrow(df)) |> 
+  mutate(
+    Facet = str_split_fixed(Item, "_", 2)[, 1],
+    # Item = str_split_fixed(Item, "_", 2)[,2],
+    Answer = case_when(
+      Answer == 0 ~ "Not at all",
+      Answer == 1 ~ "Once or twice",
+      Answer == 2 ~ "Several days",
+      Answer == 3 ~ "More than half the days",
+      TRUE ~ "Nearly every day"
+    ),
+    Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")),
+    Item = case_when(
+      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
+      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
+      Item == "Depression_3" ~ "D1: Little interest or pleasure in doing things",
+      TRUE ~ "D2: Feeling down, depressed, or hopeless"
+    ))
+
+p1 <- data |> 
+  ggplot(aes(x = Answer, fill=Facet)) +
+  geom_bar(aes(y = n), stat = "identity") +
+  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
+  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
+  labs(y = "Proportion of Answers", title = "A. Proportion of answers per item") +
+  theme_modern(axis.title.space = 10) +
+  theme(axis.text.x = element_text(angle = 45, hjust=1),
+        axis.title.x = element_blank(),
+        plot.title = element_text(face = "bold"),
+        strip.background=element_rect(fill="#EEEEEE", colour="white")) +
+  facet_wrap(~Item)
+p1
+```
+
+The "Once or twice" answer was selected in `r insight::format_percent(mean(filter(data, Answer == "Once or twice")$n))`
+
+# Anxiety
+
+## Descriptive
+```{r message=FALSE, warning=FALSE}
+df_anx <- select(df, contains("Anxiety"))
+
+performance::cronbachs_alpha(df_anx)
+```
+
+## Model
 
 ```{r message=FALSE, warning=FALSE}
-# s <- 'Anxiety = 1,2
-#       Depression = 3,4
-#       COV = Anxiety*Depression'
-# mod1 <- mirt(df, model = mirt.model(s), itemtype = 'graded', SE = TRUE)
+m_anxiety <- mirt(df_anx, model = 1, itemtype = 'graded', SE = TRUE)
 
-data <- select(df, contains("Depression"))
+m_anxiety
 
-mod1 <- mirt(data, model = 1, itemtype = 'graded', SE = TRUE)
+# Factor loadings
+summary(m_anxiety)
 
-# M2(mod1, type = "C2", calcNULL = FALSE)
+# Alpha
+a <- coef(m_anxiety)
+a <- data.frame(
+  Item = c("Anxiety 1", "Anxiety 2"),
+  Discrimination = c(paste0(format_value(a$Anxiety_1[1, 1]), ", ", format_ci(a$Anxiety_1[2, 1], a$Anxiety_1[3, 1])), paste0(format_value(a$Anxiety_2[1, 1]), ", ", format_ci(a$Anxiety_2[2, 1], a$Anxiety_2[3, 1]))
+))
+knitr::kable(a)
+```
 
-itemfit(mod1)
-
-# IRT parameters
-IRT_parms <- coef(mod1, IRTpars = TRUE, simplify = TRUE)
-IRT_parms$items
-
-# Factor
-summary(mod1)
-
+```{r message=FALSE, warning=FALSE, eval=FALSE}
 # Plots
-plot(mod1, type = 'trace', theta_lim = c(-3, 3))
-plot(mod1, type = 'infotrace', theta_lim = c(-3, 3))
-plot(mod1, type = 'score', theta_lim = c(-3, 3))
-plot(mod1, type = 'infoSE', theta_lim = c(-3, 3))
-plot(mod1, type = 'rxx', theta_lim = c(-3, 3))
+plot(m_anxiety, type = 'trace', theta_lim = c(-3, 3))
+plot(m_anxiety, type = 'infotrace', theta_lim = c(-3, 3))
+plot(m_anxiety, type = 'score', theta_lim = c(-3, 3))
+plot(m_anxiety, type = 'infoSE', theta_lim = c(-3, 3))
+plot(m_anxiety, type = 'rxx', theta_lim = c(-3, 3))
 ```
@@
 
-```{r message=FALSE, warning=FALSE}
-crc <- function(mod, data, x = c(-1.25, 2.25)) {
-  Theta <- matrix(seq(-3, 3, by = .01))
+```{r CRC_anxiety, message=FALSE, warning=FALSE}
+crc <- function(mod, data) {
+  Theta <- matrix(seq(-3.5, 3.5, by = .01))
   rez <- data.frame()
   for(i in 1:2) {
-    dat <- as.data.frame(probtrace(extract.item(mod1, i), Theta))
+    dat <- as.data.frame(probtrace(extract.item(mod, i), Theta))
     dat$Theta <- Theta[,1]
-    dat$Information <- normalize(iteminfo(extract.item(mod1, i), Theta))
+    dat$Information <- normalize(iteminfo(extract.item(mod, i), Theta))
     dat <- pivot_longer(dat, -one_of(c("Theta", "Information")), names_to = "Answer", values_to = "Probability")
@@
       TRUE ~ "Nearly every day"
-      )) |> 
-    mutate(
-      Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")))
+      ),
+      Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")),
+    Item = case_when(
+      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
+      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
+      Item == "Depression_3" ~ "D1: Little interest or pleasure in doing things",
+      TRUE ~ "D2: Feeling down, depressed, or hopeless"
+    ))
+
+  # Get close to 0.95
+  i1 <- rez[rez$Item == unique(rez$Item)[1],]
+  i2 <- rez[rez$Item == unique(rez$Item)[2],]
+  minmax <- rbind(
+    rbind(
+      i1[i1$Answer == "Not at all", ][which.min(abs(i1[i1$Answer == "Not at all", ]$Probability - 0.95)), ],
+      i1[i1$Answer == "Nearly every day", ][which.min(abs(i1[i1$Answer == "Nearly every day", ]$Probability - 0.95)), ]
+      ),
+    rbind(
+      i2[i2$Answer == "Not at all", ][which.min(abs(i2[i2$Answer == "Not at all", ]$Probability - 0.95)), ],
+      i2[i2$Answer == "Nearly every day", ][which.min(abs(i2[i2$Answer == "Nearly every day", ]$Probability - 0.95)), ]
+      )
+  )
 
@@
     geom_line(aes(y = Information), linetype = "dotted", color = "grey") +
-    geom_line(aes(color = Answer)) + 
-    geom_vline(xintercept = x, linetype = "dashed") +
+    geom_line(aes(color = Answer), size=1) +
+    # geom_vline(data = minmax, aes(xintercept = Theta), linetype = "dashed") +
+    scale_y_continuous(labels=scales::percent, expand = c(0, 0.005)) +
+    scale_color_flat_d("rainbow") +
     facet_grid(~Item) +
-    theme_modern() 
-  list(p = p, rez = rez)
+    theme_modern(axis.title.space = 5) +
+    theme(strip.background=element_rect(fill="#EEEEEE", colour="white"))
+  list(p = p, rez = rez, minmax=minmax)
 }
 
-out <- crc(mod1, data, x = c(-1.25, 2.25))
-out$p + labs(x = expression(Depression~(theta))) 
+out <- crc(m_anxiety, df_anx)
+p2a <- out$p + labs(x = expression(Anxiety~(theta)))
+p2a
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-item4 <- out$rez |> 
-  filter(Item == "Depression_4") |> 
-  filter(Theta >= -1.5 & Theta <= 2.25) |> 
-  mutate(Theta = normalize(Theta))
+normalize_scores <- function(data, minmax) {
 
-scores <- item4 |> 
+  minmax <- minmax[minmax$Theta %in% range(minmax$Theta), ]
+
+  item <- data.frame()
+  scores <- data.frame()
+  for(i in unique(data$Item)) {
+    item <- data[data$Item == i, ] |>
+      filter(Theta >= min(minmax$Theta) & Theta <= max(minmax$Theta)) |>
+      mutate(Theta_n = normalize(Theta)) |> 
+      rbind(item)
+
+    scores <- item[item$Item == i, ] |>
       group_by(Answer) |>
       filter(Probability == max(Probability)) |>
-  ungroup()
+      ungroup() |>
+      mutate(label = insight::format_value(Theta_n)) |> 
+      rbind(scores)
+  }
 
-item4 |> 
+  item |>
     ggplot(aes(x = Theta, y = Probability, color = Answer)) +
-  geom_line() + 
-  theme_modern() 
+    geom_line(size = 1) +
+    geom_segment(data=scores, aes(x=Theta, xend = Theta, y=0, yend=Probability, color = Answer), linetype = "dashed") +
+    geom_label(data=scores, aes(x=Theta, y=Probability, label=label), show.legend = FALSE) +
+    scale_y_continuous(labels=scales::percent, expand = c(0, 0.01)) +
+    scale_color_viridis_d(option = "D") +
+    facet_grid(~Item) +
+    theme_modern(axis.title.space = 5) +
+    theme(strip.background=element_rect(fill="#EEEEEE", colour="white"))
+}
+
+p3a <- normalize_scores(data=out$rez, out$minmax) + labs(x = expression(Anxiety~(theta)))
+p3a 
 ```
 
-# Anxiety
+# Depression
+
+## Descriptive
+```{r message=FALSE, warning=FALSE}
+df_dep <- select(df, contains("Depression"))
+
+performance::cronbachs_alpha(df_dep)
+```
+
+## Model
 
 ```{r message=FALSE, warning=FALSE}
-data <- select(df, contains("Anxiety"))
+m_depression <- mirt(df_dep, model = 1, itemtype = 'graded', SE = TRUE)
 
-m_anxiety <- mirt(data, model = 1, itemtype = 'graded', SE = TRUE)
+m_depression
 
-# M2(m_anxiety, type = "C2", calcNULL = FALSE)
+# Factor loadings
+summary(m_depression)
 
-itemfit(m_anxiety)
-
-# IRT parameters
-IRT_parms <- coef(m_anxiety, IRTpars = TRUE, simplify = TRUE)
-IRT_parms$items
-
-# Factor
-summary(m_anxiety)
-
-# Plots
-plot(m_anxiety, type = 'score', theta_lim = c(-3, 3))
-plot(m_anxiety, type = 'infoSE', theta_lim = c(-3, 3))
-plot(m_anxiety, type = 'rxx', theta_lim = c(-3, 3))
+# Alpha
+a2 <- coef(m_depression)
+a2 <- data.frame(
+  Item = c("Depression 1", "Depression 2"),
+  Discrimination = c(paste0(format_value(a2$Depression_3[1, 1]), ", ", format_ci(a2$Depression_3[2, 1], a2$Depression_3[3, 1])), paste0(format_value(a2$Depression_4[1, 1]), ", ", format_ci(a2$Depression_4[2, 1], a2$Depression_4[3, 1]))
+))
+knitr::kable(a2)
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-out <- crc(m_anxiety, data, x = c(-1.25, 2.25))
-out$p + labs(x = expression(Anxiety~(theta))) 
+out <- crc(m_depression, df_dep)
+p2b <- out$p + labs(x = expression(Depression~(theta)))
+p2b
 ```
+
+## Normalized scoring
+
+```{r message=FALSE, warning=FALSE}
+p3b <- normalize_scores(out$rez, out$minmax) + labs(x = expression(Depression~(theta)))
+p3b 
+```
+
+```{r message=FALSE, warning=FALSE}
+fig1 <- wrap_elements(p1) + wrap_elements(p2a / p2b + plot_annotation(title = "B.
+Information Curves per Response Type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))
+
+ggsave("figures/figure1.png", fig1, width = fig.width*1.43, height=fig.height*1.43)
+
+fig2 <- p3a / p3b + plot_annotation(title = "Normalized Scoring", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")
+
+ggsave("figures/figure2.png", fig2, width = fig.width*1.43, height=fig.height*1.43)
+```
+
+# Correlation
+
+```{r message=FALSE, warning=FALSE}
+df$Depression <- mirt::fscores(m_depression)[, 1]
+df$Anxiety <- mirt::fscores(m_anxiety)[, 1]
+
+correlation::correlation(df) |>
+  summary() |> 
+  export_table(caption = "Correlation Matrix")
+```
+
+<!-- # Multifactor model -->
+
+<!-- ```{r message=FALSE, warning=FALSE, eval=FALSE} -->
+<!-- s <- 'Anxiety = 1,2 -->
+<!--       Depression = 3,4 -->
+<!--       COV = Anxiety*Depression' -->
+<!-- mod <- mirt(df, model = mirt.model(s), itemtype = 'graded', SE = TRUE) -->
+<!-- mod <- mirt(df, model = 2, itemtype = 'graded', SE = TRUE) -->
+
+<!-- summary(mod) -->
+
+<!-- # Plots -->
+<!-- plot(mod, type = 'trace', theta_lim = c(-3, 3)) -->
+<!-- plot(mod, type = 'infotrace', theta_lim = c(-3, 3)) -->
+<!-- plot(mod, type = 'score', theta_lim = c(-3, 3)) -->
+
+<!-- Theta <- as.matrix(expand.grid(seq(-3, 3, by = 1), seq(-3, 3, by = 1))) -->
+<!-- rez <- data.frame() -->
+<!-- for(i in 1:4) { -->
+<!--   dat <- as.data.frame(probtrace(extract.item(mod, i), Theta)) -->
+<!--   dat$Theta_F1 <- Theta[, 1] -->
+<!--   dat$Theta_F2 <- Theta[, 2] -->
+<!--   # dat$Information <- normalize(iteminfo(extract.item(mod, i), Theta)) -->
+<!--   dat <- pivot_longer(dat, -one_of(c("Theta_F1", "Theta_F2")), names_to = "Answer", values_to = "Probability") -->
+<!--   dat$Item <- names(df)[i] -->
+<!--   rez <- rbind(rez, dat) -->
+<!-- } -->
+
+<!-- rez <- rez |> -->
+<!--   separate(Item, into = c("Dimension", "Item_Number")) |> -->
+<!--   mutate( -->
+<!--     Answer = case_when( -->
+<!--     Answer == "P.1" ~ "Not at all", -->
+<!--     Answer == "P.2" ~ "Once or twice", -->
+<!--     Answer == "P.3" ~ "Several days", -->
+<!--     Answer == "P.4" ~ "More than half the days", -->
+<!--     TRUE ~ "Nearly every day" -->
+<!--     )) |> -->
+<!--   mutate( -->
+<!--     Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day"))) -->
+
+<!-- rez |> -->
+<!--   ggplot(aes(x = Theta_F1, y = Probability)) + -->
+<!--   # geom_line(aes(y = Information), linetype = "dotted", color = "grey") + -->
+<!--   geom_line(aes(color = Answer)) + -->
+<!--   geom_vline(xintercept = c(-1.25, 2.25), linetype = "dashed") + -->
+<!--   facet_grid(Dimension~Item_Number) + -->
+<!--   theme_modern() -->
+<!-- ``` -->

### 2023-01-07 paper/manuscript.Rmd
manuscript

---
title             : "**Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability**"
shorttitle        : "PHQ-4R"
author:
  - name          : "Dominique Makowski"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "HSS 04-18, 48 Nanyang Avenue, Singapore"
    email         : "dom.makowski@gmail.com"
    orcid         : 0000-0001-5375-9967
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Data curation"
      - "Formal Analysis"
      - "Funding acquisition"
      - "Investigation"
      - "Methodology"
      - "Project administration"
      - "Resources"
      - "Software"
      - "Supervision"
      - "Validation"
      - "Visualization"
      - "Writing – original draft"
  - name          : "An Shu Te"
    affiliation   : "1"
    orcid         : 0000-0001-8008-2824
    role:
      - "Project administration"
      - "Resources"
      - "Investigation"
      - "Writing – original draft"
  - name          : "S.H. Annabel Chen"
    affiliation   : "1, 2, 3, 4"
    orcid         : 0000-0002-1540-5516
    role:
      - "Project administration"
      - "Supervision"
      - "Writing – review & editing"
affiliation:

  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "LKC Medicine, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "National Institute of Education, Singapore"
  - id            : "4"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
# authornote: |

abstract: |
  Something something

  **Significance Statement**. Developing reliable and sensitive instruments for mood disorders screening is critical in a global context marked by international crises (pandemics, wars), where more and more surveys as done online. In this study, we show that a small modification to the widely used PHQ-4 scale (adding the "Once or twice" response option) can increase its ability to capture the mild fluctuations prevalent in subclinical samples.

keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
wordcount         : "4085"
bibliography      : references.bib
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : yes
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
csl: utils/apa.csl
header-includes:
  - \usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
# options and parameters
options(digits = 3)

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.width = see::golden_ratio(9),
  fig.height = 9,
  fig.path = "../figures/"
)

cache <- TRUE
```

# Introduction

The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety [@kroenke2009ultra]. It consists of two items for depression (PHQ–2) and anxiety (GAD–2), each corresponding to DSM-IV Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD). It has been validated across many languages and samples [@christodoulaki2022validation, @mendoza2022factor, @materu2020psychometric], becoming one of the most popular screening instrument [@maurer2018depression].

While the scale has been validated and used in the general population [@lowe20104, @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients. This diagnostic and discriminative goal materializes in the scale's design and the existence of categorical cut-offs. However, it might not be best suited to capture subclinical variability. In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and leaves out the possibility of more subtle occurrences. While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of variability quantifications. The goal of this study is to increase, with minimal changes to the scale, the sensitivity to very mild mood alterations.

# Methods

## Original Scale

The instructions "Over the last 2 weeks, how often have you been bothered by the following problems?" are followed with 4 items (A1 - Feeling nervous, anxious or on edge; A2 - Not being able to stop or control worrying; D1 - Little interest or pleasure in doing things; D2 - Feeling down, depressed, or hopeless). The original answer options are "Not at all" (0), "Several days" (1), "More than half the days" (2), "Nearly every day" (3). The total score is computed by summing the responses of each facet.

## Revision

In order to better capture potential mild mood inflections without altering the scale scoring or structure, we added a "Once or twice" option between "Not at all" and "Several days" [see @dobson1979equidistant for the choice of the label].

## Participants

The sample consists of 500 English-speaking participants (**stats**) who were administered the revised PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.

This study was approved by the NTU Institutional Review Board (NTU **NUMBER**). All participants provided their informed consent prior to participation and were incentivized after completing the study.

# Results

The fully reproducible analysis script is accessible at https://github.com/DominiqueMakowski/PHQ4R.

## Descriptive

The reliability of the anxiety ($Cronbach's~\alpha = 0.892$; RMSEA = 0.032) and depression ($Cronbach's~\alpha = 0.822$; RMSEA = 0.032) subscales was excellent. The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option was the most prevalent response for almost all items (on average selected in **30.07\%** of cases).

## Item Response Theory

Item Response Theory (IRT) provides insights into how well items and responses capture the underlying latent trait $\theta$. For each of the subscales, we fitted a unidimensional graded response model [GRM, @samejima1997graded]. For anxiety, the two items captured 88.1\% of the variance of the latent anxiety dimension ($\theta_{anxiety}$). The discrimination parameters suggested that the first item was less precise ($\alpha = 3.20$) than the second item ($\alpha = 12.53$) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a larger range of $\theta_{anxiety}$, as can be seen in **Figure 1B**). The two depression items captured 80.6\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 15.69$) than the first ($\alpha = 2.19$). Importantly to our objective, it seems that the added "Once or twice" option does cover a selective portion of the latent space.

```{r message=FALSE, warning=FALSE, fig.cap="A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).", out.width="100%"}
knitr::include_graphics("../figures/figure1.png")
```

## Scoring

We propose two types of scoring procedures. The first aims at
<!-- In order to  -->
minimally disrupting the original scale and making its scores comparable, enabling comparisons across studies and the usage of developed cut-offs and norms. We suggest to score the new option, "Once or twice", as 0.5, occupying the middle space between 0 and 1 (preserving this way the total score range of 0 - 8 per dimension).

The second scoring method we propose takes into account the non-linear spacing between responses, as indicated by the peak of their measurement information. To develop it, we found for each dimension the lower and higher value of the latent trait $\theta$ for which the probability lower and higher probability of identification is 95\%. We considered these points at 0 and 1, and normalized the location of the peaks of each response in between.

To illustrate, a person answering "Several days" and "Once or twice" ro the first and second item of the anxiety subscale would have a score of 1.5 (1 + 0.5 following the basic scoring) and a normalized score of `r insight::format_value(mean(c(0.56, 0.45)))` ($(0.56 + 0.45) / 2$).

```{r message=FALSE, warning=FALSE, fig.cap="Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).", out.width="100%"}
knitr::include_graphics("../figures/figure2.png")
```

```{r message=FALSE, warning=FALSE, out.width="100%"}
papaja::apa_table(
  data.frame(
    Facet = c("\\textbf{Anxiety}", "", "", "", "", "", "", "", "", "", "\\textbf{Depression}", "", "", "", "", "", "", "", "", ""),
    Item = c("Feeling nervous, anxious or on edge", "", "", "", "", "Not being able to stop or control worrying", "", "", "", "", "Little interest or pleasure in doing things", "", "", "", "", "Feeling down, depressed, or hopeless", "", "", "", ""),
    Response = paste0("\\textit{", rep(c("Not at all", "Once or twice", "Several days", "More than half the days", "Nearly every day"), 4), "}"),
    Basic_Scoring = rep(c("0", "0.5", "1", "2", "3"), 4),
    Normalized_Scoring = rep(c("0", "0.5", "0.7", "0.8", "1"), 4)
    ),
  caption = "Revised Patient Health Questionnaire-4 (PHQ-4R). The instructions are \"Over the last 2 weeks, how often have you been bothered by the following problems?\". The 0.5 scoring is introduced to preserve the compatibility with the original version, and enables to use its norms and cut-offs. The normalized scoring takes into account the non-linear spacing between item responses.",  
  escape = FALSE,
  font_size = "scriptsize",
  col.names = c("Facet", "Item", "Response", 
                "Basic Scoring", 
                "Normalized Scoring",
                "Scoring"),
  midrules = c(5, 10, 15))
```

# Discussion

The objective of this study was to test the introduction of a "Once or twice" response option to the PHQ-4 to make it more sensitive to milder fluctuations. The fact that this new response option was the most prevalent is in itself evidence for its usefulness, and IRT analysis further revealed how this response captures with precision a unique portion of anxiety and depression. Our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).

# Data Availability

The dataset analysed during the current study are available in the GitHub repository https://github.com/DominiqueMakowski/PHQ4R

# Funding

This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2023-01-07 paper/manuscript.tex
manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4085}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability},
  pdfauthor={Dominique Makowski1, An Shu Te1, \& S.H. Annabel Chen1, 2, 3, 4},
  pdflang={en-EN},
  pdfkeywords={visual illusions, illusion game, Pyllusion, personality, general factor},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability}}
\author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{1}, \& S.H. Annabel Chen\textsuperscript{1, 2, 3, 4}}
\date{}

\shorttitle{PHQ-4R}

\authornote{

The authors made the following contributions. Dominique Makowski: Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Software, Supervision, Validation, Visualization, Writing -- original draft; An Shu Te: Project administration, Resources, Investigation, Writing -- original draft; S.H. Annabel Chen: Project administration, Supervision, Writing -- review \& editing.

Correspondence concerning this article should be addressed to Dominique Makowski, HSS 04-18, 48 Nanyang Avenue, Singapore. E-mail: \href{mailto:dom.makowski@gmail.com}{\nolinkurl{dom.makowski@gmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Social Sciences, Nanyang Technological University, Singapore\\\textsuperscript{2} LKC Medicine, Nanyang Technological University, Singapore\\\textsuperscript{3} National Institute of Education, Singapore\\\textsuperscript{4} Centre for Research and Development in Learning, Nanyang Technological University, Singapore}

\abstract{%
Something something

\textbf{Significance Statement}. Developing reliable and sensitive instruments for mood disorders screening is critical in a global context marked by international crises (pandemics, wars), where more and more surveys as done online. In this study, we show that a small modification to the widely used PHQ-4 scale (adding the ``Once or twice'' response option) can increase its ability to capture the mild fluctuations prevalent in subclinical samples.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety (Kroenke et al., 2009). It consists of two items for depression (PHQ--2) and anxiety (GAD--2), each corresponding to DSM-IV Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD). It has been validated across many languages and samples Materu et al. (2020), becoming one of the most popular screening instrument (Maurer et al., 2018).

While the scale has been validated and used in the general population Hajek \& König (2020), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients. This diagnostic and discriminative goal materializes in the scale's design and the existence of categorical cut-offs. However, it might not be best suited to capture subclinical variability. In particular, the gap between the two lowest possible answers, ``Not at all'' and ``Several days'', is quite large and leaves out the possibility of more subtle occurrences. While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of variability quantifications. The goal of this study is to increase, with minimal changes to the scale, the sensitivity to very mild mood alterations.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{original-scale}{%
\subsection{Original Scale}\label{original-scale}}

The instructions ``Over the last 2 weeks, how often have you been bothered by the following problems?'' are followed with 4 items (A1 - Feeling nervous, anxious or on edge; A2 - Not being able to stop or control worrying; D1 - Little interest or pleasure in doing things; D2 - Feeling down, depressed, or hopeless). The original answer options are ``Not at all'' (0), ``Several days'' (1), ``More than half the days'' (2), ``Nearly every day'' (3). The total score is computed by summing the responses of each facet.

\hypertarget{revision}{%
\subsection{Revision}\label{revision}}

In order to better capture potential mild mood inflections without altering the scale scoring or structure, we added a ``Once or twice'' option between ``Not at all'' and ``Several days'' (see Dobson \& Mothersill, 1979 for the choice of the label).

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

The sample consists of 500 English-speaking participants (\textbf{stats}) who were administered the revised PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.

This study was approved by the NTU Institutional Review Board (NTU \textbf{NUMBER}). All participants provided their informed consent prior to participation and were incentivized after completing the study.

\hypertarget{results}{%
\section{Results}\label{results}}

The fully reproducible analysis script is accessible at \url{https://github.com/DominiqueMakowski/PHQ4R}.

\hypertarget{descriptive}{%
\subsection{Descriptive}\label{descriptive}}

The reliability of the anxiety (\(Cronbach's~\alpha = 0.892\); RMSEA = 0.032) and depression (\(Cronbach's~\alpha = 0.822\); RMSEA = 0.032) subscales was excellent. The proportion of response types stratified by item (see \textbf{Figure 1A}) shows that the new ``Once or twice'' option was the most prevalent response for almost all items (on average selected in \textbf{30.07\%} of cases).

\hypertarget{item-response-theory}{%
\subsection{Item Response Theory}\label{item-response-theory}}

Item Response Theory (IRT) provides insights into how well items and responses capture the underlying latent trait \(\theta\). For each of the subscales, we fitted a unidimensional graded response model (GRM, Samejima, 1997). For anxiety, the two items captured 88.1\% of the variance of the latent anxiety dimension (\(\theta_{anxiety}\)). The discrimination parameters suggested that the first item was less precise (\(\alpha = 3.20\)) than the second item (\(\alpha = 12.53\)) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a larger range of \(\theta_{anxiety}\), as can be seen in \textbf{Figure 1B}). The two depression items captured 80.6\% of the variance of its latent trait (\(\theta_{depression}\)), and the opposite pattern was found: the first item had a higher precision (\(\alpha = 15.69\)) than the first (\(\alpha = 2.19\)).

Importantly to our objective,

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/figure1} \caption{A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).}\label{fig:unnamed-chunk-2}
\end{figure}

\hypertarget{normalized-scoring}{%
\subsection{Normalized Scoring}\label{normalized-scoring}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The objective of this study was to test the introduction of a ``Once or twice'' response option to the PHQ-4 to make it more sensitive to milder fluctuations.

The fact that the new response option was the most prevalent is in itself a testimony to its usefuleness, suggesting an increased potential to discriminate a ``normal'' mood level from lower-end extremes (the true zero scores).

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-3}Revised Patient Health Questionnaire-4 (PHQ-4R). The instructions are "Over the last 2 weeks, how often have you been bothered by the following problems?". The basic scoring is introduced to preserve the compatibility with the original version, and enables to use its norms and cut-offs. The normalized scoring takes into account the non-linear spacing between items.}

\scriptsize{

\begin{tabular}{lllll}
\toprule
Facet & Item & Response & Basic Scoring & Normalized Scoring\\
\midrule
\textbf{Anxiety} & Feeling nervous, anxious or on edge & \textit{Not at all} & 0 & 0\\
 &  & \textit{Once or twice} & 0.5 & 0.5\\
 &  & \textit{Several days} & 1 & 0.7\\
 &  & \textit{More than half the days} & 2 & 0.8\\
 &  & \textit{Nearly every day} & 3 & 1\\ \midrule
 & Not being able to stop or control worrying & \textit{Not at all} & 0 & 0\\
 &  & \textit{Once or twice} & 0.5 & 0.5\\
 &  & \textit{Several days} & 1 & 0.7\\
 &  & \textit{More than half the days} & 2 & 0.8\\
 &  & \textit{Nearly every day} & 3 & 1\\ \midrule
\textbf{Depression} & Little interest or pleasure in doing things & \textit{Not at all} & 0 & 0\\
 &  & \textit{Once or twice} & 0.5 & 0.5\\
 &  & \textit{Several days} & 1 & 0.7\\
 &  & \textit{More than half the days} & 2 & 0.8\\
 &  & \textit{Nearly every day} & 3 & 1\\ \midrule
 & Feeling down, depressed, or hopeless & \textit{Not at all} & 0 & 0\\
 &  & \textit{Once or twice} & 0.5 & 0.5\\
 &  & \textit{Several days} & 1 & 0.7\\
 &  & \textit{More than half the days} & 2 & 0.8\\
 &  & \textit{Nearly every day} & 3 & 1\\
\bottomrule
\end{tabular}

}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{data-availability}{%
\section{Data Availability}\label{data-availability}}

The dataset analysed during the current study are available in the GitHub repository \url{https://github.com/DominiqueMakowski/PHQ4R}

\hypertarget{funding}{%
\section{Funding}\label{funding}}

This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-christodoulaki2022validation}{}}%
Christodoulaki, A., Baralou, V., Konstantakopoulos, G., \& Touloumi, G. (2022). Validation of the patient health questionnaire-4 (PHQ-4) to screen for depression and anxiety in the greek general population. \emph{Journal of Psychosomatic Research}, \emph{160}, 110970.

\leavevmode\vadjust pre{\hypertarget{ref-dobson1979equidistant}{}}%
Dobson, K. S., \& Mothersill, K. J. (1979). Equidistant categorical labels for construction of likert-type scales. \emph{Perceptual and Motor Skills}, \emph{49}(2), 575--580.

\leavevmode\vadjust pre{\hypertarget{ref-hajek2020prevalence}{}}%
Hajek, A., \& König, H.-H. (2020). Prevalence and correlates of individuals screening positive for depression and anxiety on the phq-4 in the german general population: Findings from the nationally representative german socio-economic panel (GSOEP). \emph{International Journal of Environmental Research and Public Health}, \emph{17}(21), 7865.

\leavevmode\vadjust pre{\hypertarget{ref-kroenke2009ultra}{}}%
Kroenke, K., Spitzer, R. L., Williams, J. B., \& Löwe, B. (2009). An ultra-brief screening scale for anxiety and depression: The PHQ--4. \emph{Psychosomatics}, \emph{50}(6), 613--621.

\leavevmode\vadjust pre{\hypertarget{ref-lowe20104}{}}%
Löwe, B., Wahl, I., Rose, M., Spitzer, C., Glaesmer, H., Wingenfeld, K., Schneider, A., \& Brähler, E. (2010). A 4-item measure of depression and anxiety: Validation and standardization of the patient health questionnaire-4 (PHQ-4) in the general population. \emph{Journal of Affective Disorders}, \emph{122}(1-2), 86--95.

\leavevmode\vadjust pre{\hypertarget{ref-materu2020psychometric}{}}%
Materu, J., Kuringe, E., Nyato, D., Galishi, A., Mwanamsangu, A., Katebalila, M., Shao, A., Changalucha, J., Nnko, S., \& Wambura, M. (2020). The psychometric properties of PHQ-4 anxiety and depression screening scale among out of school adolescent girls and young women in tanzania: A cross-sectional study. \emph{BMC Psychiatry}, \emph{20}(1), 1--8.

\leavevmode\vadjust pre{\hypertarget{ref-maurer2018depression}{}}%
Maurer, D. M., Raymond, T. J., \& Davis, B. N. (2018). Depression: Screening and diagnosis. \emph{American Family Physician}, \emph{98}(8), 508--515.

\leavevmode\vadjust pre{\hypertarget{ref-mendoza2022factor}{}}%
Mendoza, N. B., Frondozo, C. E., Dizon, J. I. W. T., \& Buenconsejo, J. U. (2022). The factor structure and measurement invariance of the PHQ-4 and the prevalence of depression and anxiety in a southeast asian context amid the COVID-19 pandemic. \emph{Current Psychology}, 1--10.

\leavevmode\vadjust pre{\hypertarget{ref-samejima1997graded}{}}%
Samejima, F. (1997). Graded response model. In \emph{Handbook of modern item response theory} (pp. 85--100). Springer.

\end{CSLReferences}

\clearpage
\renewcommand{\listfigurename}{Figure captions}

\end{document}

### 2023-01-07 analysis.Rmd
readme

@@
 
-See [**tutorial**](https://bookdown.org/bean_jerry/using_r_for_social_work_research/item-response-theory.html) on Item Response Theory (IRT).
+<!-- See [**tutorial**](https://bookdown.org/bean_jerry/using_r_for_social_work_research/item-response-theory.html) on Item Response Theory (IRT). -->
 
-https://aidenloe.github.io/introToIRT.html
+<!-- https://aidenloe.github.io/introToIRT.html -->
 

### 2023-01-07 paper/manuscript.Rmd
readme

@@
 
-The sample consists of 500 English-speaking participants (**stats**) who were administered the revised PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
+The sample consists of 500 English-speaking participants (**stats**) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
 
@@
     ),
-  caption = "Revised Patient Health Questionnaire-4 (PHQ-4R).
+  caption = "Refined Patient Health Questionnaire-4 (PHQ-4R).
 The instructions are \"Over the last 2 weeks, how often have you been bothered by the following problems?\".
@@
                 "Basic Scoring", 
-                "Normalized Scoring",
-                "Scoring"),
+                "Normalized Scoring"),
   midrules = c(5, 10, 15))

### 2023-01-07 paper/manuscript.tex
update

@@
 
-The sample consists of 500 English-speaking participants (\textbf{stats}) who were administered the revised PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
+The sample consists of 500 English-speaking participants (\textbf{stats}) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
 
@@
 The two depression items captured 80.6\% of the variance of its latent trait (\(\theta_{depression}\)), and the opposite pattern was found: the first item had a higher precision (\(\alpha = 15.69\)) than the first (\(\alpha = 2.19\)).
-
-Importantly to our objective,
+Importantly to our objective, it seems that the added ``Once or twice'' option does cover a selective portion of the latent space.
 
@@
 
-\hypertarget{normalized-scoring}{%
-\subsection{Normalized Scoring}\label{normalized-scoring}}
+\hypertarget{scoring}{%
+\subsection{Scoring}\label{scoring}}
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+We propose two types of scoring procedures. The first aims at
+minimally disrupting the original scale and making its scores comparable, enabling comparisons across studies and the usage of developed cut-offs and norms.
+We suggest to score the new option, ``Once or twice'', as 0.5, occupying the middle space between 0 and 1 (preserving this way the total score range of 0 - 8 per dimension).
 
-The objective of this study was to test the introduction of a ``Once or twice'' response option to the PHQ-4 to make it more sensitive to milder fluctuations.
+The second scoring method we propose takes into account the non-linear spacing between responses, as indicated by the peak of their measurement information.
+To develop it, we found for each dimension the lower and higher value of the latent trait \(\theta\) for which the probability lower and higher probability of identification is 95\%.
+We considered these points at 0 and 1, and normalized the location of the peaks of each response in between.
 
-The fact that the new response option was the most prevalent is in itself a testimony to its usefuleness, suggesting an increased potential to discriminate a ``normal'' mood level from lower-end extremes (the true zero scores).
+To illustrate, a person answering ``Several days'' and ``Once or twice'' ro the first and second item of the anxiety subscale would have a score of 1.5 (1 + 0.5 following the basic scoring) and a normalized score of 0.50 (\((0.56 + 0.45) / 2\)).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{../figures/figure2} \caption{Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).}\label{fig:unnamed-chunk-3}
+\end{figure}
 
@@
 
-\caption{\label{tab:unnamed-chunk-3}Revised Patient Health Questionnaire-4 (PHQ-4R).
+\caption{\label{tab:unnamed-chunk-4}Refined Patient Health Questionnaire-4 (PHQ-4R).
 The instructions are "Over the last 2 weeks, how often have you been bothered by the following problems?".
-The basic scoring is introduced to preserve the compatibility with the original version, and enables to use its norms and cut-offs.
-The normalized scoring takes into account the non-linear spacing between items.}
+The 0.5 scoring is introduced to preserve the compatibility with the original version, and enables to use its norms and cut-offs.
+The normalized scoring takes into account the non-linear spacing between item responses.}
 
@@
 \end{table}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+The objective of this study was to test the introduction of a ``Once or twice'' response option to the PHQ-4 to make it more sensitive to milder fluctuations.
+The fact that this new response option was the most prevalent is in itself evidence for its usefulness, and IRT analysis further revealed how this response captures with precision a unique portion of anxiety and depression.
+Our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
 

### 2023-01-08 analysis.Rmd
update

@@
 
-df <- read.csv("../IllusionGameReliability/data/preprocessed_questionnaires.csv") |> 
-# df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |> 
-  select(starts_with("Item_PHQ4")) |> 
+df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
+# df <- read.csv("../IllusionGameReliability/data/preprocessed_questionnaires.csv") |> 
+  select(Sex, Age, starts_with("Item_PHQ4")) |> 
   datawizard::remove_empty_rows()
@@
 
+# Participants
+
+```{r message=FALSE, warning=FALSE}
+report::report_participants(df)
+```
+
 # Distribution
@@
 ```{r distribution, message=FALSE, warning=FALSE, fig.width=fig.height}
+df <- select(df, -Sex, -Age)
+
 data <- df |> 

### 2023-01-08 paper/manuscript.Rmd
update

@@
       - "Resources"
+      - "Software"
       - "Investigation"
@@
     institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
-# authornote: |
 
 abstract: |
-  Something something
+  The PHQ-4 is an ultra brief (4 items) screening questionnaire for depression and anxiety.
+We propose to add one aditional response option ("Once or twice", in between "Not at all" and "Several days") to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
+Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide scoring procedure that preserve the compatability with past versions.
 
@@
 
-keywords          : "visual illusions, illusion game, Pyllusion, personality, general factor"
-wordcount         : "4085"
+keywords          : "PHQ-4, depression, anxiety, brief questionnaire, short scale"
+wordcount         : "1011"
 bibliography      : references.bib
@@
 While the scale has been validated and used in the general population [@lowe20104, @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This diagnostic and discriminative goal materializes in the scale's design and the existence of categorical cut-offs.
-However, it might not be best suited to capture subclinical variability.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, but does not necessarily entail a strong sensitivity to milder mood alterations.
 In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and leaves out the possibility of more subtle occurrences.
-While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of variability quantifications.
-The goal of this study is to increase, with minimal changes to the scale, the sensitivity to very mild mood alterations.
+While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
+The goal of this study is increase, with minimal changes to the scale, the sensitivity to mild mood level inflections.
 
@@
 
-The instructions "Over the last 2 weeks, how often have you been bothered by the following problems?" are followed with 4 items (A1 - Feeling nervous, anxious or on edge; A2 - Not being able to stop or control worrying; D1 - Little interest or pleasure in doing things; D2 - Feeling down, depressed, or hopeless).
+The instructions *"Over the last 2 weeks, how often have you been bothered by the following problems?"* are followed with 4 items (A1 - *Feeling nervous, anxious or on edge*; A2 - *Not being able to stop or control worrying*; D1 - *Little interest or pleasure in doing things*; D2 - *Feeling down, depressed, or hopeless*).
 The original answer options are "Not at all" (0), "Several days" (1), "More than half the days" (2), "Nearly every day" (3).
@@
 
-The sample consists of 500 English-speaking participants (**stats**) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
+The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: [18, 73]; Sex: 49.9\% females, 50.1\% males) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
 
@@
 
-The fully reproducible analysis script is accessible at https://github.com/DominiqueMakowski/PHQ4R.
-
 ## Descriptive
 
-The reliability of the anxiety ($Cronbach's~\alpha = 0.892$; RMSEA = 0.032) and depression ($Cronbach's~\alpha = 0.822$; RMSEA = 0.032) subscales was excellent.
-The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option was the most prevalent response for almost all items (on average selected in **30.07\%** of cases).
+The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales is excellent.
+The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in **29.12\%** of cases).
 
@@
 For each of the subscales, we fitted a unidimensional graded response model [GRM, @samejima1997graded].
-For anxiety, the two items captured 88.1\% of the variance of the latent anxiety dimension ($\theta_{anxiety}$).
-The discrimination parameters suggested that the first item was less precise ($\alpha = 3.20$) than the second item ($\alpha = 12.53$) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a larger range of $\theta_{anxiety}$, as can be seen in **Figure 1B**).
-The two depression items captured 80.6\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 15.69$) than the first ($\alpha = 2.19$).
-Importantly to our objective, it seems that the added "Once or twice" option does cover a selective portion of the latent space.
+For anxiety, the two items captured 89.2\% of the variance of the latent anxiety dimension ($\theta_{anxiety}$).
+The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a more exclusive range of $\theta_{anxiety}$, as can be seen in **Figure 1B**).
+The two depression items captured 82.8\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the first ($\alpha = 2.41$).
+However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
+Importantly to our objective, it seems that the added "Once or twice" option does cover a selective and unique portion of the latent space.
 
@@
 
-To illustrate, a person answering "Several days" and "Once or twice" ro the first and second item of the anxiety subscale would have a score of 1.5 (1 + 0.5 following the basic scoring) and a normalized score of `r insight::format_value(mean(c(0.56, 0.45)))` ($(0.56 + 0.45) / 2$).
+To illustrate, a person answering "Several days" and "Once or twice" to the first and second item of the anxiety subscale would have a score of 1.5 (1 + 0.5 following the basic scoring) and a normalized score of `r insight::format_value(mean(c(0.55, 0.44)))` ($(0.55 + 0.44) / 2$).
 
@@
     Basic_Scoring = rep(c("0", "0.5", "1", "2", "3"), 4),
-    Normalized_Scoring = rep(c("0", "0.5", "0.7", "0.8", "1"), 4)
+    Normalized_Scoring = c("0", "0.34", "0.55", "0.71", "1",
+                           "0", "0.44", "0.58", "0.71", "1",
+                           "0", "0.38", "0.51", "0.62", "1",
+                           "0", "0.35", "0.52", "0.66", "1")
     ),
@@
 Our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
+Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by comparing it with other more comprehensive measures.
 

### 2023-01-08 paper/manuscript.tex
update

@@
 {}{\PatchFailed}
-\keywords{visual illusions, illusion game, Pyllusion, personality, general factor\newline\indent Word count: 4085}
+\keywords{PHQ-4, depression, anxiety, brief questionnaire, short scale\newline\indent Word count: 1011}
 \usepackage{lineno}
@@
   pdflang={en-EN},
-  pdfkeywords={visual illusions, illusion game, Pyllusion, personality, general factor},
+  pdfkeywords={PHQ-4, depression, anxiety, brief questionnaire, short scale},
   hidelinks,
@@
 The authors made the following contributions.
-Dominique Makowski: Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Software, Supervision, Validation, Visualization, Writing -- original draft; An Shu Te: Project administration, Resources, Investigation, Writing -- original draft; S.H.
+Dominique Makowski: Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Resources, Software, Supervision, Validation, Visualization, Writing -- original draft; An Shu Te: Project administration, Resources, Software, Investigation, Writing -- original draft; S.H.
 Annabel Chen: Project administration, Supervision, Writing -- review \& editing.
@@
 \abstract{%
-Something something
+The PHQ-4 is an ultra brief (4 items) screening questionnaire for depression and anxiety.
+We propose to add one aditional response option (``Once or twice'', in between ``Not at all'' and ``Several days'') to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
+Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide scoring procedure that preserve the compatability with past versions.
 
@@
 While the scale has been validated and used in the general population Hajek \& König (2020), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This diagnostic and discriminative goal materializes in the scale's design and the existence of categorical cut-offs.
-However, it might not be best suited to capture subclinical variability.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, but does not necessarily entail a strong sensitivity to milder mood alterations.
 In particular, the gap between the two lowest possible answers, ``Not at all'' and ``Several days'', is quite large and leaves out the possibility of more subtle occurrences.
-While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of variability quantifications.
-The goal of this study is to increase, with minimal changes to the scale, the sensitivity to very mild mood alterations.
+While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
+The goal of this study is increase, with minimal changes to the scale, the sensitivity to mild mood level inflections.
 
@@
 
-The instructions ``Over the last 2 weeks, how often have you been bothered by the following problems?'' are followed with 4 items (A1 - Feeling nervous, anxious or on edge; A2 - Not being able to stop or control worrying; D1 - Little interest or pleasure in doing things; D2 - Feeling down, depressed, or hopeless).
+The instructions \emph{``Over the last 2 weeks, how often have you been bothered by the following problems?''} are followed with 4 items (A1 - \emph{Feeling nervous, anxious or on edge}; A2 - \emph{Not being able to stop or control worrying}; D1 - \emph{Little interest or pleasure in doing things}; D2 - \emph{Feeling down, depressed, or hopeless}).
 The original answer options are ``Not at all'' (0), ``Several days'' (1), ``More than half the days'' (2), ``Nearly every day'' (3).
@@
 
-The sample consists of 500 English-speaking participants (\textbf{stats}) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
+The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: {[}18, 73{]}; Sex: 49.9\% females, 50.1\% males) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
 
@@
 
-The fully reproducible analysis script is accessible at \url{https://github.com/DominiqueMakowski/PHQ4R}.
-
 \hypertarget{descriptive}{%
@@
 
-The reliability of the anxiety (\(Cronbach's~\alpha = 0.892\); RMSEA = 0.032) and depression (\(Cronbach's~\alpha = 0.822\); RMSEA = 0.032) subscales was excellent.
-The proportion of response types stratified by item (see \textbf{Figure 1A}) shows that the new ``Once or twice'' option was the most prevalent response for almost all items (on average selected in \textbf{30.07\%} of cases).
+The reliability of the anxiety (\(Cronbach's~\alpha = 0.903\); RMSEA = 0.031) and depression (\(Cronbach's~\alpha = 0.841\); RMSEA = 0.044) subscales is excellent.
+The proportion of response types stratified by item (see \textbf{Figure 1A}) shows that the new ``Once or twice'' option was the most prevalent response for all items (on average selected in \textbf{29.12\%} of cases).
 
@@
 For each of the subscales, we fitted a unidimensional graded response model (GRM, Samejima, 1997).
-For anxiety, the two items captured 88.1\% of the variance of the latent anxiety dimension (\(\theta_{anxiety}\)).
-The discrimination parameters suggested that the first item was less precise (\(\alpha = 3.20\)) than the second item (\(\alpha = 12.53\)) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a larger range of \(\theta_{anxiety}\), as can be seen in \textbf{Figure 1B}).
-The two depression items captured 80.6\% of the variance of its latent trait (\(\theta_{depression}\)), and the opposite pattern was found: the first item had a higher precision (\(\alpha = 15.69\)) than the first (\(\alpha = 2.19\)).
-Importantly to our objective, it seems that the added ``Once or twice'' option does cover a selective portion of the latent space.
+For anxiety, the two items captured 89.2\% of the variance of the latent anxiety dimension (\(\theta_{anxiety}\)).
+The discrimination parameters suggested that the first item was less precise (\(\alpha = 3.42\)) than the second item (\(\alpha = 12.55\)) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a more exclusive range of \(\theta_{anxiety}\), as can be seen in \textbf{Figure 1B}).
+The two depression items captured 82.8\% of the variance of its latent trait (\(\theta_{depression}\)), and the opposite pattern was found: the first item had a higher precision (\(\alpha = 16.46\)) than the first (\(\alpha = 2.41\)).
+However, it is important to note that the ``less precise'' items were also the ones covering a larger portion of the latent space (being more sensitive on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
+Importantly to our objective, it seems that the added ``Once or twice'' option does cover a selective and unique portion of the latent space.
 
@@
 
-To illustrate, a person answering ``Several days'' and ``Once or twice'' ro the first and second item of the anxiety subscale would have a score of 1.5 (1 + 0.5 following the basic scoring) and a normalized score of 0.50 (\((0.56 + 0.45) / 2\)).
+To illustrate, a person answering ``Several days'' and ``Once or twice'' to the first and second item of the anxiety subscale would have a score of 1.5 (1 + 0.5 following the basic scoring) and a normalized score of 0.49 (\((0.55 + 0.44) / 2\)).
 
@@
 \textbf{Anxiety} & Feeling nervous, anxious or on edge & \textit{Not at all} & 0 & 0\\
- &  & \textit{Once or twice} & 0.5 & 0.5\\
- &  & \textit{Several days} & 1 & 0.7\\
- &  & \textit{More than half the days} & 2 & 0.8\\
+ &  & \textit{Once or twice} & 0.5 & 0.34\\
+ &  & \textit{Several days} & 1 & 0.55\\
+ &  & \textit{More than half the days} & 2 & 0.71\\
  &  & \textit{Nearly every day} & 3 & 1\\ \midrule
  & Not being able to stop or control worrying & \textit{Not at all} & 0 & 0\\
- &  & \textit{Once or twice} & 0.5 & 0.5\\
- &  & \textit{Several days} & 1 & 0.7\\
- &  & \textit{More than half the days} & 2 & 0.8\\
+ &  & \textit{Once or twice} & 0.5 & 0.44\\
+ &  & \textit{Several days} & 1 & 0.58\\
+ &  & \textit{More than half the days} & 2 & 0.71\\
  &  & \textit{Nearly every day} & 3 & 1\\ \midrule
 \textbf{Depression} & Little interest or pleasure in doing things & \textit{Not at all} & 0 & 0\\
- &  & \textit{Once or twice} & 0.5 & 0.5\\
- &  & \textit{Several days} & 1 & 0.7\\
- &  & \textit{More than half the days} & 2 & 0.8\\
+ &  & \textit{Once or twice} & 0.5 & 0.38\\
+ &  & \textit{Several days} & 1 & 0.51\\
+ &  & \textit{More than half the days} & 2 & 0.62\\
  &  & \textit{Nearly every day} & 3 & 1\\ \midrule
  & Feeling down, depressed, or hopeless & \textit{Not at all} & 0 & 0\\
- &  & \textit{Once or twice} & 0.5 & 0.5\\
- &  & \textit{Several days} & 1 & 0.7\\
- &  & \textit{More than half the days} & 2 & 0.8\\
+ &  & \textit{Once or twice} & 0.5 & 0.35\\
+ &  & \textit{Several days} & 1 & 0.52\\
+ &  & \textit{More than half the days} & 2 & 0.66\\
  &  & \textit{Nearly every day} & 3 & 1\\

### 2023-01-08 paper/manuscript.Rmd
Update manuscript.Rmd

@@
 
-The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: [18, 73]; Sex: 49.9\% females, 50.1\% males) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
+The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: [18, 73]; Sex: 49.9\% females, 50.1\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
 

### 2023-01-08 paper/manuscript.Rmd

@@
     institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
-
 abstract: |
@@
 The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety [@kroenke2009ultra].
-It consists of two items for depression (PHQ–2) and anxiety (GAD–2), each corresponding to DSM-IV Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
-It has been validated across many languages and samples [@christodoulaki2022validation, @mendoza2022factor, @materu2020psychometric], becoming one of the most popular screening instrument [@maurer2018depression].
+It consists of two items for depression (PHQ–2; @kroenke2003patient) and anxiety (GAD–2; @kroenke2007anxiety), each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
+It has been validated across many languages and different populations [@christodoulaki2022validation, @mendoza2022factor, @materu2020psychometric], becoming one of the most popular screening instruments for depression and anxiety [@maurer2018depression].
 
 While the scale has been validated and used in the general population [@lowe20104, @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, but does not necessarily entail a strong sensitivity to milder mood alterations.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, but in doing so, limits its sensitivity to milder mood alterations.
 In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and leaves out the possibility of more subtle occurrences.
-While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
-The goal of this study is increase, with minimal changes to the scale, the sensitivity to mild mood level inflections.
+While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of psychological distress on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
+The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.
 
@@
 
-This study was approved by the NTU Institutional Review Board (NTU **NUMBER**).
+This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187) and all procedures performed were in accordance with the ethical standards of the institutional board and with the 1964 Helsinki Declaration.
 All participants provided their informed consent prior to participation and were incentivized after completing the study.
@@
 
-The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales is excellent.
-The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in **29.12\%** of cases).
+The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales are excellent (>0.8).
+The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option we added was the most prevalent response for all items (on average selected in **29.12\%** of cases).
 
@@
 
-Item Response Theory (IRT) provides insights into how well items and responses capture the underlying latent trait $\theta$.
+Item Response Theory (IRT) provides insights into how well items and responses capture an underlying latent trait $\theta$.
 For each of the subscales, we fitted a unidimensional graded response model [GRM, @samejima1997graded].
@@
 However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
-Importantly to our objective, it seems that the added "Once or twice" option does cover a selective and unique portion of the latent space.
+Importantly, for our objective, it seems that the added "Once or twice" option does cover a selective and unique portion of the latent space.
 
@@
 minimally disrupting the original scale and making its scores comparable, enabling comparisons across studies and the usage of developed cut-offs and norms.
-We suggest to score the new option, "Once or twice", as 0.5, occupying the middle space between 0 and 1 (preserving this way the total score range of 0 - 8 per dimension).
+We suggest to score the new option, "Once or twice", as 0.5, occupying the middle space between 0 and 1 (preserving in this way the total score range of 0 - 8 per dimension).
 
 The second scoring method we propose takes into account the non-linear spacing between responses, as indicated by the peak of their measurement information.
-To develop it, we found for each dimension the lower and higher value of the latent trait $\theta$ for which the probability lower and higher probability of identification is 95\%.
-We considered these points at 0 and 1, and normalized the location of the peaks of each response in between.
+To develop this measurement, we found for each dimension the lower and higher values of the latent trait $\theta$ for which the probability of identification is 95\%.
+**In other words, the probability that an individual responds in a particular category given a trait level is 95%.
+(not sure if this is what this means?)** We considered these points at 0 and 1, and normalized the location of the peaks of each response in between.
 
@@
 
-The objective of this study was to test the introduction of a "Once or twice" response option to the PHQ-4 to make it more sensitive to milder fluctuations.
-The fact that this new response option was the most prevalent is in itself evidence for its usefulness, and IRT analysis further revealed how this response captures with precision a unique portion of anxiety and depression.
+The objective of this study was to test the introduction of a "Once or twice" response option to the PHQ-4 to enhance its sensitivity to milder mood fluctuations.
+The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness.
+IRT analysis further revealed that this response captures with precision a unique portion of anxiety and depression.
 Our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
-Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by comparing it with other more comprehensive measures.
+Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by cross-validating it with other more comprehensive measures.
 

### 2023-01-08 paper/manuscript.Rmd
minor

@@
 The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety [@kroenke2009ultra].
-It consists of two items for depression (PHQ–2; @kroenke2003patient) and anxiety (GAD–2; @kroenke2007anxiety), each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
-It has been validated across many languages and different populations [@christodoulaki2022validation, @mendoza2022factor, @materu2020psychometric], becoming one of the most popular screening instruments for depression and anxiety [@maurer2018depression].
+It consists of two items for depression [PHQ–2, @kroenke2003patient] and anxiety [GAD–2, @kroenke2007anxiety], each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
+It has been validated across many languages and populations [@christodoulaki2022validation, @mendoza2022factor, @materu2020psychometric], becoming one of the most popular screening instruments for depression and anxiety [@maurer2018depression].
 
 While the scale has been validated and used in the general population [@lowe20104, @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, but in doing so, limits its sensitivity to milder mood alterations.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations.
 In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and leaves out the possibility of more subtle occurrences.
-While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of psychological distress on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
+While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
 The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.
@@
 
-This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187) and all procedures performed were in accordance with the ethical standards of the institutional board and with the 1964 Helsinki Declaration.
+This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187).
 All participants provided their informed consent prior to participation and were incentivized after completing the study.
@@
 
-The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales are excellent (>0.8).
-The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option we added was the most prevalent response for all items (on average selected in **29.12\%** of cases).
+The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales is excellent.
+The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in **29.12\%** of cases).
 
@@
 The second scoring method we propose takes into account the non-linear spacing between responses, as indicated by the peak of their measurement information.
-To develop this measurement, we found for each dimension the lower and higher values of the latent trait $\theta$ for which the probability of identification is 95\%.
-**In other words, the probability that an individual responds in a particular category given a trait level is 95%.
-(not sure if this is what this means?)** We considered these points at 0 and 1, and normalized the location of the peaks of each response in between.
+Specifically, we found for each dimension the lower and higher values of the latent trait $\theta$ for which the probability of identification is 95\%.
+We considered these points as the 0 and 1 ends, and normalized the location of the peaks of each response in between.
 
@@
 IRT analysis further revealed that this response captures with precision a unique portion of anxiety and depression.
-Our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
+Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
 Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by cross-validating it with other more comprehensive measures.

### 2023-01-08 paper/manuscript.tex
minor

@@
 The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety (Kroenke et al., 2009).
-It consists of two items for depression (PHQ--2) and anxiety (GAD--2), each corresponding to DSM-IV Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
-It has been validated across many languages and samples Materu et al.
-(2020), becoming one of the most popular screening instrument (Maurer et al., 2018).
+It consists of two items for depression (PHQ--2, Kroenke et al., 2003) and anxiety (GAD--2, Kroenke et al., 2007), each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
+It has been validated across many languages and populations Materu et al.
+(2020), becoming one of the most popular screening instruments for depression and anxiety (Maurer et al., 2018).
 
 While the scale has been validated and used in the general population Hajek \& König (2020), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, but does not necessarily entail a strong sensitivity to milder mood alterations.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations.
 In particular, the gap between the two lowest possible answers, ``Not at all'' and ``Several days'', is quite large and leaves out the possibility of more subtle occurrences.
-While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
-The goal of this study is increase, with minimal changes to the scale, the sensitivity to mild mood level inflections.
+While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
+The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.
 
@@
 
-The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: {[}18, 73{]}; Sex: 49.9\% females, 50.1\% males) who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
+The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: {[}18, 73{]}; Sex: 49.9\% females, 50.1\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
 
-This study was approved by the NTU Institutional Review Board (NTU \textbf{NUMBER}).
+This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187).
 All participants provided their informed consent prior to participation and were incentivized after completing the study.
@@
 
-Item Response Theory (IRT) provides insights into how well items and responses capture the underlying latent trait \(\theta\).
+Item Response Theory (IRT) provides insights into how well items and responses capture an underlying latent trait \(\theta\).
 For each of the subscales, we fitted a unidimensional graded response model (GRM, Samejima, 1997).
@@
 However, it is important to note that the ``less precise'' items were also the ones covering a larger portion of the latent space (being more sensitive on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
-Importantly to our objective, it seems that the added ``Once or twice'' option does cover a selective and unique portion of the latent space.
+Importantly, for our objective, it seems that the added ``Once or twice'' option does cover a selective and unique portion of the latent space.
 
@@
 minimally disrupting the original scale and making its scores comparable, enabling comparisons across studies and the usage of developed cut-offs and norms.
-We suggest to score the new option, ``Once or twice'', as 0.5, occupying the middle space between 0 and 1 (preserving this way the total score range of 0 - 8 per dimension).
+We suggest to score the new option, ``Once or twice'', as 0.5, occupying the middle space between 0 and 1 (preserving in this way the total score range of 0 - 8 per dimension).
 
 The second scoring method we propose takes into account the non-linear spacing between responses, as indicated by the peak of their measurement information.
-To develop it, we found for each dimension the lower and higher value of the latent trait \(\theta\) for which the probability lower and higher probability of identification is 95\%.
-We considered these points at 0 and 1, and normalized the location of the peaks of each response in between.
+Specifically, we found for each dimension the lower and higher values of the latent trait \(\theta\) for which the probability of identification is 95\%.
+We considered these points as the 0 and 1 ends, and normalized the location of the peaks of each response in between.
 
@@
 
-The objective of this study was to test the introduction of a ``Once or twice'' response option to the PHQ-4 to make it more sensitive to milder fluctuations.
-The fact that this new response option was the most prevalent is in itself evidence for its usefulness, and IRT analysis further revealed how this response captures with precision a unique portion of anxiety and depression.
-Our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
-Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by comparing it with other more comprehensive measures.
+The objective of this study was to test the introduction of a ``Once or twice'' response option to the PHQ-4 to enhance its sensitivity to milder mood fluctuations.
+The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness.
+IRT analysis further revealed that this response captures with precision a unique portion of anxiety and depression.
+Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
+Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by cross-validating it with other more comprehensive measures.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kroenke2003patient}{}}%
+Kroenke, K., Spitzer, R. L., \& Williams, J. B. (2003). The patient health questionnaire-2: Validity of a two-item depression screener. \emph{Medical Care}, 1284--1292.
+
 \leavevmode\vadjust pre{\hypertarget{ref-kroenke2009ultra}{}}%
 Kroenke, K., Spitzer, R. L., Williams, J. B., \& Löwe, B. (2009). An ultra-brief screening scale for anxiety and depression: The PHQ--4. \emph{Psychosomatics}, \emph{50}(6), 613--621.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroenke2007anxiety}{}}%
+Kroenke, K., Spitzer, R.
+L., Williams, J.
+B., Monahan, P.
+O., \& Löwe, B.
+(2007).
+Anxiety disorders in primary care: Prevalence, impairment, comorbidity, and detection.
+\emph{Annals of Internal Medicine}, \emph{146}(5), 317--325.
 

### 2023-01-08 paper/manuscript.Rmd
formatting

@@
 It consists of two items for depression [PHQ–2, @kroenke2003patient] and anxiety [GAD–2, @kroenke2007anxiety], each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
-It has been validated across many languages and populations [@christodoulaki2022validation, @mendoza2022factor, @materu2020psychometric], becoming one of the most popular screening instruments for depression and anxiety [@maurer2018depression].
+It has been validated across many languages and populations [@christodoulaki2022validation; @mendoza2022factor; @materu2020psychometric], becoming one of the most popular screening instruments for depression and anxiety [@maurer2018depression].
 
-While the scale has been validated and used in the general population [@lowe20104, @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
+While the scale has been validated and used in the general population [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
 This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations.

### 2023-01-08 paper/manuscript.tex
formatting

@@
 It consists of two items for depression (PHQ--2, Kroenke et al., 2003) and anxiety (GAD--2, Kroenke et al., 2007), each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
-It has been validated across many languages and populations Materu et al.
-(2020), becoming one of the most popular screening instruments for depression and anxiety (Maurer et al., 2018).
+It has been validated across many languages and populations (Christodoulaki et al., 2022; Materu et al., 2020; Mendoza et al., 2022), becoming one of the most popular screening instruments for depression and anxiety (Maurer et al., 2018).
 
-While the scale has been validated and used in the general population Hajek \& König (2020), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
+While the scale has been validated and used in the general population (Hajek \& König, 2020; Löwe et al., 2010), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
 This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations.

### 2023-01-08 paper/manuscript.Rmd
fix

@@
 The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales is excellent.
-The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in **29.12\%** of cases).
+The proportion of response types stratified by item (see **Figure 1A**) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in 29.12\% of cases).
 
@@
 For anxiety, the two items captured 89.2\% of the variance of the latent anxiety dimension ($\theta_{anxiety}$).
-The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a more exclusive range of $\theta_{anxiety}$, as can be seen in **Figure 1B**).
+The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of $\theta_{anxiety}$, as can be seen in **Figure 1B**).
 The two depression items captured 82.8\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the first ($\alpha = 2.41$).
-However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
-Importantly, for our objective, it seems that the added "Once or twice" option does cover a selective and unique portion of the latent space.
+However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive especially on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
+Importantly for our objective, the added "Once or twice" option does cover a selective and unique portion of the latent space.
 

### 2023-01-08 paper/manuscript.tex
fix

@@
 The reliability of the anxiety (\(Cronbach's~\alpha = 0.903\); RMSEA = 0.031) and depression (\(Cronbach's~\alpha = 0.841\); RMSEA = 0.044) subscales is excellent.
-The proportion of response types stratified by item (see \textbf{Figure 1A}) shows that the new ``Once or twice'' option was the most prevalent response for all items (on average selected in \textbf{29.12\%} of cases).
+The proportion of response types stratified by item (see \textbf{Figure 1A}) shows that the new ``Once or twice'' option was the most prevalent response for all items (on average selected in 29.12\% of cases).
 
@@
 For anxiety, the two items captured 89.2\% of the variance of the latent anxiety dimension (\(\theta_{anxiety}\)).
-The discrimination parameters suggested that the first item was less precise (\(\alpha = 3.42\)) than the second item (\(\alpha = 12.55\)) in its ability to discriminate between various levels of anxiety (i.e., each response on item 1 covers a more exclusive range of \(\theta_{anxiety}\), as can be seen in \textbf{Figure 1B}).
+The discrimination parameters suggested that the first item was less precise (\(\alpha = 3.42\)) than the second item (\(\alpha = 12.55\)) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of \(\theta_{anxiety}\), as can be seen in \textbf{Figure 1B}).
 The two depression items captured 82.8\% of the variance of its latent trait (\(\theta_{depression}\)), and the opposite pattern was found: the first item had a higher precision (\(\alpha = 16.46\)) than the first (\(\alpha = 2.41\)).
-However, it is important to note that the ``less precise'' items were also the ones covering a larger portion of the latent space (being more sensitive on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
-Importantly, for our objective, it seems that the added ``Once or twice'' option does cover a selective and unique portion of the latent space.
+However, it is important to note that the ``less precise'' items were also the ones covering a larger portion of the latent space (being more sensitive especially on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
+Importantly for our objective, the added ``Once or twice'' option does cover a selective and unique portion of the latent space.
 

### 2023-01-09 analysis.Rmd
scoring function

@@
   select(Sex, Age, starts_with("Item_PHQ4")) |> 
-  datawizard::remove_empty_rows()
+  filter(!is.na(Item_PHQ4_Anxiety_1))
 
@@
 
+# Scores
+
+```{r message=FALSE, warning=FALSE}
+data <- sapply(cbind(df_anx, df_dep), function(x) {
+  case_when(
+      x == 0 ~ "Not at all",
+      x == 1 ~ "Once or twice",
+      x == 2 ~ "Several days",
+      x == 3 ~ "More than half the days",
+      TRUE ~ "Nearly every day"
+    )
+}) |> 
+  as.data.frame()
+
+source("score_PHQ4.R")
+
+data <- score_PHQ4(A1=data$Anxiety_1, A2=data$Anxiety_2, D1=data$Depression_3, D2=data$Depression_4, method="basic") |> 
+  datawizard::data_addsuffix("_Basic") |> 
+  cbind(
+    score_PHQ4(A1=data$Anxiety_1, A2=data$Anxiety_2, D1=data$Depression_3, D2=data$Depression_4, method="normalized") |> 
+      datawizard::data_addsuffix("_Normalized")) |> 
+  select(matches("Depression|Anxiety")) |> 
+  pivot_longer(everything()) |> 
+  separate(name, into = c("Dimension", "Scoring"))
+
+data |> 
+  ggplot(aes(x=value)) + 
+  geom_area(data=estimate_density(data, at = c("Dimension", "Scoring"), method = "KernSmooth") |> 
+              group_by(Scoring) |> 
+              normalize(select = "y"),
+            aes(x = x, y = y, fill=Dimension), alpha = 0.05, position="identity") +
+  geom_hline(yintercept = 0.5, linetype = "dotted") +
+  stat_ecdf(aes(color=Dimension), geom = "step", size=1) +
+  facet_wrap(~Scoring, scales = "free") +
+  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
+  scale_color_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63")) +
+  theme_modern(axis.title.space = 5) +
+  labs(title = "Distribution and Cumulative Density of the Sample", x = "Score", y = "Cumulative Proportion")
+```
+
 # Correlation

### 2023-01-09 paper/manuscript.Rmd
scoring function

@@
 We propose to add one aditional response option ("Once or twice", in between "Not at all" and "Several days") to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
-Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide scoring procedure that preserve the compatability with past versions.
+Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide a scoring procedure that preserve the compatability with past versions.
 
   **Significance Statement**.
-Developing reliable and sensitive instruments for mood disorders screening is critical in a global context marked by international crises (pandemics, wars), where more and more surveys as done online.
+Developing reliable and sensitive instruments for mood disorders screening is critical in a global context marked by international crises (pandemics, wars), where increasingly more surveys are done online.
 In this study, we show that a small modification to the widely used PHQ-4 scale (adding the "Once or twice" response option) can increase its ability to capture the mild fluctuations prevalent in subclinical samples.
@@
 The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety [@kroenke2009ultra].
-It consists of two items for depression [PHQ–2, @kroenke2003patient] and anxiety [GAD–2, @kroenke2007anxiety], each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
+It consists of two items for depression [PHQ–2, @kroenke2003patient] and anxiety [GAD–2, @kroenke2007anxiety], each corresponding to DSM-5 diagnostic symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
 It has been validated across many languages and populations [@christodoulaki2022validation; @mendoza2022factor; @materu2020psychometric], becoming one of the most popular screening instruments for depression and anxiety [@maurer2018depression].
@@
 
-## Original Scale
+## Original Scale and Revision
 
@@
 
-## Revision
-
 In order to better capture potential mild mood inflections without altering the scale scoring or structure, we added a "Once or twice" option between "Not at all" and "Several days" [see @dobson1979equidistant for the choice of the label].
@@
 
-The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: [18, 73]; Sex: 49.9\% females, 50.1\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
-
+The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
 This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187).
@@
 
-## Descriptive
+## Descriptive Statistics
 
@@
 Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
-Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by cross-validating it with other more comprehensive measures.
+
+Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations), cross-validate it with other more comprehensive measures, and refine the normalized scoring estimates in larger sample.
 
@@
 
-The dataset analysed during the current study are available in the GitHub repository https://github.com/DominiqueMakowski/PHQ4R
+The dataset analysed during the current study are available in the GitHub repository https://github.com/DominiqueMakowski/PHQ4R.
+We have created an R function for easy scoring according to the two methods described, which can be used as follows:
+
+```{r message=FALSE, warning=FALSE}
+# Load the function
+source()
+```
 

### 2023-01-09 score_PHQ4.R

@@ -1,0 +1,49 @@
+score_PHQ4 <- function(A1=0, A2=0, D1=0, D2=0, method="basic") {
+  df <- data.frame(A1 = A1, A2 = A2, D1 = D1, D2 = D2)
+  df <- as.data.frame(sapply(df, tolower))
+  if (method == "basic") {
+    df <- data.frame(sapply(df, function(x) {
+      case_when(
+        x == "not at all" ~ 0,
+        x == "once or twice" ~ 0.5,
+        x == "several days" ~ 1,
+        x == "more than half the days" ~ 2,
+        TRUE ~ 3
+      )
+    }))
+    df$Anxiety <- df$A1 + df$A2
+    df$Depression <- df$D1 + df$D2
+  } else {
+    df$A1 <- case_when(
+      df$A1 == "not at all" ~ 0,
+      df$A1 == "once or twice" ~ 0.34,
+      df$A1 == "several days" ~ 0.55,
+      df$A1 == "more than half the days" ~ 0.71,
+      TRUE ~ 1
+    )
+    df$A2 <- case_when(
+      df$A2 == "not at all" ~ 0,
+      df$A2 == "once or twice" ~ 0.44,
+      df$A2 == "several days" ~ 0.58,
+      df$A2 == "more than half the days" ~ 0.71,
+      TRUE ~ 1
+    )
+    df$D1 <- case_when(
+      df$D1 == "not at all" ~ 0,
+      df$D1 == "once or twice" ~ 0.38,
+      df$D1 == "several days" ~ 0.51,
+      df$D1 == "more than half the days" ~ 0.62,
+      TRUE ~ 1
+    )
+    df$D2 <- case_when(
+      df$D2 == "not at all" ~ 0,
+      df$D2 == "once or twice" ~ 0.35,
+      df$D2 == "several days" ~ 0.52,
+      df$D2 == "more than half the days" ~ 0.66,
+      TRUE ~ 1
+    )
+    df$Anxiety <- (df$A1 + df$A2) / 2
+    df$Depression <- (df$D1 + df$D2) / 2
+  }
+  df
+}

### 2023-01-09 paper/manuscript.Rmd
update

@@
 
-Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations), cross-validate it with other more comprehensive measures, and refine the normalized scoring estimates in larger sample.
+The added value in terms of finer sensitivity that we hypothesize should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression.
+Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new "Once or twice" option pulls participants that would have otherwise answer "Not at all" or "Several days".
+Finally, future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) to see if it affects the cut-offs proposed in the literature (and whether it allows for a better discrimination of mood disorders), and refine if need be the normalized scoring estimates in larger sample.
 
@@
 # Load the function
-source()
+source("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/score_PHQ4.R")
+
+score_PHQ4(A1 = "Not at all", A2 = "Once or twice",
+           D1 = "Once or twice", D2 = "several days",
+           method = "basic")
 ```

### 2023-01-09 score_PHQ4.R

@@ -1,41 +1,57 @@
-score_PHQ4 <- function(A1=0, A2=0, D1=0, D2=0, method="basic") {
+score_PHQ4 <- function(A1, A2, D1, D2, method="basic") {
   df <- data.frame(A1 = A1, A2 = A2, D1 = D1, D2 = D2)
-  df <- as.data.frame(sapply(df, tolower))
+  df <- as.data.frame(sapply(df, tolower, simplify=FALSE))
+
+  # Sanity check
+  for(i in names(df)) {
+    if(!all(df[[i]] %in% c("not at all",
+                    "once or twice",
+                    "several days",
+                    "more than half the days",
+                    "nearly every day"))){
+      stop(paste0("Non-allowed answers in ",
+                  i,
+                  ". Answers should be 'not at all', 'once or twice'",
+                  ", 'several days', 'more than half the days'",
+                  ", or 'nearly every day'."))
+    }
+  }
+
   if (method == "basic") {
     df <- data.frame(sapply(df, function(x) {
-      case_when(
+      dplyr::case_when(
         x == "not at all" ~ 0,
         x == "once or twice" ~ 0.5,
         x == "several days" ~ 1,
         x == "more than half the days" ~ 2,
         TRUE ~ 3
       )
-    }))
+    }, simplify=FALSE))
     df$Anxiety <- df$A1 + df$A2
     df$Depression <- df$D1 + df$D2
   } else {
-    df$A1 <- case_when(
+    df$A1 <- dplyr::case_when(
       df$A1 == "not at all" ~ 0,
       df$A1 == "once or twice" ~ 0.34,
       df$A1 == "several days" ~ 0.55,
       df$A1 == "more than half the days" ~ 0.71,
       TRUE ~ 1
     )
-    df$A2 <- case_when(
+    df$A2 <- dplyr::case_when(
       df$A2 == "not at all" ~ 0,
       df$A2 == "once or twice" ~ 0.44,
       df$A2 == "several days" ~ 0.58,
       df$A2 == "more than half the days" ~ 0.71,
       TRUE ~ 1
     )
-    df$D1 <- case_when(
+    df$D1 <- dplyr::case_when(
       df$D1 == "not at all" ~ 0,
       df$D1 == "once or twice" ~ 0.38,
       df$D1 == "several days" ~ 0.51,
       df$D1 == "more than half the days" ~ 0.62,
       TRUE ~ 1
     )
[…]

### 2023-01-09 paper/manuscript.Rmd
update

@@
 keywords          : "PHQ-4, depression, anxiety, brief questionnaire, short scale"
-wordcount         : "1011"
+wordcount         : "1158"
 bibliography      : references.bib
@@
 
-The added value in terms of finer sensitivity that we hypothesize should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression.
+The hypothesized added value in terms of finer sensitivity should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression.
 Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new "Once or twice" option pulls participants that would have otherwise answer "Not at all" or "Several days".
 Finally, future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) to see if it affects the cut-offs proposed in the literature (and whether it allows for a better discrimination of mood disorders), and refine if need be the normalized scoring estimates in larger sample.
+
+\newpage
 
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r echo=TRUE, message=FALSE, warning=FALSE}
 # Load the function
@@
 
+# Enter the responses
 score_PHQ4(A1 = "Not at all", A2 = "Once or twice",

### 2023-01-09 paper/manuscript.tex
update

@@
 \usepackage{xcolor}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx}
@@
 {}{\PatchFailed}
-\keywords{PHQ-4, depression, anxiety, brief questionnaire, short scale\newline\indent Word count: 1011}
+\keywords{PHQ-4, depression, anxiety, brief questionnaire, short scale\newline\indent Word count: 1158}
 \usepackage{lineno}
@@
 We propose to add one aditional response option (``Once or twice'', in between ``Not at all'' and ``Several days'') to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
-Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide scoring procedure that preserve the compatability with past versions.
+Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide a scoring procedure that preserve the compatability with past versions.
 
 \textbf{Significance Statement}.
-Developing reliable and sensitive instruments for mood disorders screening is critical in a global context marked by international crises (pandemics, wars), where more and more surveys as done online.
+Developing reliable and sensitive instruments for mood disorders screening is critical in a global context marked by international crises (pandemics, wars), where increasingly more surveys are done online.
 In this study, we show that a small modification to the widely used PHQ-4 scale (adding the ``Once or twice'' response option) can increase its ability to capture the mild fluctuations prevalent in subclinical samples.
@@
 The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety (Kroenke et al., 2009).
-It consists of two items for depression (PHQ--2, Kroenke et al., 2003) and anxiety (GAD--2, Kroenke et al., 2007), each corresponding to DSM-V Diagnostic Criterion A symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
+It consists of two items for depression (PHQ--2, Kroenke et al., 2003) and anxiety (GAD--2, Kroenke et al., 2007), each corresponding to DSM-5 diagnostic symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD).
 It has been validated across many languages and populations (Christodoulaki et al., 2022; Materu et al., 2020; Mendoza et al., 2022), becoming one of the most popular screening instruments for depression and anxiety (Maurer et al., 2018).
@@
 
-\hypertarget{original-scale}{%
-\subsection{Original Scale}\label{original-scale}}
+\hypertarget{original-scale-and-revision}{%
+\subsection{Original Scale and Revision}\label{original-scale-and-revision}}
 
@@
 The total score is computed by summing the responses of each facet.
-
-\hypertarget{revision}{%
-\subsection{Revision}\label{revision}}
 
@@
 
-The sample consists of 497 English-speaking participants (Mean age = 30.0, SD = 10.1, range: {[}18, 73{]}; Sex: 49.9\% females, 50.1\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
-
+The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
 This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187).
@@
 
-\hypertarget{descriptive}{%
-\subsection{Descriptive}\label{descriptive}}
+\hypertarget{descriptive-statistics}{%
+\subsection{Descriptive Statistics}\label{descriptive-statistics}}
 
@@
 Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
-Future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) and by cross-validating it with other more comprehensive measures.
+
+The hypothesized added value in terms of finer sensitivity should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression.
+Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new ``Once or twice'' option pulls participants that would have otherwise answer ``Not at all'' or ``Several days''.
+Finally, future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) to see if it affects the cut-offs proposed in the literature (and whether it allows for a better discrimination of mood disorders), and refine if need be the normalized scoring estimates in larger sample.
+
+\newpage
 
@@
 
-The dataset analysed during the current study are available in the GitHub repository \url{https://github.com/DominiqueMakowski/PHQ4R}
+The dataset analysed during the current study are available in the GitHub repository \url{https://github.com/DominiqueMakowski/PHQ4R}.
+We have created an R function for easy scoring according to the two methods described, which can be used as follows:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Load the function}
+\FunctionTok{source}\NormalTok{(}\StringTok{"https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/score\_PHQ4.R"}\NormalTok{)}
+
+\CommentTok{\# Enter the responses}
+\FunctionTok{score\_PHQ4}\NormalTok{(}\AttributeTok{A1 =} \StringTok{"Not at all"}\NormalTok{, }\AttributeTok{A2 =} \StringTok{"Once or twice"}\NormalTok{,}
+           \AttributeTok{D1 =} \StringTok{"Once or twice"}\NormalTok{, }\AttributeTok{D2 =} \StringTok{"several days"}\NormalTok{,}
+           \AttributeTok{method =} \StringTok{"basic"}\NormalTok{)}
+\DocumentationTok{\#\#   A1  A2  D1 D2 Anxiety Depression}
+\DocumentationTok{\#\# 1  0 0.5 0.5  1     0.5        1.5}
+\end{Highlighting}
+\end{Shaded}
 

### 2023-01-17 analysis.Rmd
update JAD

@@
 fig.height <- 7
+set.seed(333)
 
@@
 
+```{r message=FALSE, warning=FALSE}
+add_labels <- function(x) {
+  x <- case_when(
+      x == 0 ~ "Not at all",
+      x == 1 ~ "Once or twice",
+      x == 2 ~ "Several days",
+      x == 3 ~ "More than half the days",
+      TRUE ~ "Nearly every day"
+    )
+  fct_relevel(x, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day"))
+}
+```
+
 ```{r distribution, message=FALSE, warning=FALSE, fig.width=fig.height}
@@
     # Item = str_split_fixed(Item, "_", 2)[,2],
-    Answer = case_when(
-      Answer == 0 ~ "Not at all",
-      Answer == 1 ~ "Once or twice",
-      Answer == 2 ~ "Several days",
-      Answer == 3 ~ "More than half the days",
-      TRUE ~ "Nearly every day"
-    ),
-    Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")),
+    Answer = add_labels(Answer),
     Item = case_when(
@@
   scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
-  labs(y = "Proportion of Answers", title = "A. Proportion of answers per item") +
+  labs(y = "Proportion of Answers") +
   theme_modern(axis.title.space = 10) +
@@
         plot.title = element_text(face = "bold"),
+        strip.text = element_text(size=rel(0.7)),
         strip.background=element_rect(fill="#EEEEEE", colour="white")) +
   facet_wrap(~Item)
-p1
+p1 + ggtitle("Proportion of answers per item")
 ```
@@
 The "Once or twice" answer was selected in `r insight::format_percent(mean(filter(data, Answer == "Once or twice")$n))`
+
+```{r joint_prevalence, message=FALSE, warning=FALSE, fig.width=fig.width}
+p1b <- rbind(
+  data.frame(Scores = paste0(df$Anxiety_1, "_", df$Anxiety_2),
+             Facet = "Anxiety"),
+  data.frame(Scores = paste0(df$Depression_3, "_", df$Depression_4),
+             Facet = "Depression")
+) |> 
+  group_by(Facet, Scores) |> 
+  summarize(n = n() / nrow(df)) |> 
+  separate(Scores, into = c("Q1", "Q2")) |> 
+  mutate(Label = ifelse(Q1 == Q2, format_percent(n), "")) |> 
+  mutate(Q1 = add_labels(as.numeric(Q1)),
+         Q2 = add_labels(as.numeric(Q2))) |> 
+  ggplot(aes(x = Q1, y = Q2)) +
+  geom_tile(aes(fill = n)) +
+  geom_text(aes(label = Label), size=2, color="white") +
+  scale_y_discrete(expand = c(0, 0)) +
+  scale_x_discrete(expand = c(0, 0)) +
+  scale_fill_gradientn(colors=c("white", "#2196F3", "#3F51B5", "#673AB7"), labels = scales::percent) +
+  labs(fill = "% Participants", x = "Item 1", y = "Item 2") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust=1),
+        panel.grid.major = element_blank(),
+        plot.title = element_text(face = "bold", hjust=0),
+        strip.background=element_rect(fill="#EEEEEE", colour="white"),
+        strip.text = element_text(face = "bold")) +
+  facet_wrap(~Facet) 
+p1b + ggtitle("Joint prevalence of responses")
+```
 
@@
 
+# Figures
+
 ```{r message=FALSE, warning=FALSE}
-fig1 <- wrap_elements(p1) + wrap_elements(p2a / p2b + plot_annotation(title = "B.
-Information Curves per Response Type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))
+fig1 <- wrap_elements(wrap_elements(p1 + plot_annotation(title = "A. Proportion of answers per item", theme = list(plot.title = element_text(face = "bold")))) / 
+                        wrap_elements(p1b + plot_annotation(title = "B. Joint prevalence of responses", theme = list(plot.title = element_text(face = "bold"))))) + 
+  wrap_elements(p2a / p2b + plot_annotation(title = "C.
+Information curves per response type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))
 

### 2023-01-17 paper/1_PA/manuscript.Rmd
update JAD

--- 2023-01-09 paper/manuscript.Rmd
@@
 
-keywords          : "PHQ-4, depression, anxiety, brief questionnaire, short scale"
+keywords          : "PHQ-4; depression; anxiety; brief questionnaire; short scale"
 wordcount         : "1158"
@@
 draft             : no
-mask              : no
+mask              : yes
 figurelist        : yes
@@
 output            : papaja::apa6_pdf
-csl: utils/apa.csl
+csl: "../utils/apa.csl"
 header-includes:
@@
   fig.height = 9,
-  fig.path = "../figures/"
+  fig.path = "../../figures/"
 )
@@
 The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
-This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187).
+This study was approved by the
+Institutional Review Board **[Masked for Review]**.
+<!-- NTU Institutional Review Board (NTU IRB-2022-187).  -->
 All participants provided their informed consent prior to participation and were incentivized after completing the study.
@@
 Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).", out.width="100%"}
-knitr::include_graphics("../figures/figure1.png")
+knitr::include_graphics("../../figures/figure1.png")
 ```
@@
 ```{r message=FALSE, warning=FALSE, fig.cap="Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).", out.width="100%"}
-knitr::include_graphics("../figures/figure2.png")
+knitr::include_graphics("../../figures/figure2.png")
 ```
@@
 
-The dataset analysed during the current study are available in the GitHub repository https://github.com/DominiqueMakowski/PHQ4R.
+The dataset analysed during the current study are available in the GitHub repository **[Masked for Review].** 
+<!-- https://github.com/DominiqueMakowski/PHQ4R. -->
 We have created an R function for easy scoring according to the two methods described, which can be used as follows:
@@
 
-This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
+This work was supported by the **[Masked for Review]**.
+<!-- Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM). -->
 

### 2023-01-17 paper/1_PA/manuscript.tex
update JAD

--- 2023-01-09 paper/manuscript.tex
@@
 \documentclass[
-  man,floatsintext]{apa6}
+  man,mask,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
   pdftitle={Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability},
-  pdfauthor={Dominique Makowski1, An Shu Te1, \& S.H. Annabel Chen1, 2, 3, 4},
   pdflang={en-EN},
@@
 The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
-This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187).
+This study was approved by the
+Institutional Review Board \textbf{{[}Masked for Review{]}}.
 All participants provided their informed consent prior to participation and were incentivized after completing the study.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/figure1} \caption{A.
+\includegraphics[width=1\linewidth]{../../figures/figure1} \caption{A.
 Proportion of answers of each type.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/figure2} \caption{Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).}\label{fig:unnamed-chunk-3}
+\includegraphics[width=1\linewidth]{../../figures/figure2} \caption{Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 
-The dataset analysed during the current study are available in the GitHub repository \url{https://github.com/DominiqueMakowski/PHQ4R}.
+The dataset analysed during the current study are available in the GitHub repository \textbf{{[}Masked for Review{]}.}
 We have created an R function for easy scoring according to the two methods described, which can be used as follows:
@@
 
-This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
+This work was supported by the \textbf{{[}Masked for Review{]}}.
 

### 2023-01-17 paper/2_JAD/manuscript.Rmd
update JAD

--- 2023-01-17 paper/1_PA/manuscript.Rmd
@@
 abstract: |
-  The PHQ-4 is an ultra brief (4 items) screening questionnaire for depression and anxiety.
-We propose to add one aditional response option ("Once or twice", in between "Not at all" and "Several days") to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
-Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide a scoring procedure that preserve the compatability with past versions.
+  The PHQ-4 is an ultra-brief (4 items) screening questionnaire for depression and anxiety.
+We propose to add one additional response option ("Once or twice", in between "Not at all" and "Several days") to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
+Using Item Response Theory (IRT), we provide evidence that this new option does indeed capture specific portions of the measured constructs in the general population.
 
@@
 keywords          : "PHQ-4; depression; anxiety; brief questionnaire; short scale"
-wordcount         : "1158"
+wordcount         : "1249"
 bibliography      : references.bib
@@
 draft             : no
-mask              : yes
+mask              : no
 figurelist        : yes
@@
 
-While the scale has been validated and used in the general population [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
+While the scale has been validated and used in the general population and non-clinical samples [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
 This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations.
@@
 
-The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
+The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population recruited via *Prolific*, a crowd-sourcing platform recognized for providing high quality data [@peer2022data].
+The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
+Note that the participants were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.
 This study was approved by the
-Institutional Review Board **[Masked for Review]**.
-<!-- NTU Institutional Review Board (NTU IRB-2022-187).  -->
+<!-- Institutional Review Board **[Masked for Review]**. -->
+NTU Institutional Review Board (NTU IRB-2022-187).
 All participants provided their informed consent prior to participation and were incentivized after completing the study.
@@
 
-```{r message=FALSE, warning=FALSE, fig.cap="A.
-Proportion of answers of each type.
-B.
-Item Information Curves from IRT showing the coverage by each item and response of the latent dimension.
+```{r message=FALSE, warning=FALSE, fig.cap="A) Proportion of answers of each type.
+B) Prevalence of answer pairs.
+C) Item Information Curves from IRT showing the coverage by each item and response of the latent dimension.
 Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).", out.width="100%"}
@@
 We propose two types of scoring procedures. The first aims at
-<!-- In order to  -->
 minimally disrupting the original scale and making its scores comparable, enabling comparisons across studies and the usage of developed cut-offs and norms.
@@
 The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness.
-IRT analysis further revealed that this response captures with precision a unique portion of anxiety and depression.
+IRT analysis further revealed that this response captures with precision a unique portion of the variability in the latent factors measured by the instrument.
 Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
 
-The hypothesized added value in terms of finer sensitivity should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression.
+It is important to emphasize that our study is a statistical validation in the general population, which analysed how different responses relate to one another.
+It is a first step providing supporting evidence for the proposed change.
+However, the structural gains that we observed in terms off finer sensitivity should be confirmed 1) by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression, and 2) testing the usefulness of this revision on more targeted samples (e.g., clinically characterized populations) to investigate the impact on existing cut-offs and on the detection of mood disorders.
 Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new "Once or twice" option pulls participants that would have otherwise answer "Not at all" or "Several days".
-Finally, future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) to see if it affects the cut-offs proposed in the literature (and whether it allows for a better discrimination of mood disorders), and refine if need be the normalized scoring estimates in larger sample.
+<!-- , and refine if need be the normalized scoring estimates in larger sample. -->
 
@@
 
-The dataset analysed during the current study are available in the GitHub repository **[Masked for Review].** 
-<!-- https://github.com/DominiqueMakowski/PHQ4R. -->
+The dataset analysed during the current study are available in the GitHub repository 
+<!-- **[Masked for Review].**  -->
+https://github.com/DominiqueMakowski/PHQ4R.
 We have created an R function for easy scoring according to the two methods described, which can be used as follows:
@@
 
-This work was supported by the **[Masked for Review]**.
-<!-- Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM). -->
+This work was supported by the 
+<!-- **[Masked for Review]**. -->
+Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
 

### 2023-01-17 paper/2_JAD/manuscript.tex
update JAD

--- 2023-01-09 paper/manuscript.tex
@@
 {}{\PatchFailed}
-\keywords{PHQ-4, depression, anxiety, brief questionnaire, short scale\newline\indent Word count: 1158}
+\keywords{PHQ-4; depression; anxiety; brief questionnaire; short scale\newline\indent Word count: 1249}
 \usepackage{lineno}
@@
   pdflang={en-EN},
-  pdfkeywords={PHQ-4, depression, anxiety, brief questionnaire, short scale},
+  pdfkeywords={PHQ-4; depression; anxiety; brief questionnaire; short scale},
   hidelinks,
@@
 \abstract{%
-The PHQ-4 is an ultra brief (4 items) screening questionnaire for depression and anxiety.
-We propose to add one aditional response option (``Once or twice'', in between ``Not at all'' and ``Several days'') to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
-Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide a scoring procedure that preserve the compatability with past versions.
+The PHQ-4 is an ultra-brief (4 items) screening questionnaire for depression and anxiety.
+We propose to add one additional response option (``Once or twice'', in between ``Not at all'' and ``Several days'') to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
+Using Item Response Theory (IRT), we provide evidence that this new option does indeed capture specific portions of the measured constructs in the general population.
 
@@
 
-While the scale has been validated and used in the general population (Hajek \& König, 2020; Löwe et al., 2010), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
+While the scale has been validated and used in the general population and non-clinical samples (Hajek \& König, 2020; Löwe et al., 2010), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
 This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations.
@@
 
-The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
-This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187).
+The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population recruited via \emph{Prolific}, a crowd-sourcing platform recognized for providing high quality data (Peer et al., 2022).
+The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
+Note that the participants were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}.
+This study was approved by the
+NTU Institutional Review Board (NTU IRB-2022-187).
 All participants provided their informed consent prior to participation and were incentivized after completing the study.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../figures/figure1} \caption{A.
-Proportion of answers of each type.
-B.
-Item Information Curves from IRT showing the coverage by each item and response of the latent dimension.
+\includegraphics[width=1\linewidth]{../../figures/figure1} \caption{A) Proportion of answers of each type.
+B) Prevalence of answer pairs.
+C) Item Information Curves from IRT showing the coverage by each item and response of the latent dimension.
 Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).}\label{fig:unnamed-chunk-2}
@@
 The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness.
-IRT analysis further revealed that this response captures with precision a unique portion of anxiety and depression.
+IRT analysis further revealed that this response captures with precision a unique portion of the variability in the latent factors measured by the instrument.
 Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
 
-The hypothesized added value in terms of finer sensitivity should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression.
+It is important to emphasize that our study is a statistical validation in the general population, which analysed how different responses relate to one another.
+It is a first step providing supporting evidence for the proposed change.
+However, the structural gains that we observed in terms off finer sensitivity should be confirmed 1) by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression, and 2) testing the usefulness of this revision on more targeted samples (e.g., clinically characterized populations) to investigate the impact on existing cut-offs and on the detection of mood disorders.
 Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new ``Once or twice'' option pulls participants that would have otherwise answer ``Not at all'' or ``Several days''.
-Finally, future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) to see if it affects the cut-offs proposed in the literature (and whether it allows for a better discrimination of mood disorders), and refine if need be the normalized scoring estimates in larger sample.
 
@@
 
-The dataset analysed during the current study are available in the GitHub repository \url{https://github.com/DominiqueMakowski/PHQ4R}.
+The dataset analysed during the current study are available in the GitHub repository
+\url{https://github.com/DominiqueMakowski/PHQ4R}.
 We have created an R function for easy scoring according to the two methods described, which can be used as follows:
@@
 
-This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
+This work was supported by the
+Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-peer2022data}{}}%
+Peer, E., Rothschild, D., Gordon, A., Evernden, Z., \& Damer, E.
+(2022).
+Data quality of platforms and panels for online behavioral research.
+\emph{Behavior Research Methods}, \emph{54}(4), 1643--1662.
+
 \leavevmode\vadjust pre{\hypertarget{ref-samejima1997graded}{}}%

### 2023-10-31 paper/3_Assessment/manuscript.Rmd
prepare study 2

--- 2023-01-17 paper/2_JAD/manuscript.Rmd
@@
   - name          : "An Shu Te"
-    affiliation   : "1"
+    affiliation   : "2"
     orcid         : 0000-0001-8008-2824
@@
   - name          : "S.H. Annabel Chen"
-    affiliation   : "1, 2, 3, 4"
+    affiliation   : "2, 3, 4, 5"
     orcid         : 0000-0002-1540-5516
@@
 affiliation:
-
   - id            : "1"
+    institution   : "School of Psychology, University of Sussex, UK"
+  - id            : "2"
     institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
-  - id            : "2"
+  - id            : "3"
     institution   : "LKC Medicine, Nanyang Technological University, Singapore"
-  - id            : "3"
+  - id            : "4"
     institution   : "National Institute of Education, Singapore"
-  - id            : "4"
+  - id            : "5"
     institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
@@
   The PHQ-4 is an ultra-brief (4 items) screening questionnaire for depression and anxiety.
-We propose to add one additional response option ("Once or twice", in between "Not at all" and "Several days") to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
-Using Item Response Theory (IRT), we provide evidence that this new option does indeed capture specific portions of the measured constructs in the general population.
+In this brief report, we test the benefits of adding one additional response option ("Once or twice", in between "Not at all" and "Several days") to improve the scale's sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
+Using Item Response Theory (IRT), we provide evidence that this new option does indeed extends the sensitivity to the measured constructs on the lower end of the spectrum.
 
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_docx
 csl: "../utils/apa.csl"
@@
 
-While the scale has been validated and used in the general population and non-clinical samples [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations.
-In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and leaves out the possibility of more subtle occurrences.
-While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications.
-The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.
+While the scale has been validated and used in the general population and non-clinical samples [@lowe20104; @hajek2020prevalence], it has to be noted that the initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a focus on the sensitivity to milder mood alterations.
+In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and possibly leaves out the possibility of more subtle occurrences.
+While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantification.
+Thus, the goal of this brief report is to test the possibility of enhancing, with minimal changes to the original scale, the PHQ-4 sensitivity to mild mood level inflections.
 

### 2023-10-31 paper/3_Assessment/manuscript.tex
prepare study 2

--- 2023-01-17 paper/2_JAD/manuscript.tex
@@
   pdftitle={Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability},
-  pdfauthor={Dominique Makowski1, An Shu Te1, \& S.H. Annabel Chen1, 2, 3, 4},
+  pdfauthor={Dominique Makowski1, An Shu Te2, \& S.H. Annabel Chen2, 3, 4, 5},
   pdflang={en-EN},
@@
 \title{\textbf{Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability}}
-\author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{1}, \& S.H. Annabel Chen\textsuperscript{1, 2, 3, 4}}
+\author{Dominique Makowski\textsuperscript{1}, An Shu Te\textsuperscript{2}, \& S.H. Annabel Chen\textsuperscript{2, 3, 4, 5}}
 \date{}
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Social Sciences, Nanyang Technological University, Singapore\\\textsuperscript{2} LKC Medicine, Nanyang Technological University, Singapore\\\textsuperscript{3} National Institute of Education, Singapore\\\textsuperscript{4} Centre for Research and Development in Learning, Nanyang Technological University, Singapore}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} School of Psychology, University of Sussex, UK\\\textsuperscript{2} School of Social Sciences, Nanyang Technological University, Singapore\\\textsuperscript{3} LKC Medicine, Nanyang Technological University, Singapore\\\textsuperscript{4} National Institute of Education, Singapore\\\textsuperscript{5} Centre for Research and Development in Learning, Nanyang Technological University, Singapore}
 

### 2023-10-31 study2/experiment/demographics.js

@@ -1,0 +1,111 @@
+// Retrieve and save browser info
+var demographics_browser_info = {
+    type: jsPsychBrowserCheck,
+    data: {
+        screen: "browser_info",
+        date: new Date().toLocaleDateString("fr-FR"),
+        time: new Date().toLocaleTimeString("fr-FR"),
+    },
+    on_finish: function () {
+        data = jsPsych.data.get().filter({ screen: "browser_info" }).values()[0]
+        jsPsych.data.addProperties({
+            ["screen_height"]: data["height"],
+            ["screen_width"]: data["width"],
+        })
+        for (var key in data) {
+            if (
+                [
+                    "vsync_rate",
+                    "os",
+                    "mobile",
+                    "browser",
+                    "browser_version",
+                ].includes(key)
+            ) {
+                jsPsych.data.addProperties({
+                    [key]: data[key],
+                })
+            }
+        }
+        jsPsych.data.addProperties()
+    },
+}
+
+// Participant ID
+var demographics_participant_id = {
+    type: jsPsychSurveyText,
+    questions: [
+        {
+            prompt: "Enter participant ID:",
+            placeholder: "001",
+            name: "Participant_ID",
+        },
+    ],
+    data: {
+        screen: "participant_id",
+    },
+    on_finish: function () {
+        // Store `participant_id` so that it can be reused later
+        jsPsych.data.addProperties({
+            participant_id: jsPsych.data.get().last().values()[0]["response"][
+                "Participant_ID"
+            ],
+        })
+    },
+}
+
+// Consent form
+var demographics_consent = {
+    type: jsPsychHtmlButtonResponse,
+    stimulus:
+        // Logo
+        "<img src='https://blogs.brighton.ac.uk/sussexwrites/files/2019/06/University-of-Sussex-logo-transparent.png' width='150px' align='right'/><br><br><br><br><br>" +
+        // Title
+        "<h1>Informed Consent</h1>" +
+        "<p align='left'>Thank you for considering participating in our research 🤗 This study contains various questionnaires about your personality, feelings and current state of mind.</p>" +
+        "<p align='left'>Your participation in this research will be kept completely confidential. Your responses are entirely anonymous, and no IP address or any identifiers is collected.</p>" +
+        "<p align='left'><b>By participating, you agree to follow the instructions and provide honest answers.</b> If you do not wish to participate this survey, simply close your browser.</p>" +
[…]

### 2023-10-31 study2/experiment/questionnaires.js

@@ -1,0 +1,318 @@
+// Mini IPIP6 questionnaire
+var ipip6_items = [
+    // I am the life of the party
+    "Я - душа любой вечеринки",
+    // I sympathise with others' feelings
+    "Я сопереживаю эмоциям других людей",
+    // I get chores done right away
+    "Я сразу выполняю домашние дела",
+    // I have frequent mood swings
+    "Мое настроение часто меняется",
+    // I have a vivid imagination
+    "У меня богатое воображение",
+    // I feel entitled to more of everything
+    "Я считаю, что мне должно доставаться больше, чем то, что я получаю сейчас",
+    // I don't talk a lot
+    "Я не многословен/многословна",
+    // I am not interested in other people's problems
+    "Мне не интересны проблемы других людей",
+    // I have difficulty understanding abstract ideas
+    "Мне трудно понимать абстрактные идеи",
+    // I like order
+    "Я люблю порядок",
+    // I make a mess of things
+    "Я устраиваю беспорядок",
+    // I deserve more things in life
+    "Я заслуживаю большего в жизни",
+    // I do not have a good imagination
+    "У меня не очень хорошее воображение",
+    // I feel others' emotions
+    "Я чувствую эмоции других людей",
+    // I am relaxed most of the time
+    "Большую часть времени я расслаблен",
+    // I get upset easily
+    "Меня легко расстроить",
+    // I seldom feel blue
+    "Я редко чувствую себя грустным(ой)",
+    // I would like to be seen driving around in a really expensive car
+    "Я бы хотел(а), что бы меня видели за рулем очень дорогой машины",
+    // I keep in the background
+    "Я стараюсь оставаться в тени",
+    // I am not really interested in others
+    "Меня не очень интересуют другие люди",
+    // I am not interested in abstract ideas
+    "Меня не интересуют абстрактные идеи",
+    // I often forget to put things back in their proper place
+    "Я часто забываю положить вещи на место",
+    // I talk to a lot of different people at parties
+    "На вечеринках я общаюсь с большим количеством разных людей",
+    // I would get a lot of pleasure from owning expensive luxury goods
+    "Я получил(а) бы много удовольствия от владения дорогими предметами роскоши",
+]
+var ipip6_dimensions = [
+    "Extraversion_1",
+    "Agreeableness_2",
+    "Conscientiousness_3",
+    "Neuroticism_4",
+    "Openness_5",
+    "HonestyHumility_6_R",
+    "Extraversion_7_R",
[…]

### 2023-11-01 study2/experiment/demographics.js
update

@@ -1,4 +1,4 @@
-// Retrieve and save browser info
+// Retrieve and save browser info ========================================================
 var demographics_browser_info = {
     type: jsPsychBrowserCheck,
     data: {
@@ -31,7 +31,7 @@ var demographics_browser_info = {
     },
 }
 
-// Participant ID
+// Participant ID ========================================================================
 var demographics_participant_id = {
     type: jsPsychSurveyText,
     questions: [
@@ -54,7 +54,7 @@ var demographics_participant_id = {
     },
 }
 
-// Consent form
+// Consent form ========================================================================
 var demographics_consent = {
     type: jsPsychHtmlButtonResponse,
     stimulus:
@@ -62,7 +62,7 @@ var demographics_consent = {
         "<img src='https://blogs.brighton.ac.uk/sussexwrites/files/2019/06/University-of-Sussex-logo-transparent.png' width='150px' align='right'/><br><br><br><br><br>" +
         // Title
         "<h1>Informed Consent</h1>" +
-        "<p align='left'>Thank you for considering participating in our research 🤗 This study contains various questionnaires about your personality, feelings and current state of mind.</p>" +
+        "<p align='left'>Thank you for considering participating in our research. This study contains various questionnaires about your personality, feelings and current state of mind.</p>" +
         "<p align='left'>Your participation in this research will be kept completely confidential. Your responses are entirely anonymous, and no IP address or any identifiers is collected.</p>" +
         "<p align='left'><b>By participating, you agree to follow the instructions and provide honest answers.</b> If you do not wish to participate this survey, simply close your browser.</p>" +
         // "<p>Please note that various checks will be performed to ensure the validity of the data.<br>We reserve the right to return your participation or prorate reimbursement should we detect non-valid responses (e.g., random pattern of answers, instructions not read, ...).</p>"
@@ -71,17 +71,45 @@ var demographics_consent = {
 
     choices: ["I consent"],
     data: { screen: "consent" },
-    on_finish: function () {
-        jsPsych.data.addProperties({
-            participant_id: participant_id,
-            study_id: study_id,
-            session_id: session_id,
-        })
+}
+
+// Demographic info ========================================================================
+var demographics_multichoice = {
+    type: jsPsychSurveyMultiChoice,
+    questions: [
+        {
+            prompt: "What is your gender?",
+            options: ["Male", "Female", "Other"],
+            name: "gender",
+        },
+        // {
[…]

### 2023-11-01 study2/experiment/questionnaires.js

@@ -1,318 +1,82 @@
-// Mini IPIP6 questionnaire
-var ipip6_items = [
-    // I am the life of the party
-    "Я - душа любой вечеринки",
-    // I sympathise with others' feelings
-    "Я сопереживаю эмоциям других людей",
-    // I get chores done right away
-    "Я сразу выполняю домашние дела",
-    // I have frequent mood swings
-    "Мое настроение часто меняется",
-    // I have a vivid imagination
-    "У меня богатое воображение",
-    // I feel entitled to more of everything
-    "Я считаю, что мне должно доставаться больше, чем то, что я получаю сейчас",
-    // I don't talk a lot
-    "Я не многословен/многословна",
-    // I am not interested in other people's problems
-    "Мне не интересны проблемы других людей",
-    // I have difficulty understanding abstract ideas
-    "Мне трудно понимать абстрактные идеи",
-    // I like order
-    "Я люблю порядок",
-    // I make a mess of things
-    "Я устраиваю беспорядок",
-    // I deserve more things in life
-    "Я заслуживаю большего в жизни",
-    // I do not have a good imagination
-    "У меня не очень хорошее воображение",
-    // I feel others' emotions
-    "Я чувствую эмоции других людей",
-    // I am relaxed most of the time
-    "Большую часть времени я расслаблен",
-    // I get upset easily
-    "Меня легко расстроить",
-    // I seldom feel blue
-    "Я редко чувствую себя грустным(ой)",
-    // I would like to be seen driving around in a really expensive car
-    "Я бы хотел(а), что бы меня видели за рулем очень дорогой машины",
-    // I keep in the background
-    "Я стараюсь оставаться в тени",
-    // I am not really interested in others
-    "Меня не очень интересуют другие люди",
-    // I am not interested in abstract ideas
-    "Меня не интересуют абстрактные идеи",
-    // I often forget to put things back in their proper place
-    "Я часто забываю положить вещи на место",
-    // I talk to a lot of different people at parties
-    "На вечеринках я общаюсь с большим количеством разных людей",
-    // I would get a lot of pleasure from owning expensive luxury goods
-    "Я получил(а) бы много удовольствия от владения дорогими предметами роскоши",
-]
-var ipip6_dimensions = [
-    "Extraversion_1",
-    "Agreeableness_2",
-    "Conscientiousness_3",
-    "Neuroticism_4",
-    "Openness_5",
-    "HonestyHumility_6_R",
-    "Extraversion_7_R",
[…]

### 2023-12-31 study2/analysis/1_cleaning.qmd
add cleaning

---
title: "PHQ4R - Data Cleaning"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
```

```{r}
df <- read.csv("../data/data_raw.csv")
```

The initial sample consisted of `r report::report_participants(df, age="Age", gender="Gender", education="Education")`.

### Compute Scores

#### PHQ-4

```{r}
recode_phq <- function(x) {
  case_when(
    x == 1 ~ 0.5,
    x == 2 ~ 1,
    x == 3 ~ 2,
    x == 4 ~ 3,
    .default = x
  )
}

df <- df |>
  mutate(
    PHQ4_Anxiety_1 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Anxiety_1), PHQ4_Anxiety_1),
    PHQ4_Anxiety_2 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Anxiety_2), PHQ4_Anxiety_2),
    PHQ4_Depression_3 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Depression_3), PHQ4_Depression_3),
    PHQ4_Depression_4 = ifelse(PHQ4_Condition == "PHQ4 - Revised", recode_phq(PHQ4_Depression_4), PHQ4_Depression_4),
    PHQ4_Anxiety = PHQ4_Anxiety_1 + PHQ4_Anxiety_2,
    PHQ4_Depression = PHQ4_Depression_3 + PHQ4_Depression_4,
    PHQ4_Total = PHQ4_Anxiety + PHQ4_Depression
  )
```

```{r}
df |>
  select(matches("PHQ4_")) |>
  pivot_longer(cols = matches("Depression_|Anxiety_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count") +
  facet_grid(item ~ PHQ4_Condition) +
  labs(x = "Score", y = "N answers")
```

#### STAI-5

```{r}
df |>
  select(matches("STAI5_"), -STAI5_Duration, -STAI5_Order) |>
  pivot_longer(cols = matches("STAI5_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers")
```

```{r}
df <- df |> 
  mutate(STAI5_General = (STAI5_1 + STAI5_2 + STAI5_3 + STAI5_4 + STAI5_5) / 5)
```

#### BDI-2

```{r}
df |>
  select(matches("BDI2_"), -BDI2_Duration, -BDI2_Order) |>
  pivot_longer(cols = matches("BDI2_"), names_to = "item", values_to = "score") |>
  ggplot(aes(x = as.factor(score))) +
  geom_bar(stat="count") +
  facet_wrap(~item) +
  labs(x = "Score", y = "N answers")
```

```{r}
df <- df |> 
  mutate(BDI2_Total = rowSums(select(df, matches("BDI2_"), -BDI2_Duration, -BDI2_Order)))
```

### Recruitment History

```{r}
# Consecutive count of participants per day (as area)
df |>
  mutate(
    Date = as.Date(Date, format = "%d/%m/%Y"),
    N = 1:nrow(df)
  ) |>
  ggplot(aes(x = Date, y = N)) +
  geom_area() +
  scale_y_continuous(expand = c(0, 0)) +
  labs(
    title = "Recruitment History",
    x = "Date",
    y = "Total Number of Participants"
  ) +
  see::theme_modern()
```

## Outliers

### Experiment Duration

```{r}
df |>
  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
         Category = ifelse(Experiment_Duration > 50, "extra", "ok"),
         Duration = ifelse(Experiment_Duration > 50, 50, Experiment_Duration)) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_point(aes(color = Category, shape = Category)) +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  scale_color_manual(values = c("extra" = "red", ok = "black")) +
  guides(color = "none", shape = "none") +
  ggside::geom_xsidedensity(fill = "grey", color=NA) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Experiment Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  see::theme_modern() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3) 
```

## Descriptive Statistics

### Participants

The final sample includes `r report::report_participants(df, age="Age", gender="Gender", education="Education")`.

### Table 1

```{r}
table1 <- df |> 
  select(Age, 
         PHQ4_Condition, 
         STAI5 = STAI5_General,
         BDI2 = BDI2_Total) |> 
  report::report_sample(group_by = "PHQ4_Condition")

table1$Variable <- str_remove_all(table1$Variable, fixed("Mean "))
table1$Variable <- str_remove_all(table1$Variable, fixed(" (SD)"))

insight::display(table1)
```

```{r}
rbind(
  as.data.frame(report::report(t.test(df$Age ~ df$PHQ4_Condition))),
  as.data.frame(report::report(t.test(df$STAI5_General ~ df$PHQ4_Condition))),
  as.data.frame(report::report(t.test(df$BDI2_Total ~ df$PHQ4_Condition)))
) |> 
  select(Parameter, t, df_error, p) |> 
  mutate(Parameter = str_remove_all(Parameter, fixed("df$"))) |>
  insight::display()
```

### Distributions

```{r}
df |> 
  select(PHQ4_Condition, 
         STAI5 = STAI5_General,
         BDI2 = BDI2_Total,
         PHQ4_Total) |> 
  standardize() |> 
  estimate_density(at = "PHQ4_Condition") |> 
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color = Parameter)) +
  labs(x = "Standardized Score", y = "Distribution") +
  facet_wrap(~PHQ4_Condition)
```

## Save

```{r}
#| code-fold: false

write.csv(df, "../data/data.csv", row.names = FALSE)
```

### 2023-12-31 study2/analysis/0_preprocessing.py

@@ -1,25 +1,48 @@
 import json
-import os
 
-import osfclient
 import pandas as pd
 
+
+# Get files from OSF ======================================================
+def osf_listfiles(data_subproject="", token="", after_date=None):
+    try:
+        import osfclient
+    except ImportError:
+        raise ImportError("Please install 'osfclient' (`pip install osfclient`)")
+    osf = osfclient.OSF(token=token).project(data_subproject)  # Connect to project
+    storage = [s for s in osf.storages][0]  # Access storage component
+    files = [
+        {
+            "name": file.name.replace(".csv", ""),
+            "date": pd.to_datetime(file.date_created),
+            "url": file._download_url,
+            "size": file.size,
+            "file": file,
+        }
+        for file in storage.files
+    ]
+
+    if after_date is not None:
+        date = pd.to_datetime(after_date, format="%d/%m/%Y", utc=True)
+        files = [f for f, d in zip(files, [f["date"] > date for f in files]) if d]
+    return files
+
+
 token = ""  # Paste OSF token here to access private repositories
-data_subproject = "au695"  # Data subproject ID
-data_all = pd.DataFrame()  # Initialize empty dataframe
+files = osf_listfiles(
+    token=token,
+    data_subproject="au695",  # Data subproject ID
+    after_date="18/12/2023",
+)
 
-osf = osfclient.OSF(token=token).project(data_subproject)  # Connect to project
-storage = [s for s in osf.storages][0]  # Access storage component
-n_files = sum([1 for _ in storage.files])
-for i, file in enumerate(storage.files):
-    print(f"File N°{i+1}/{n_files}")
-    file_name = file.name.replace(".csv", "")
-    response = file._get(file._download_url, stream=True)
-    data = pd.read_csv(response.raw)
 
-    # Collection date
-    if data["date"].dropna().unique()[0] < "17/11/2023":
-        continue
+# Loop through files ======================================================
+alldata = pd.DataFrame()  # Initialize empty dataframe
+
+for i, file in enumerate(files):
+    print(f"File N°{i+1}/{len(files)}")
+
+    data = pd.read_csv(file["file"]._get(file["url"], stream=True).raw)
 
     # Participant ========================================================
[…]

### 2024-01-06 study2/analysis/1_cleaning.qmd
add cleaning

@@
 The initial sample consisted of `r report::report_participants(df, age="Age", gender="Gender", education="Education")`.
+
+### Recruitment History
+
+```{r}
+# Consecutive count of participants per day (as area)
+df |>
+  mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
+  arrange(Date) |>
+  mutate(N = 1:nrow(df)) |>
+  ggplot(aes(x = Date, y = N)) +
+  geom_area() +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(
+    title = "Recruitment History",
+    x = "Date",
+    y = "Total Number of Participants"
+  ) +
+  see::theme_modern()
+```
 
@@
 
-### Recruitment History
-
-```{r}
-# Consecutive count of participants per day (as area)
-df |>
-  mutate(
-    Date = as.Date(Date, format = "%d/%m/%Y"),
-    N = 1:nrow(df)
-  ) |>
-  ggplot(aes(x = Date, y = N)) +
-  geom_area() +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(
-    title = "Recruitment History",
-    x = "Date",
-    y = "Total Number of Participants"
-  ) +
-  see::theme_modern()
-```
-
 ## Outliers
@@
          PHQ4_Condition, 
+         PHQ4 = PHQ4_Total,
          STAI5 = STAI5_General,
@@
   as.data.frame(report::report(t.test(df$Age ~ df$PHQ4_Condition))),
+  as.data.frame(report::report(t.test(df$PHQ4_Total ~ df$PHQ4_Condition))),
   as.data.frame(report::report(t.test(df$STAI5_General ~ df$PHQ4_Condition))),

### 2024-01-06 study2/analysis/0_preprocessing.py

@@ -86,7 +86,7 @@ for i, file in enumerate(files):
 
     # Ethnicity
     race = demo2["ethnicity"].title().rstrip()
-    race = "Caucasian" if race in ["White"] else race
+    race = "Caucasian" if race in ["White", "White British"] else race
     race = "South Asian" if race in ["Pakistani"] else race
     race = "Arab" if race in ["Middle Eastern"] else race
     race = (
@@ -147,6 +147,7 @@ for i, file in enumerate(files):
     for item in ias:
         df[item] = ias[item]
 
+    # Save data ----------------------------------------------------------
     alldata = pd.concat([alldata, df], axis=0, ignore_index=True)
 
 # Save data ==============================================================

### 2024-01-30 study2/analysis/1_cleaning.qmd
update cleaning

@@
 ---
-title: "PHQ4R - Data Cleaning"
+title: "PHQ4R - Data Cleaning (Study 2)"
 editor: source
@@
 ```{r}
-df <- read.csv("../data/data_raw.csv")
+#| code-fold: false
+
+df <- read.csv("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data_raw.csv")
 ```
@@
 # Consecutive count of participants per day (as area)
+# df |>
+#   mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
+#   arrange(Date) |>
+#   mutate(N = 1:nrow(df)) |>
+#   ggplot(aes(x = Date, y = N)) +
+#   geom_area() +
+#   scale_y_continuous(expand = c(0, 0)) +
+#   labs(
+#     title = "Recruitment History",
+#     x = "Date",
+#     y = "Total Number of Participants"
+#   ) +
+#   see::theme_modern()
+```
+
+```{r}
+#| message: false
+
+# Consecutive count of participants per day (as area)
 df |>
   mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
-  arrange(Date) |>
-  mutate(N = 1:nrow(df)) |>
+  group_by(Date, Experimenter) |> 
+  summarize(N = n()) |> 
+  ungroup() |>
+  complete(Date, Experimenter, fill = list(N = 0)) |> 
+  group_by(Experimenter) |>
+  mutate(N = cumsum(N)) |>
   ggplot(aes(x = Date, y = N)) +
-  geom_area() +
+  geom_area(aes(fill=Experimenter)) +
   scale_y_continuous(expand = c(0, 0)) +
@@
 ```{r}
+#| echo: true
+#| code-fold: show
+
 recode_phq <- function(x) {
@@
   ggplot(aes(x = as.factor(score))) +
-  geom_bar(stat="count") +
+  geom_bar(stat="count", fill="#9C27B0") +
   facet_grid(item ~ PHQ4_Condition) +
-  labs(x = "Score", y = "N answers")
+  labs(x = "Score", y = "N answers") +
+  theme_bw()
 ```
@@
   ggplot(aes(x = as.factor(score))) +
-  geom_bar(stat="count") +
+  geom_bar(stat="count", fill="#E91E63") +
   facet_wrap(~item) +
-  labs(x = "Score", y = "N answers")
+  labs(x = "Score", y = "N answers") +
+  theme_bw()
 ```
@@
 ```{r}
+#| echo: true
+#| code-fold: show
+
 df <- df |> 
@@
   ggplot(aes(x = as.factor(score))) +
-  geom_bar(stat="count") +
+  geom_bar(stat="count", fill="#2196F3") +
   facet_wrap(~item) +
-  labs(x = "Score", y = "N answers")
+  labs(x = "Score", y = "N answers") +
+  theme_bw()
 ```
@@
 
-## Outliers
+## Exclusions
 
@@
   mutate(Participant = fct_reorder(Participant, Experiment_Duration),
-         Category = ifelse(Experiment_Duration > 50, "extra", "ok"),
-         Duration = ifelse(Experiment_Duration > 50, 50, Experiment_Duration)) |>
+         Category = ifelse(Experiment_Duration > 30, "extra", "ok"),
+         Duration = ifelse(Experiment_Duration > 30, 30, Experiment_Duration)) |>
   ggplot(aes(y = Participant, x = Duration)) +
+  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
   geom_point(aes(color = Category, shape = Category)) +
@@
   )  +
-  see::theme_modern() +
+  theme_bw() +
   ggside::theme_ggside_void() +
-  theme(ggside.panel.scale = .3) 
+  theme(ggside.panel.scale = .3,
+        panel.border = element_blank()) 
 ```
 
-## Descriptive Statistics
+## Final Sample
 
@@
   ggplot(aes(x=x, y=y)) +
-  geom_line(aes(color = Parameter)) +
+  geom_line(aes(color = Parameter), linewidth=2) +
+  scale_color_manual(values = c("STAI5" = "#E91E63", "BDI2" = "#2196F3", "PHQ4_Total" = "#9C27B0")) +
   labs(x = "Standardized Score", y = "Distribution") +
-  facet_wrap(~PHQ4_Condition)
+  facet_wrap(~PHQ4_Condition) +
+  theme_bw()
 ```

### 2024-01-30 study2/analysis/0_preprocessing.py

@@ -86,7 +86,7 @@ for i, file in enumerate(files):
 
     # Ethnicity
     race = demo2["ethnicity"].title().rstrip()
-    race = "Caucasian" if race in ["White", "White British"] else race
+    race = "Caucasian" if race in ["White", "White British", "White English"] else race
     race = "South Asian" if race in ["Pakistani"] else race
     race = "Arab" if race in ["Middle Eastern"] else race
     race = (

### 2024-02-10 study2/experiment/demographics.js
add MAIA

@@ -85,6 +85,38 @@ function demographics_consent(experimenter = "DEFAULT") {
     }
 }
 
+var demographics_consent_incentivized = {
+    type: jsPsychHtmlButtonResponse,
+    css_classes: ["narrow-text"],
+    stimulus:
+        // Logo and title
+        "<img src='https://blogs.brighton.ac.uk/sussexwrites/files/2019/06/University-of-Sussex-logo-transparent.png' width='150px' align='right'/><br><br><br><br><br>" +
+        "<h1>Informed Consent</h1>" +
+        // Overview
+        "<p align='left'><b>Invitation to Take Part</b><br>" +
+        "You are being invited to take part in a research study to further our understanding of Human psychology. Thank you for carefully reading this information sheet. This study is being conducted by Dr Dominique Makowski from the School of Psychology, University of Sussex, who is happy to be contacted (D.Makowski@sussex.ac.uk) if you have any questions.</p>" +
+        // Description
+        "<p align='left'><b>Why have I been invited and what will I do?</b><br>" +
+        "We are surveying adults to understand how mood fluctuations and mood disorders symptoms (or absence thereof) are expressed and what difficulties they can generate. This study contains various questionnaires about your personality, feelings and current state of mind. The whole experiment will take you <b>about 10 min</b> to complete. Please make you sure that you are in a quiet environment, and that you have time to complete it in one go.</p>" +
+        // Results and personal information
+        "<p align='left'><b>What will happen to the results and my personal information?</b><br>" +
+        "The results of this research may be written into a scientific publication. Your anonymity will be ensured in the way described in the consent information below. Please read this information carefully and then, if you wish to take part, please acknowledge that you have fully understood this sheet, and that you consent to take part in the study as it is described here.</p>" +
+        "<p align='left'><b>Consent</b><br></p>" +
+        // Bullet points
+        "<li align='left'>I understand that by signing below I am agreeing to take part in the University of Sussex research described here, and that I have read and understood this information sheet</li>" +
+        "<li align='left'>I understand that my participation is entirely voluntary, that I can choose not to participate in part or all of the study, and that I can withdraw at any stage by closing the browser without having to give a reason and without being penalised in any way (e.g., if I am a student, my decision whether or not to take part will not affect my grades).</li>" +
+        "<li align='left'>I understand that since the study is anonymous, it will be impossible to withdraw my data once I have completed and submitted the test/questionnaire.</li>" +
[…]

### 2024-02-10 study2/experiment/questionnaires.js

@@ -10,20 +10,12 @@ var PHQ4_items = [
     "Little interest or pleasure in doing things",
 ]
 
-var PHQ4_dimensions = [
-    "PHQ4_Anxiety_1",
-    "PHQ4_Anxiety_2",
-    "PHQ4_Depression_3",
-    "PHQ4_Depression_4",
-]
+var PHQ4_dimensions = ["PHQ4_Anxiety_1", "PHQ4_Anxiety_2", "PHQ4_Depression_3", "PHQ4_Depression_4"]
 
 // Questionnaire ========================================================================
 
 function phq4() {
-    condition = jsPsych.randomization.sampleWithoutReplacement(
-        ["PHQ4", "PHQ4R"],
-        1
-    )[0]
+    condition = jsPsych.randomization.sampleWithoutReplacement(["PHQ4", "PHQ4R"], 1)[0]
 
     if (condition == "PHQ4") {
         labels = [
@@ -80,12 +72,7 @@ for (const [index, element] of stai5_items.entries()) {
     stai5_questions.push({
         prompt: "<b>" + element + "</b>",
         name: stai5_dimensions[index],
-        labels: [
-            "<br>Not at all",
-            "<br>Somewhat",
-            "<br>Moderately so",
-            "<br>Very much so",
-        ],
+        labels: ["<br>Not at all", "<br>Somewhat", "<br>Moderately so", "<br>Very much so"],
         required: true,
     })
 }
@@ -114,6 +101,7 @@ var IAS_items = [
     "I can always accurately perceive when I need to urinate",
     "I can always accurately perceive when I need to defecate",
     "I can always accurately perceive when I encounter different tastes",
+    "I can always accurately perceive when I am about to blink", // Attentional check
     "I can always accurately perceive when I am going to vomit",
     "I can always accurately perceive when I am going to sneeze",
     "I can always accurately perceive when I am going to cough",
@@ -137,6 +125,7 @@ var IAS_dimensions = [
     "IAS_5",
     "IAS_6",
     "IAS_7",
+    "AttentionCheck_1",
     "IAS_8",
     "IAS_9",
     "IAS_10",
@@ -395,6 +384,17 @@ var bdi2 = {
             required: true,
         },
         {
+            prompt: "<b>19. Attention</b>", // Attention check
+            options: [
+                "0. I am not paying attention",
+                "1. I am cyrrebtky answering the questions at random",
+                "2. I am carefuly reading and answering the questions",
[…]

### 2024-02-12 study1/analysis.qmd
update analysis of study 1

---
title: "PHQ4R - Data Analysis (Study 1)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    fig-dpi: 300
    fig-width: 11.33
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(mirt)
```

```{r}
#| code-fold: false

df <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGameReliability/main/data/preprocessed_questionnaires.csv") |>
  select(Sex, Age, starts_with("Item_PHQ4")) |>
  filter(!is.na(Item_PHQ4_Anxiety_1))

names(df) <- str_remove_all(names(df), "Item_PHQ4_")

df2 <- read.csv("../study2/data/data_raw.csv") |>
  filter(PHQ4_Condition == "PHQ4 - Revised") |>
  select(Sex = Gender, Age, starts_with("PHQ4_"), -PHQ4_Duration, -PHQ4_Condition, -PHQ4_Order)

names(df2) <- str_remove_all(names(df2), "PHQ4_")
```

## Participants

- The first sample includes `r report::report_participants(df)`.
- The first sample includes `r report::report_participants(df2)`.

## Distribution

```{r}
add_labels <- function(x) {
  x <- case_when(
    x == 0 ~ "Not at all",
    x == 1 ~ "Once or twice",
    x == 2 ~ "Several days",
    x == 3 ~ "More than half the days",
    TRUE ~ "Nearly every day"
  )
  fct_relevel(x, c("Not at all", "Once or twice", "Several days", "More than half the days", "Nearly every day"))
}
```

::: {.panel-tabset}

### Sample 1

```{r}
#| message: false

df <- select(df, -Sex, -Age)

data <- df |>
  pivot_longer(everything(), names_to = "Item", values_to = "Answer") |>
  group_by(Item, Answer) |>
  summarise(n = n() / nrow(df)) |>
  mutate(
    Facet = str_split_fixed(Item, "_", 2)[, 1],
    # Item = str_split_fixed(Item, "_", 2)[,2],
    Answer = add_labels(Answer),
    Item = case_when(
      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
      Item == "Depression_3" ~ "D1: Feeling down, depressed, or hopeless",
      Item == "Depression_4" ~ "D2: Little interest or pleasure in doing things"
    )
  )

p1 <- data |>
  ggplot(aes(x = Answer, fill = Facet)) +
  geom_bar(aes(y = n), stat = "identity") +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
  labs(y = "Proportion of Answers") +
  theme_modern(axis.title.space = 10) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    plot.title = element_text(face = "bold"),
    strip.text = element_text(size = rel(0.7)),
    strip.background = element_rect(fill = "#EEEEEE", colour = "white")
  ) +
  facet_wrap(~Item)
p1 + ggtitle("Proportion of answers per item")
```

The "Once or twice" answer was selected in `r insight::format_percent(mean(filter(data, Answer == "Once or twice")$n))`

```{r}
#| message: false

p1b <- rbind(
  data.frame(Scores = paste0(df$Anxiety_1, "_", df$Anxiety_2),
             Facet = "Anxiety"),
  data.frame(Scores = paste0(df$Depression_3, "_", df$Depression_4),
             Facet = "Depression")
) |> 
  group_by(Facet, Scores) |> 
  summarize(n = n() / nrow(df)) |> 
  separate(Scores, into = c("Q1", "Q2")) |> 
  mutate(Label = ifelse(Q1 == Q2, format_percent(n), "")) |> 
  mutate(Q1 = add_labels(as.numeric(Q1)),
         Q2 = add_labels(as.numeric(Q2))) |> 
  ggplot(aes(x = Q1, y = Q2)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = Label), size=2, color="white") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_gradientn(colors=c("white", "#2196F3", "#3F51B5", "#673AB7"), labels = scales::percent) +
  labs(fill = "% Participants", x = "Item 1", y = "Item 2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        panel.grid.major = element_blank(),
        plot.title = element_text(face = "bold", hjust=0),
        strip.background=element_rect(fill="#EEEEEE", colour="white"),
        strip.text = element_text(face = "bold")) +
  facet_wrap(~Facet) 
p1b + ggtitle("Joint prevalence of responses")
```

### Sample 2

```{r}
#| message: false

df2 <- select(df2, -Sex, -Age)

data <- df2 |>
  pivot_longer(everything(), names_to = "Item", values_to = "Answer") |>
  group_by(Item, Answer) |>
  summarise(n = n() / nrow(df2)) |>
  mutate(
    Facet = str_split_fixed(Item, "_", 2)[, 1],
    # Item = str_split_fixed(Item, "_", 2)[,2],
    Answer = add_labels(Answer),
    Item = case_when(
      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
      Item == "Depression_3" ~ "D1: Feeling down, depressed, or hopeless",
      Item == "Depression_4" ~ "D2: Little interest or pleasure in doing things"
    )
  )

p1_s2 <- data |>
  ggplot(aes(x = Answer, fill = Facet)) +
  geom_bar(aes(y = n), stat = "identity") +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
  labs(y = "Proportion of Answers") +
  theme_modern(axis.title.space = 10) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    plot.title = element_text(face = "bold"),
    strip.text = element_text(size = rel(0.7)),
    strip.background = element_rect(fill = "#EEEEEE", colour = "white")
  ) +
  facet_wrap(~Item)
p1_s2 + ggtitle("Proportion of answers per item")
```

The "Once or twice" answer was selected in `r insight::format_percent(mean(filter(data, Answer == "Once or twice")$n))`

```{r}
#| message: false

p1b_s2 <- rbind(
  data.frame(Scores = paste0(df2$Anxiety_1, "_", df2$Anxiety_2),
             Facet = "Anxiety"),
  data.frame(Scores = paste0(df2$Depression_3, "_", df2$Depression_4),
             Facet = "Depression")
) |> 
  group_by(Facet, Scores) |> 
  summarize(n = n() / nrow(df2)) |> 
  separate(Scores, into = c("Q1", "Q2")) |> 
  mutate(Label = ifelse(Q1 == Q2, format_percent(n), "")) |> 
  mutate(Q1 = add_labels(as.numeric(Q1)),
         Q2 = add_labels(as.numeric(Q2))) |> 
  ggplot(aes(x = Q1, y = Q2)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = Label), size=2, color="white") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_gradientn(colors=c("white", "#2196F3", "#3F51B5", "#673AB7"), labels = scales::percent) +
  labs(fill = "% Participants", x = "Item 1", y = "Item 2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        panel.grid.major = element_blank(),
        plot.title = element_text(face = "bold", hjust=0),
        strip.background=element_rect(fill="#EEEEEE", colour="white"),
        strip.text = element_text(face = "bold")) +
  facet_wrap(~Facet) 
p1b_s2 + ggtitle("Joint prevalence of responses")
```

:::

## Anxiety

```{r}
df_anx <- select(df, contains("Anxiety"))
df_anx2 <- select(df2, contains("Anxiety"))
```

### Consistency

- Sample 1: Cronbach's alpha is `r insight::format_value(performance::cronbachs_alpha(df_anx), digits=3)`.
- Sample 2: Cronbach's alpha is `r insight::format_value(performance::cronbachs_alpha(df_anx2), digits=3)`.

### Model

::: {.panel-tabset}

#### Sample 1

```{r}
m_anxiety <- mirt(df_anx, model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)

m_anxiety

# Factor loadings
summary(m_anxiety)

# Alpha
a <- coef(m_anxiety)
a <- data.frame(
  Item = c("Anxiety 1", "Anxiety 2"),
  Discrimination = c(paste0(format_value(a$Anxiety_1[1, 1]), ", ", format_ci(a$Anxiety_1[2, 1], a$Anxiety_1[3, 1])), paste0(format_value(a$Anxiety_2[1, 1]), ", ", format_ci(a$Anxiety_2[2, 1], a$Anxiety_2[3, 1])))
)
knitr::kable(a)
```

```{r}
#| eval: false

# Plots
plot(m_anxiety, type = "trace", theta_lim = c(-3, 3))
plot(m_anxiety, type = "infotrace", theta_lim = c(-3, 3))
plot(m_anxiety, type = "score", theta_lim = c(-3, 3))
plot(m_anxiety, type = "infoSE", theta_lim = c(-3, 3))
plot(m_anxiety, type = "rxx", theta_lim = c(-3, 3))
```

#### Sample 2

```{r}
m_anxiety2 <- mirt(df_anx2, model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)

m_anxiety2

# Factor loadings
summary(m_anxiety2)

# Alpha
a <- coef(m_anxiety2)
a <- data.frame(
  Item = c("Anxiety 1", "Anxiety 2"),
  Discrimination = c(paste0(format_value(a$Anxiety_1[1, 1]), ", ", format_ci(a$Anxiety_1[2, 1], a$Anxiety_1[3, 1])), paste0(format_value(a$Anxiety_2[1, 1]), ", ", format_ci(a$Anxiety_2[2, 1], a$Anxiety_2[3, 1])))
)
knitr::kable(a)
```

:::

### Category Characteristic Curves (CRC) and Item Information Curves

Typically, an optimally informative item will have a large location and broad category coverage (as indicated by location parameters) over theta.

::: {.panel-tabset}

#### Sample 1

```{r}
crc <- function(mod, data) {
  Theta <- matrix(seq(-3.5, 3.5, by = .01))
  rez <- data.frame()
  for (i in 1:2) {
    dat <- as.data.frame(probtrace(extract.item(mod, i), Theta))
    dat$Theta <- Theta[, 1]
    dat$Information <- normalize(iteminfo(extract.item(mod, i), Theta))
    dat <- pivot_longer(dat, -one_of(c("Theta", "Information")), names_to = "Answer", values_to = "Probability")
    dat$Item <- names(data)[i]
    rez <- rbind(rez, dat)
  }

  rez <- rez |>
    mutate(
      Answer = case_when(
        Answer == "P.1" ~ "Not at all",
        Answer == "P.2" ~ "Once or twice",
        Answer == "P.3" ~ "Several days",
        Answer == "P.4" ~ "More than half the days",
        TRUE ~ "Nearly every day"
      ),
      Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days", "Nearly every day")),
      Item = case_when(
        Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
        Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
        Item == "Depression_3" ~ "D1: Feeling down, depressed, or hopeless",
        Item == "Depression_4" ~ "D2: Little interest or pleasure in doing things"
      )
    )

  # Get close to 0.95
  i1 <- rez[rez$Item == unique(rez$Item)[1], ]
  i2 <- rez[rez$Item == unique(rez$Item)[2], ]
  minmax <- rbind(
    rbind(
      i1[i1$Answer == "Not at all", ][which.min(abs(i1[i1$Answer == "Not at all", ]$Probability - 0.95)), ],
      i1[i1$Answer == "Nearly every day", ][which.min(abs(i1[i1$Answer == "Nearly every day", ]$Probability - 0.95)), ]
    ),
    rbind(
      i2[i2$Answer == "Not at all", ][which.min(abs(i2[i2$Answer == "Not at all", ]$Probability - 0.95)), ],
      i2[i2$Answer == "Nearly every day", ][which.min(abs(i2[i2$Answer == "Nearly every day", ]$Probability - 0.95)), ]
    )
  )

  p <- rez |>
    ggplot(aes(x = Theta, y = Probability)) +
    geom_line(aes(y = Information), linetype = "dotted", color = "grey") +
    geom_line(aes(color = Answer), linewidth = 1) +
    # geom_vline(data = minmax, aes(xintercept = Theta), linetype = "dashed") +
    scale_y_continuous(labels = scales::percent, expand = c(0, 0.005)) +
    scale_color_flat_d("rainbow") +
    facet_grid(~Item) +
    theme_modern(axis.title.space = 5) +
    theme(strip.background = element_rect(fill = "#EEEEEE", colour = "white"))
  list(p = p, rez = rez, minmax = minmax)
}

out <- crc(m_anxiety, df_anx)
p2a <- out$p + labs(x = expression(Anxiety ~ (theta)))
p2a
```

#### Sample 2

```{r}
out2 <- crc(m_anxiety2, df_anx2)
p2a_s2 <- out2$p + labs(x = expression(Anxiety ~ (theta)))
p2a_s2
```

:::

### Normalized scoring

::: {.panel-tabset}

#### Sample 1

```{r}
normalize_scores <- function(data, minmax) {
  minmax <- minmax[minmax$Theta %in% range(minmax$Theta), ]

  item <- data.frame()
  scores <- data.frame()
  for (i in unique(data$Item)) {
    item <- data[data$Item == i, ] |>
      filter(Theta >= min(minmax$Theta) & Theta <= max(minmax$Theta)) |>
      mutate(Theta_n = normalize(Theta)) |>
      rbind(item)

    scores <- item[item$Item == i, ] |>
      group_by(Answer) |>
      filter(Probability == max(Probability)) |>
      ungroup() |>
      mutate(label = insight::format_value(Theta_n)) |>
      rbind(scores)
  }

  item |>
    ggplot(aes(x = Theta, y = Probability, color = Answer)) +
    geom_line(linewidth = 1) +
    geom_segment(data = scores, aes(x = Theta, xend = Theta, y = 0, yend = Probability, color = Answer), linetype = "dashed") +
    geom_label(data = scores, aes(x = Theta, y = Probability, label = label), show.legend = FALSE) +
    scale_y_continuous(labels = scales::percent, expand = c(0, 0.01)) +
    scale_color_viridis_d(option = "D") +
    facet_grid(~Item) +
    theme_modern(axis.title.space = 5) +
    theme(strip.background = element_rect(fill = "#EEEEEE", colour = "white"))
}

p3a <- normalize_scores(data = out$rez, out$minmax) + labs(x = expression(Anxiety ~ (theta)))
p3a
```

#### Sample 2

```{r}
p3a_s2 <- normalize_scores(data = out2$rez, out2$minmax) + labs(x = expression(Anxiety ~ (theta)))
p3a_s2
```

:::

## Depression

```{r}
df_dep <- select(df, contains("Depression"))
df_dep2 <- select(df2, contains("Depression"))
```

### Consistency

- Sample 1: Cronbach's alpha is `r insight::format_value(performance::cronbachs_alpha(df_dep), digits=3)`.
- Sample 2: Cronbach's alpha is `r insight::format_value(performance::cronbachs_alpha(df_dep2), digits=3)`.

## Model

::: {.panel-tabset}

#### Sample 1

```{r}
m_depression <- mirt(df_dep, model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)

m_depression

# Factor loadings
summary(m_depression)

# Alpha
a2 <- coef(m_depression)
a2 <- data.frame(
  Item = c("Depression 1", "Depression 2"),
  Discrimination = c(paste0(format_value(a2$Depression_3[1, 1]), ", ", format_ci(a2$Depression_3[2, 1], a2$Depression_3[3, 1])), paste0(format_value(a2$Depression_4[1, 1]), ", ", format_ci(a2$Depression_4[2, 1], a2$Depression_4[3, 1])))
)
knitr::kable(a2)
```

#### Sample 2

```{r}
m_depression2 <- mirt(df_dep2, model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)

m_depression2

# Factor loadings
summary(m_depression2)

# Alpha
a2 <- coef(m_depression2)
a2 <- data.frame(
  Item = c("Depression 1", "Depression 2"),
  Discrimination = c(paste0(format_value(a2$Depression_3[1, 1]), ", ", format_ci(a2$Depression_3[2, 1], a2$Depression_3[3, 1])), paste0(format_value(a2$Depression_4[1, 1]), ", ", format_ci(a2$Depression_4[2, 1], a2$Depression_4[3, 1])))
)
knitr::kable(a2)
```

:::

## Category Characteristic Curves (CRC) and Item Information Curves

::: {.panel-tabset}

#### Sample 1

```{r}
out <- crc(m_depression, df_dep)
p2b <- out$p + labs(x = expression(Depression ~ (theta)))
p2b
```

#### Sample 2

```{r}
out2 <- crc(m_depression2, df_dep2)
p2b_s2 <- out2$p + labs(x = expression(Depression ~ (theta)))
p2b_s2
```

:::

## Normalized scoring

::: {.panel-tabset}

#### Sample 1

```{r}
p3b <- normalize_scores(out$rez, out$minmax) + labs(x = expression(Depression ~ (theta)))
p3b
```

#### Sample 2

```{r}
p3b_s2 <- normalize_scores(out2$rez, out2$minmax) + labs(x = expression(Depression ~ (theta)))
p3b_s2
```

:::

## Figures

::: {.panel-tabset}

#### Sample 1

```{r}
#| fig-width: 16.20
#| fig-height: 10

fig1 <- wrap_elements(wrap_elements(p1 + plot_annotation(title = "A. Proportion of answers per item", theme = list(plot.title = element_text(face = "bold")))) /
  wrap_elements(p1b + plot_annotation(title = "B. Joint prevalence of responses", theme = list(plot.title = element_text(face = "bold"))))) +
  wrap_elements(p2a / p2b + plot_annotation(title = "C. Information curves per response type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))

fig1
```

```{r}
#| fig-width: 16.20
#| fig-height: 10

fig2 <- p3a / p3b + plot_annotation(title = "Normalized Scoring", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")

fig2
```

```{r}
fig.width <- see::golden_ratio(7)
fig.height <- 7

ggsave("figures/figure1.png", fig1, width = fig.width * 1.43, height = fig.height * 1.43)
ggsave("figures/figure2.png", fig2, width = fig.width * 1.43, height = fig.height * 1.43)
```

#### Sample 2

```{r}
#| fig-width: 16.20
#| fig-height: 10

fig1_s2 <- wrap_elements(wrap_elements(p1_s2 + plot_annotation(title = "A. Proportion of answers per item", theme = list(plot.title = element_text(face = "bold")))) /
  wrap_elements(p1b_s2 + plot_annotation(title = "B. Joint prevalence of responses", theme = list(plot.title = element_text(face = "bold"))))) +
  wrap_elements(p2a_s2 / p2b_s2 + plot_annotation(title = "C. Information curves per response type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))
fig1_s2
```

```{r}
#| fig-width: 16.20
#| fig-height: 10

fig2_s2 <- p3a_s2 / p3b_s2 + plot_annotation(title = "Normalized Scoring", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")
fig2_s2
```

```{r}
ggsave("figures/figure1_s2.png", fig1_s2, width = fig.width * 1.43, height = fig.height * 1.43)
ggsave("figures/figure2_s2.png", fig2_s2, width = fig.width * 1.43, height = fig.height * 1.43)
```

:::

## Scores

::: {.panel-tabset}

#### Sample 1

```{r message=FALSE, warning=FALSE}
source("../score_PHQ4.R")

make_scores <- function(df) {
  data <- score_PHQ4(A1=df$Anxiety_1, A2=df$Anxiety_2, D1=df$Depression_3, D2=df$Depression_4, method="basic") |> 
    datawizard::data_addsuffix("_Basic") |> 
    cbind(
      score_PHQ4(A1=df$Anxiety_1, A2=df$Anxiety_2, D1=df$Depression_3, D2=df$Depression_4, method="normalized") |> 
        datawizard::data_addsuffix("_Normalized")) |> 
    select(matches("Depression|Anxiety")) |> 
    pivot_longer(everything()) |> 
    separate(name, into = c("Dimension", "Scoring"))

  data |> 
    ggplot(aes(x=value)) + 
    geom_area(data=estimate_density(data, at = c("Dimension", "Scoring"), method = "KernSmooth") |> 
                group_by(Scoring) |> 
                normalize(select = "y"),
              aes(x = x, y = y, fill=Dimension), alpha = 0.05, position="identity") +
    geom_hline(yintercept = 0.5, linetype = "dotted") +
    stat_ecdf(aes(color=Dimension), geom = "step", linewidth=1) +
    facet_wrap(~Scoring, scales = "free") +
    scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
    scale_color_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63")) +
    theme_modern(axis.title.space = 5) +
    labs(title = "Distribution and Cumulative Density of the Sample", x = "Score", y = "Cumulative Proportion")
}

make_scores(df)
```

#### Sample 2

```{r message=FALSE, warning=FALSE}
make_scores(df2)
```

:::

## Correlation

::: {.panel-tabset}

#### Sample 1

```{r message=FALSE, warning=FALSE}
data.frame(Depression = mirt::fscores(m_depression)[, 1],
           Anxiety = mirt::fscores(m_anxiety)[, 1]) |> 
  correlation::correlation(df) |>
  summary() |> 
  export_table(caption = "Correlation Matrix")
```

#### Sample 2

```{r message=FALSE, warning=FALSE}
data.frame(Depression = mirt::fscores(m_depression2)[, 1],
           Anxiety = mirt::fscores(m_anxiety2)[, 1]) |> 
  correlation::correlation(df2) |>
  summary() |> 
  export_table(caption = "Correlation Matrix")
```

:::

<!-- ## Multifactor model -->

<!-- ```{r message=FALSE, warning=FALSE, eval=FALSE} -->
<!-- s <- 'Anxiety = 1,2 -->
<!--       Depression = 3,4 -->
<!--       COV = Anxiety*Depression' -->
<!-- mod <- mirt(df, model = mirt.model(s), itemtype = 'graded', SE = TRUE) -->
<!-- mod <- mirt(df, model = 2, itemtype = 'graded', SE = TRUE) -->

<!-- summary(mod) -->

<!-- # Plots -->
<!-- plot(mod, type = 'trace', theta_lim = c(-3, 3)) -->
<!-- plot(mod, type = 'infotrace', theta_lim = c(-3, 3)) -->
<!-- plot(mod, type = 'score', theta_lim = c(-3, 3)) -->

<!-- Theta <- as.matrix(expand.grid(seq(-3, 3, by = 1), seq(-3, 3, by = 1))) -->
<!-- rez <- data.frame() -->
<!-- for(i in 1:4) { -->
<!--   dat <- as.data.frame(probtrace(extract.item(mod, i), Theta)) -->
<!--   dat$Theta_F1 <- Theta[, 1] -->
<!--   dat$Theta_F2 <- Theta[, 2] -->
<!--   # dat$Information <- normalize(iteminfo(extract.item(mod, i), Theta)) -->
<!--   dat <- pivot_longer(dat, -one_of(c("Theta_F1", "Theta_F2")), names_to = "Answer", values_to = "Probability") -->
<!--   dat$Item <- names(df)[i] -->
<!--   rez <- rbind(rez, dat) -->
<!-- } -->

<!-- rez <- rez |> -->
<!--   separate(Item, into = c("Dimension", "Item_Number")) |> -->
<!--   mutate( -->
<!--     Answer = case_when( -->
<!--     Answer == "P.1" ~ "Not at all", -->
<!--     Answer == "P.2" ~ "Once or twice", -->
<!--     Answer == "P.3" ~ "Several days", -->
<!--     Answer == "P.4" ~ "More than half the days", -->
<!--     TRUE ~ "Nearly every day" -->
<!--     )) |> -->
<!--   mutate( -->
<!--     Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day"))) -->

<!-- rez |> -->
<!--   ggplot(aes(x = Theta_F1, y = Probability)) + -->
<!--   # geom_line(aes(y = Information), linetype = "dotted", color = "grey") + -->
<!--   geom_line(aes(color = Answer)) + -->
<!--   geom_vline(xintercept = c(-1.25, 2.25), linetype = "dashed") + -->
<!--   facet_grid(Dimension~Item_Number) + -->
<!--   theme_modern() -->
<!-- ``` -->

### 2024-02-12 study2/analysis/1_cleaning.qmd
update analysis of study 1

@@
 ---
-title: "PHQ4R - Data Cleaning (Study 2)"
+title: "PHQ-4-R - Data Cleaning (Study 2)"
 editor: source
@@
 
-df <- read.csv("https://raw.githubusercontent.com/DominiqueMakowski/PHQ4R/main/study2/data/data_raw.csv")
+df <- read.csv("../data/data_raw.csv")
 ```
@@
 ### Recruitment History
-
-```{r}
-# Consecutive count of participants per day (as area)
-# df |>
-#   mutate(Date = as.Date(Date, format = "%d/%m/%Y")) |> 
-#   arrange(Date) |>
-#   mutate(N = 1:nrow(df)) |>
-#   ggplot(aes(x = Date, y = N)) +
-#   geom_area() +
-#   scale_y_continuous(expand = c(0, 0)) +
-#   labs(
-#     title = "Recruitment History",
-#     x = "Date",
-#     y = "Total Number of Participants"
-#   ) +
-#   see::theme_modern()
-```
 
@@
 
+# TODO: compute modified score
+# Maybe use the score_PHQ4 function from the README to demo it here
+
 recode_phq <- function(x) {
@@
   )
+
+df <- df |>
+  mutate(
+    Expe_PHQ4_Anxiety_1 = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised",
+      case_when(
+        PHQ4_Anxiety_1 == 0 ~ 0,
+        PHQ4_Anxiety_1 == 0.5 ~ 0.34,
+        PHQ4_Anxiety_1 == 1 ~ 0.55,
+        PHQ4_Anxiety_1 == 2 ~ 0.71,
+        PHQ4_Anxiety_1 == 3 ~ 1,
+        .default = PHQ4_Anxiety_1
+      ), PHQ4_Anxiety_1
+    ),
+    Expe_PHQ4_Anxiety_2 = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised",
+      case_when(
+        PHQ4_Anxiety_2 == 0 ~ 0,
+        PHQ4_Anxiety_2 == 0.5 ~ 0.44,
+        PHQ4_Anxiety_2 == 1 ~ 0.58,
+        PHQ4_Anxiety_2 == 2 ~ 0.71,
+        PHQ4_Anxiety_2 == 3 ~ 1,
+        .default = PHQ4_Anxiety_2
+      ), PHQ4_Anxiety_2
+    ),
+    Expe_PHQ4_Depression_3 = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised",
+      case_when(
+        PHQ4_Depression_3 == 0 ~ 0,
+        PHQ4_Depression_3 == 0.5 ~ 0.38,
+        PHQ4_Depression_3 == 1 ~ 0.51,
+        PHQ4_Depression_3 == 2 ~ 0.62,
+        PHQ4_Depression_3 == 3 ~ 1,
+        .default = PHQ4_Depression_3
+      ), PHQ4_Depression_3
+    ),
+    Expe_PHQ4_Depression_4 = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised",
+      case_when(
+        PHQ4_Depression_4 == 0 ~ 0,
+        PHQ4_Depression_4 == 0.5 ~ 0.35,
+        PHQ4_Depression_4 == 1 ~ 0.52,
+        PHQ4_Depression_4 == 2 ~ 0.66,
+        PHQ4_Depression_4 == 3 ~ 1,
+        .default = PHQ4_Depression_4
+      ), PHQ4_Depression_4
+    ),
+    Expe_PHQ4_Anxiety = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Expe_PHQ4_Anxiety_1 + Expe_PHQ4_Anxiety_2, range = c(0, 2), to = c(0, 6)), PHQ4_Anxiety_1 + PHQ4_Anxiety_2
+    ),
+    Expe_PHQ4_Depression = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Expe_PHQ4_Depression_3 + Expe_PHQ4_Depression_4, range = c(0, 2), to = c(0, 6)), PHQ4_Depression_3 + PHQ4_Depression_4
+    ),
+    Expe_PHQ4_Total = Expe_PHQ4_Anxiety + Expe_PHQ4_Depression
+  )
+
+# df |>
+#   select(starts_with("Expe_PHQ4_"), PHQ4_Condition) |>
+#   pivot_longer(cols = -PHQ4_Condition, names_to = "item", values_to = "score") |>
+#   ggplot(aes(x = score)) +
+#   geom_density(fill="#9C27B0", adjust=1/2)  +
+#   facet_grid(item ~ PHQ4_Condition)
 ```
@@
 df |>
-  select(matches("PHQ4_")) |>
+  select(starts_with("PHQ4_")) |>
   pivot_longer(cols = matches("Depression_|Anxiety_"), names_to = "item", values_to = "score") |>
@@
 
+<!-- df[is.na(df$STAI5_General), c("Participant")] -->
+
 #### BDI-2
@@
 
+<!-- df[is.na(df$BDI2_Total), c("Participant")] -->
+
 ## Exclusions
@@
 ### Experiment Duration
+
+The experiment's median duration is `r insight::format_value(median(df$Experiment_Duration))` min (`r insight::format_ci(bayestestR::hdi(df$Experiment_Duration, ci=0.5))`).
 
@@
   geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
-  geom_point(aes(color = Category, shape = Category)) +
+  geom_point(aes(color = Experimenter, shape = Category)) +
+  geom_text(aes(color = Experimenter, label=ifelse(Experiment_Duration < 4.5, as.character(Participant), '')), hjust=-1, vjust=0) +
   scale_shape_manual(values = c("extra" = 3, ok = 19)) +
-  scale_color_manual(values = c("extra" = "red", ok = "black")) +
-  guides(color = "none", shape = "none") +
-  ggside::geom_xsidedensity(fill = "grey", color=NA) +
+  guides(shape = "none") +
+  ggside::geom_xsidedensity(aes(fill=Experimenter), color=NA, alpha=0.5) +
   ggside::scale_xsidey_continuous(expand = c(0, 0)) +
@@
   theme(ggside.panel.scale = .3,
-        panel.border = element_blank()) 
+        panel.border = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank()) 
+```
+
+### Response Coherence
+
+```{r}
+SD_per_dim <- function(x, dims="") {
+  m <- matrix(nrow=nrow(x), ncol=0)
+  for(s in dims) {
+    m <- cbind(m, sapply(as.data.frame(t(x[grepl(s, names(x))])), sd))
+  }
+  m
+}
+
+df$BDI2_SD <- select(df, matches("BDI2_"), -BDI2_Duration, -BDI2_Order, -BDI2_Total) |> 
+  SD_per_dim(dims=c("BDI2_")) |> 
+  rowMeans()
+df$STAI5_SD <- select(df, matches("STAI5_"), -STAI5_Duration, -STAI5_Order, -STAI5_General) |> 
+  SD_per_dim(dims=c("STAI5_")) |> 
+  rowMeans()
+```
+
+```{r}
+outliers <- df |> 
+  select(Participant, BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
+  standardize(robust=TRUE) |> 
+  mutate(SD = rowSums(across(ends_with("SD"))), Duration = rowSums(across(ends_with("Duration"))))
+
+# df |> 
+#   ggplot(aes(x=BDI2_Duration, y=BDI2_SD)) +
+#   geom_point()
+
+table <- df |> 
+  select(Participant, BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
+  mutate_at(vars(ends_with("Duration")), \(x) ifelse(x > 10, 10, x)) |>
+  mutate(Participant = fct_reorder(Participant, outliers$Duration)) |> 
+  arrange(Participant)
+
+data.frame(Participant = c("Average"), t(sapply(table[2:ncol(table)], mean, na.rm = TRUE))) |>
+  rbind(table) |> 
+  gt::gt() |> 
+  gt::fmt_number() |> 
+  # gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% outliers_half, "orange", "white")) |>
+  gt::data_color(columns = ends_with("Duration"), 
+                 direction="column", palette = "RdBu") |> 
+  gt::data_color(columns = ends_with("SD"), 
+                 direction="column", palette = "RdBu", reverse=TRUE) |> 
+  gt::data_color(direction="row", rows=1, palette = "grey") |> 
+  gt::opt_interactive(use_compact_mode = TRUE)
+```
+
+### Attentional Checks
+
+```{r}
+df |> 
+  select(starts_with("IAS_"), -IAS_Duration, -IAS_Order) |>
+  pivot_longer(starts_with("IAS_"), names_to = "item", values_to = "score") |> 
+  ggplot(aes(x = score)) +
+  geom_density() +
+  facet_wrap(~item)
 ```
@@
 The final sample includes `r report::report_participants(df, age="Age", gender="Gender", education="Education")`.
+
+### Age
+
+```{r}
+p_age <- df |>  
+  mutate(Gender = ifelse(is.na(Gender), "Missing", Gender)) |>
+  ggplot(aes(x = x)) +
+  geom_histogram(aes(x = Age, fill=Gender), bins=28) +
+  geom_vline(xintercept = mean(df$Age), color = "red", linewidth=1.5) +
+  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
+  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
+  theme_modern(axis.title.space = 10) +
+  theme(
+    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
+    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
+    axis.text.y = element_text(size = rel(1.1)),
+    axis.text.x = element_text(size = rel(1.1)),
+    axis.title.x = element_blank()
+  )
+p_age
+```
+
+<!-- ## Ethnicity -->
+
+<!-- ```{r p_demographics_ethnicity} -->
+<!-- plot_waffle <- function(sub, what = "Nationality", title = what, rows = 8, size = 3) { -->
+<!--   ggwaffle::waffle_iron(sub, what, rows = rows) |> -->
+<!--     ggplot(aes(x, y)) + -->
+<!--     geom_point(aes(color = group), shape = "square", size = size) + -->
+<!--     coord_equal() + -->
+<!--     ggtitle(title) + -->
+<!--     labs(fill = "", color = "", subtitle = "N participants per self-declared group") + # (each square represents one participant)" -->
+<!--     theme_void() + -->
+<!--     theme( -->
+<!--       plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0), -->
+<!--       plot.subtitle = element_text(size = rel(1.2)), -->
+<!--       legend.key.height = unit(1, "mm"), -->
+<!--       legend.key.width = unit(1, "mm") -->
+<!--     ) -->
+<!-- } -->
+
+<!-- p_ethni <- plot_waffle(sub, "Ethnicity", rows = 20, size = 8) + -->
+<!--   scale_color_manual(values = c("Hispanic" = "#FF5722", "Caucasian" = "#2196F3", "African" = "#4CAF50", "Asian" = "#FFC107", "Other" = "#795548")) -->
+<!-- p_ethni -->
+<!-- ``` -->
+
+### Education
+
+```{r}
+p_edu <- df |>
+  mutate(
+    Education = ifelse(Education == "", NA, Education),
+    Education = fct_relevel(Education, "Other", "High School", "Bachelor", "Master", "Doctorate")) |> 
+  ggplot(aes(x = Education)) +
+  geom_bar(aes(fill = Education)) +
+  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
+  scale_fill_viridis_d(guide = "none") +
+  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
+  theme_modern(axis.title.space = 15) +
+  theme(
+    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
+    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
+    axis.text.y = element_text(size = rel(1.1)),
+    axis.text.x = element_text(size = rel(1.1)),
+    axis.title.x = element_blank()
+  )
+p_edu
+```
 
@@
 ```{r}
+add_bf <- function(r) {
+  x <- as.data.frame(r)
+  f <- as.formula(paste(str_remove(x$Parameter, fixed("df$")), "~ PHQ4_Condition"))
+  x$BF <- as.data.frame(BayesFactor::ttestBF(formula=f, data=df))$bf
+  x
+}
+
 rbind(
-  as.data.frame(report::report(t.test(df$Age ~ df$PHQ4_Condition))),
-  as.data.frame(report::report(t.test(df$PHQ4_Total ~ df$PHQ4_Condition))),
-  as.data.frame(report::report(t.test(df$STAI5_General ~ df$PHQ4_Condition))),
-  as.data.frame(report::report(t.test(df$BDI2_Total ~ df$PHQ4_Condition)))
+  add_bf(report::report(t.test(df$Age ~ df$PHQ4_Condition))),
+  add_bf(report::report(t.test(df$PHQ4_Total ~ df$PHQ4_Condition))),
+  add_bf(report::report(t.test(df$STAI5_General ~ df$PHQ4_Condition))),
+  add_bf(report::report(t.test(df$BDI2_Total ~ df$PHQ4_Condition)))
 ) |> 
-  select(Parameter, t, df_error, p) |> 
+  select(Parameter, t, df_error, p, BF) |> 
   mutate(Parameter = str_remove_all(Parameter, fixed("df$"))) |>
@@
 
+### Mood Disorders
+
+### Diagnostic
+
+```{r}
+p1 <- select(df, Participant, Gender, starts_with("Disorder_")) |> 
+  pivot_longer(cols = starts_with("Disorder_"), names_to = "Disorder", values_to = "Value") |> 
+  mutate(Disorder = str_remove_all(Disorder, fixed("Disorder_"))) |> 
+  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Disorder")) |> 
+  mutate(N_tot = sum(N), .by="Disorder") |>
+  mutate(Disorder = fct_reorder(Disorder, desc(N_tot))) |>
+  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
+  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
+  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p2 <- select(df, Participant, Gender, starts_with("DisorderTreatment_")) |> 
+  pivot_longer(cols = starts_with("DisorderTreatment_"), names_to = "Treatment", values_to = "Value") |> 
+  mutate(Treatment = str_remove_all(Treatment, fixed("DisorderTreatment_"))) |> 
+  summarize(N = sum(Value) / nrow(df), .by=c("Gender", "Treatment")) |> 
+  mutate(N_tot = sum(N), .by="Treatment") |>
+  mutate(Treatment = fct_reorder(Treatment, desc(N_tot))) |>
+  ggplot(aes(x = Treatment, y = N, fill=Gender)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
+  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
+  labs(title = "Treatment", y = "Number of Participants", subtitle = "Participants per treatment", x="Type") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1 / p2
+```
+
+```{r}
+df$MoodDisorder <- rowSums(select(df, DisorderTreatment_MoodStabilizer, DisorderTreatment_Antidepressant, DisorderTreatment_Anxiolytic))
+df$MoodDisorder <- ifelse(df$MoodDisorder > 0, TRUE, FALSE)
+```
+
+`r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment.
+
 ## Save

### 2024-02-12 study2/analysis/2_analysis.qmd
update analysis of study 1

---
title: "PHQ4R - Data Analysis (Study 2)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
---

## Data Preparation

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
```

```{r}
df <- read.csv("../data/data.csv") |> 
  mutate(
    PHQ4_Condition = fct_relevel(PHQ4_Condition, "PHQ4 - Original"),
    PHQ4_Condition = fct_recode(PHQ4_Condition, "Original" = "PHQ4 - Original", "Revised" = "PHQ4 - Revised"),
    BDI2_Category = ifelse(BDI2_Total <= 19, "Mild", "Moderate+"),
    STAI5_Category = ifelse(STAI5_General < 2, "Mild", "Moderate+"),
  )
```

## Depression

### Whole Sample

```{r}
model <- lm(BDI2_Total ~ PHQ4_Depression * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Depression, y = BDI2_Total), width=0.1) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(BDI2_Total ~ BDI2_Category / (PHQ4_Depression * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()

# model <- rstanarm::stan_glm(BDI2_Total ~ BDI2_Category / (PHQ4_Depression * PHQ4_Condition), data = df)
# parameters::parameters(model)
# bayestestR::bayesfactor_parameters(model)
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Depression, y = BDI2_Total)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=BDI2_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### General Score

```{r}
lm(BDI2_Total ~ PHQ4_Total * PHQ4_Condition, data = df) |> 
  parameters::parameters() |> 
  display()

lm(BDI2_Total ~ BDI2_Category / (PHQ4_Total * PHQ4_Condition), data = df) |> 
  parameters::parameters() |> 
  display()
```

```{r}
m0 <- lm(BDI2_Total ~ PHQ4_Depression, data = filter(df, PHQ4_Condition == "Original"))
m1 <- lm(BDI2_Total ~ PHQ4_Depression, data = filter(df, PHQ4_Condition == "Revised"))

performance::compare_performance(m0, m1)
```

## Anxiety

### Whole Sample

```{r}
model <- lm(STAI5_General ~ PHQ4_Anxiety * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(STAI5_General ~ STAI5_Category / (PHQ4_Anxiety * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=STAI5_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### General Score

```{r}
lm(STAI5_General ~ PHQ4_Total * PHQ4_Condition, data = df) |> 
  parameters::parameters() |> 
  display()

lm(STAI5_General ~ BDI2_Category / (PHQ4_Total * PHQ4_Condition), data = df) |> 
  parameters::parameters() |> 
  display()
```

## Correlation Difference

### Test

```{r}
dat <- df |> 
  group_by(PHQ4_Condition) |>
  summarize(
    N = n(),
    r_depression = cor.test(BDI2_Total, PHQ4_Depression)$estimate,
    r_anxiety = cor.test(STAI5_General, PHQ4_Anxiety)$estimate,
  )

psych::r.test(n=dat$N[1], n2=dat$N[2], r12=dat$r_depression[1], r34=dat$r_depression[2])
psych::r.test(n=dat$N[1], n2=dat$N[2], r12=dat$r_anxiety[1], r34=dat$r_anxiety[2])
```

### Bootstrapping

#### Whole Sample

```{r}
bootstrap_diffs <- function(df_original, df_revised) {
  dat <- data.frame()
  for(i in 1:1000) {
    newdf_ori <- df_original[sample(nrow(df_original), replace = TRUE), ]
    newdf_rev <- df_revised[sample(nrow(df_revised), replace = TRUE), ]

    r_depression_ori = cor.test(newdf_ori$BDI2_Total, newdf_ori$PHQ4_Depression)$estimate
    r_depression_rev = cor.test(newdf_rev$BDI2_Total, newdf_rev$PHQ4_Depression)$estimate
    r_anxiety_ori = cor.test(newdf_ori$STAI5_General, newdf_ori$PHQ4_Anxiety)$estimate
    r_anxiety_rev = cor.test(newdf_rev$STAI5_General, newdf_rev$PHQ4_Anxiety)$estimate

    dat <- rbind(dat,
      data.frame(
        d_r_depression = r_depression_rev - r_depression_ori,
        d_r_anxiety = r_anxiety_rev - r_anxiety_ori,
        iteration = i
      )
    )
  }
  dat
}

dat <- bootstrap_diffs(
  filter(df, PHQ4_Condition == "Original"),
  filter(df, PHQ4_Condition == "Revised"))

dat |> 
  pivot_longer(cols = c(d_r_depression, d_r_anxiety)) |>
  ggplot(aes(x = value)) +
  geom_density(aes(fill = name), alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(x = "Difference in Correlation", y = "Density") 
```

#### Subsamples

```{r}
dat <- data.frame()
for(cat in c("Mild", "Moderate+")) {
  dat <- rbind(
    dat,
    bootstrap_diffs(
      filter(df, PHQ4_Condition == "Original", BDI2_Category == cat),
      filter(df, PHQ4_Condition == "Revised", BDI2_Category == cat)) |> 
      mutate(Category = paste("Depression -", cat)),
    bootstrap_diffs(
      filter(df, PHQ4_Condition == "Original", STAI5_Category == cat),
      filter(df, PHQ4_Condition == "Revised", STAI5_Category == cat)) |> 
      mutate(Category = paste("Anxiety -", cat)) 
  )
}

dat |> 
  separate(Category, into = c("Type", "Category"), sep = " - ") |>
  pivot_longer(cols = c(d_r_depression, d_r_anxiety)) |>
  ggplot(aes(x = value)) +
  geom_density(aes(fill = name), alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  labs(x = "Difference in Correlation", y = "Density") +
  facet_grid(Category~Type)
```

## Predictive Value

### Mood Disorder

```{r}
glm(MoodDisorder ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(MoodDisorder ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(MoodDisorder ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()
```

### MDD

```{r}
glm(Disorder_MDD ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_MDD ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_MDD ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()
```

### GAD

```{r}
glm(Disorder_GAD ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_GAD ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()

glm(Disorder_GAD ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
  parameters::parameters() |> 
  display()
```

## Modified - Depression

### Whole Sample

```{r}
model <- lm(BDI2_Total ~ Expe_PHQ4_Depression * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Depression, y = BDI2_Total), width=0.1) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(BDI2_Total ~ BDI2_Category / (Expe_PHQ4_Depression * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Depression, y = BDI2_Total)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=BDI2_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

## Modified - Anxiety

### Whole Sample

```{r}
model <- lm(STAI5_General ~ Expe_PHQ4_Anxiety * PHQ4_Condition, data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### Subsamples

```{r}
model <- lm(STAI5_General ~ STAI5_Category / (Expe_PHQ4_Anxiety * PHQ4_Condition), data = df)

parameters::parameters(model) |> 
  display()
```

```{r}
df |> 
  ggplot(aes(x = Expe_PHQ4_Anxiety, y = STAI5_General)) +
  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
  geom_smooth(aes(color = PHQ4_Condition, linetype=STAI5_Category), method = "lm") +
  theme_bw() +
  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
  ggside::theme_ggside_void() 
```

### 2024-02-12 score_PHQ4.R

@@ -1,4 +1,21 @@
 score_PHQ4 <- function(A1, A2, D1, D2, method="basic") {
+
+  # Rescore ----
+  int_to_char <- function(x) {
+    x <- as.character(x)
+    x[x=="0"] <- "Not at all"
+    x[x=="1"] <- "Once or twice"
+    x[x=="2"] <- "Several days"
+    x[x=="3"] <- "More than half the days"
+    x[x=="4"] <- "Nearly every day"
+    x[x=="NA"] <- NA
+    x
+  }
+  if(is.numeric(A1)) A1 <- int_to_char(A1)
+  if(is.numeric(A2)) A2 <- int_to_char(A2)
+  if(is.numeric(D1)) D1 <- int_to_char(D1)
+  if(is.numeric(D2)) D2 <- int_to_char(D2)
+
   df <- data.frame(A1 = A1, A2 = A2, D1 = D1, D2 = D2)
   df <- as.data.frame(sapply(df, tolower, simplify=FALSE))
 

### 2024-02-12 study2/analysis/0_preprocessing.py

@@ -1,5 +1,6 @@
 import json
 
+import numpy as np
 import pandas as pd
 
 
@@ -28,7 +29,7 @@ def osf_listfiles(data_subproject="", token="", after_date=None):
     return files
 
 
-token = ""  # Paste OSF token here to access private repositories
+token = "zYboMoukFI8HKabenQ35DH6tESHJo6oZll5BvOPma6Dppjqc2jnIB6sPCERCuaqO0UrHAa"  # Paste OSF token here to access private repositories
 files = osf_listfiles(
     token=token,
     data_subproject="au695",  # Data subproject ID
@@ -38,6 +39,7 @@ files = osf_listfiles(
 
 # Loop through files ======================================================
 alldata = pd.DataFrame()  # Initialize empty dataframe
+prolific_ids = {}
 
 for i, file in enumerate(files):
     print(f"File N°{i+1}/{len(files)}")
@@ -48,40 +50,64 @@ for i, file in enumerate(files):
     # data["screen"].unique()
 
     # Browser info -------------------------------------------------------
-    brower = data[data["screen"] == "browser_info"].iloc[0]
+    browser = data[data["screen"] == "browser_info"].iloc[0]
+
+    # Experimenter
+    if "experimenter" in browser.index:
+        experimenter = browser["experimenter"]
+    else:
+        experimenter = browser["researcher"]
+    if "prolific_id" in browser.index:
+        if isinstance(browser["prolific_id"], str):
+            experimenter = "Prolific"
+    if isinstance(experimenter, float):
+        experimenter = "Experimenter" + str(int(experimenter))
 
     df = pd.DataFrame(
         {
             "Participant": file["name"],
-            "Experimenter": brower["experimenter"],
+            "Experimenter": experimenter,
             "Experiment_Duration": data["time_elapsed"].max() / 1000 / 60,
-            "Date": brower["date"],
-            "Time": brower["time"],
-            "Browser": brower["browser"],
-            "Mobile": brower["mobile"],
-            "Platform": brower["os"],
-            "Screen_Width": brower["screen_width"],
-            "Screen_Height": brower["screen_height"],
+            "Date_OSF": file["date"],
+            "Date": browser["date"],
[…]

### 2024-02-19 study2/analysis/1_cleaning.qmd
update cleaning and data

@@
   mutate(
-    Expe_PHQ4_Anxiety_1 = ifelse(
+    Weighted_PHQ4_Anxiety_1 = ifelse(
       PHQ4_Condition == "PHQ4 - Revised",
@@
     ),
-    Expe_PHQ4_Anxiety_2 = ifelse(
+    Weighted_PHQ4_Anxiety_2 = ifelse(
       PHQ4_Condition == "PHQ4 - Revised",
@@
     ),
-    Expe_PHQ4_Depression_3 = ifelse(
+    Weighted_PHQ4_Depression_3 = ifelse(
       PHQ4_Condition == "PHQ4 - Revised",
@@
     ),
-    Expe_PHQ4_Depression_4 = ifelse(
+    Weighted_PHQ4_Depression_4 = ifelse(
       PHQ4_Condition == "PHQ4 - Revised",
@@
     ),
-    Expe_PHQ4_Anxiety = ifelse(
-      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Expe_PHQ4_Anxiety_1 + Expe_PHQ4_Anxiety_2, range = c(0, 2), to = c(0, 6)), PHQ4_Anxiety_1 + PHQ4_Anxiety_2
+    Weighted_PHQ4_Anxiety = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Weighted_PHQ4_Anxiety_1 + Weighted_PHQ4_Anxiety_2, range = c(0, 2), to = c(0, 6)), PHQ4_Anxiety_1 + PHQ4_Anxiety_2
     ),
-    Expe_PHQ4_Depression = ifelse(
-      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Expe_PHQ4_Depression_3 + Expe_PHQ4_Depression_4, range = c(0, 2), to = c(0, 6)), PHQ4_Depression_3 + PHQ4_Depression_4
+    Weighted_PHQ4_Depression = ifelse(
+      PHQ4_Condition == "PHQ4 - Revised", datawizard::rescale(Weighted_PHQ4_Depression_3 + Weighted_PHQ4_Depression_4, range = c(0, 2), to = c(0, 6)), PHQ4_Depression_3 + PHQ4_Depression_4
     ),
-    Expe_PHQ4_Total = Expe_PHQ4_Anxiety + Expe_PHQ4_Depression
+    Weighted_PHQ4_Total = Weighted_PHQ4_Anxiety + Weighted_PHQ4_Depression
   )
@@
 # df |>
-#   select(starts_with("Expe_PHQ4_"), PHQ4_Condition) |>
+#   select(starts_with("Weighted_PHQ4_"), PHQ4_Condition) |>
 #   pivot_longer(cols = -PHQ4_Condition, names_to = "item", values_to = "score") |>
@@
   geom_point(aes(color = Experimenter, shape = Category)) +
-  geom_text(aes(color = Experimenter, label=ifelse(Experiment_Duration < 4.5, as.character(Participant), '')), hjust=-1, vjust=0) +
+  geom_text(aes(color = Experimenter, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
   scale_shape_manual(values = c("extra" = 3, ok = 19)) +
@@
 df$MoodDisorder <- ifelse(df$MoodDisorder > 0, TRUE, FALSE)
+df$DisorderHistory <- ifelse(df$DisorderHistory == "Yes", TRUE, FALSE)
 ```
@@
 `r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment.
+`r sum(df$DisorderHistory)` participants (`r insight::format_percent(sum(df$DisorderHistory) / nrow(df))`) reported having a history of mood disorder.
 

### 2024-02-27 study2/analysis/1_cleaning.qmd
add table
add progress bar

@@
 
-df <- read.csv("../data/data_raw.csv")
+df <- read.csv("../data/data_raw.csv") |> 
+   mutate(across(everything(), ~ifelse(.x == "", NA, .x))) 
 ```
@@
 
+```{r}
+#| message: false
+
+# Table
+summarize(df, N = n(), .by=c("Experimenter")) |> 
+  arrange(desc(N)) |> 
+  gt::gt() |> 
+  gt::opt_stylize() |> 
+  gt::opt_interactive(use_compact_mode = TRUE) |> 
+  gt::tab_header("Number of participants per recruitment source")
+```
+
 ### Compute Scores
@@
 #| code-fold: show
-
-# TODO: compute modified score
-# Maybe use the score_PHQ4 function from the README to demo it here
 

### 2024-03-01 study2/experiment/questionnaires.js
typo

@@ -387,7 +387,7 @@ var bdi2 = {
             prompt: "<b>19. Attention</b>", // Attention check
             options: [
                 "0. I am not paying attention",
-                "1. I am cyrrebtky answering the questions at random",
+                "1. I am currently answering the questions at random",
                 "2. I am carefuly reading and answering the questions",
                 "3. I am not reading to try finish the questionnaire quickly",
             ],

### 2024-05-20 study2/analysis/0_preprocessing.py
Update 0_preprocessing.py

@@ -29,7 +29,7 @@ def osf_listfiles(data_subproject="", token="", after_date=None):
     return files
 
 
-token = ""  # Paste OSF token here to access private repositories
+token = "zYboMoukFI8HKabenQ35DH6tESHJo6oZll5BvOPma6Dppjqc2jnIB6sPCERCuaqO0UrHAa"  # Paste OSF token here to access private repositories
 files = osf_listfiles(
     token=token,
     data_subproject="au695",  # Data subproject ID
@@ -116,9 +116,15 @@ for i, file in enumerate(files):
         continue
     if df["SONA_ID"].values[0] == 30884 and df["Mobile"].values[0] == True:
         continue
+    if df["SONA_ID"].values[0] == 30913 and df["Experiment_Duration"].values[0] < 10:
+        continue
     if df["SONA_ID"].values[0] == 31769 and df["Experiment_Duration"].values[0] > 15:
         continue
+    if df["SONA_ID"].values[0] == 31811 and df["Experiment_Duration"].values[0] < 15:
+        continue
     if df["SONA_ID"].values[0] == 31886 and df["Experiment_Duration"].values[0] < 10:
+        continue
+    if df["SONA_ID"].values[0] == 30733 and df["Experiment_Duration"].values[0] < 10:
         continue
     if df["SONA_ID"].values[0] == 31943 and df["Experiment_Duration"].values[0] < 10:
         continue
@@ -379,13 +385,19 @@ sona_credited = [
     30615,
     30616,
     30617,  # Check 2 failed
+    30624,
+    30626,
     30627,
+    30631,
     30636,
+    30652,
     30656,
+    30665,
     30682,
     30690,
     30697,
     30712,  # Not in list
+    # 30733,
     30742,
     30743,
     30744,
@@ -395,14 +407,23 @@ sona_credited = [
     30748,
     30758,
     30759,  # Not in list
+    30761,
     30765,
     30770,
     30783,  # Check 3 Failed
     30798,
+    30799,
     30834,
+    30838,
+    30846,  # Check 3 failed (awarded because email)
     30863,  # Not in list
     30867,
+    30873,
     30878,  # Not in list - Check 3 failed
+    30835,
+    30909,
+    30913,  # Failed checks
+    30926,
     30942,
     30957,  # Check 3 failed
     30970,
@@ -410,13 +431,22 @@ sona_credited = [
     30981,  # Not in list
     30986,
[…]

### 2024-05-20 study2/analysis/0_preprocessing.py
Update 0_preprocessing.py

@@ -29,7 +29,7 @@ def osf_listfiles(data_subproject="", token="", after_date=None):
     return files
 
 
-token = "zYboMoukFI8HKabenQ35DH6tESHJo6oZll5BvOPma6Dppjqc2jnIB6sPCERCuaqO0UrHAa"  # Paste OSF token here to access private repositories
+token = ""  # Paste OSF token here to access private repositories
 files = osf_listfiles(
     token=token,
     data_subproject="au695",  # Data subproject ID
@@ -108,9 +108,23 @@ for i, file in enumerate(files):
     # Filter duplicates
     if df["SONA_ID"].values[0] == 30609 and df["Experiment_Duration"].values[0] < 6:
         continue
-    if df["SONA_ID"].values[0] == 30679 and df["Mobile"].values[0] == True:
+    if df["SONA_ID"].values[0] == 30679 and df["Date"].values[0] != "22/03/2024":
+        continue
+    if df["SONA_ID"].values[0] == 30688 and df["Time"].values[0] != "13:02:42":
         continue
     if df["SONA_ID"].values[0] == 30746 and df["Mobile"].values[0] == True:
+        continue
+    if df["SONA_ID"].values[0] == 30746 and df["Date"].values[0] != "19/04/2024":
+        continue
+    if df["SONA_ID"].values[0] == 30759 and df["Experiment_Duration"].values[0] > 10:
+        continue
+    if df["SONA_ID"].values[0] == 31787 and df["Experiment_Duration"].values[0] < 8:
+        continue
+    if df["SONA_ID"].values[0] == 30796 and df["Mobile"].values[0] == True:
+        continue
+    if df["SONA_ID"].values[0] == 30796 and df["Experiment_Duration"].values[0] > 20:
+        continue
+    if df["SONA_ID"].values[0] == 30836 and df["Experiment_Duration"].values[0] < 10:
         continue
     if df["SONA_ID"].values[0] == 30878 and df["Experiment_Duration"].values[0] < 10:
         continue
@@ -118,7 +132,21 @@ for i, file in enumerate(files):
         continue
     if df["SONA_ID"].values[0] == 30913 and df["Experiment_Duration"].values[0] < 10:
         continue
+    if df["SONA_ID"].values[0] == 30939:
+        continue
+    if df["SONA_ID"].values[0] == 30979 and df["Experiment_Duration"].values[0] < 10:
+        continue
+    if df["SONA_ID"].values[0] == 31030 and df["Experiment_Duration"].values[0] < 10:
[…]

### 2024-05-21 study1/analysis.qmd
update

@@
 df_anx2 <- select(df2, contains("Anxiety"))
+df_anxfull <- rbind(df_anx, df_anx2)
 ```
@@
 knitr::kable(a)
+```
+
+```{r}
+#| eval: false
+
+# Plots
+plot(m_anxiety2, type = "trace", theta_lim = c(-3, 3))
+plot(m_anxiety2, type = "infotrace", theta_lim = c(-3, 3))
+plot(m_anxiety2, type = "score", theta_lim = c(-3, 3))
+plot(m_anxiety2, type = "infoSE", theta_lim = c(-3, 3))
+plot(m_anxiety2, type = "rxx", theta_lim = c(-3, 3))
+```
+
+#### Full
+
+```{r}
+m_anxietyfull <- mirt(df_anxfull, model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)
+
+plot(m_anxietyfull, type = "trace", theta_lim = c(-3, 3))
+# plot(m_anxietyfull, type = "infotrace", theta_lim = c(-3, 3))
 ```
@@
 
-## Model
+### Model
 
@@
 
-## Category Characteristic Curves (CRC) and Item Information Curves
+### Category Characteristic Curves (CRC) and Item Information Curves
 
@@
 
-## Normalized scoring
+### Normalized scoring
 

### 2024-05-21 study2/analysis/1_cleaning.qmd
update

@@
   gt::tab_header("Number of participants per recruitment source")
+```
+
+```{r}
+#| code-fold: false
+
+df$Sample <- case_when(
+  df$Experimenter %in% c("Experimenter1", "Experimenter2") ~ "Student",
+  df$Experimenter %in% c("dom", "Unknown", "readme") ~ "Other",
+  .default = df$Experimenter
+)
 ```
@@
   geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
-  geom_point(aes(color = Experimenter, shape = Category)) +
-  geom_text(aes(color = Experimenter, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
+  geom_point(aes(color = Sample, shape = Category)) +
+  geom_text(aes(color = Sample, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
   scale_shape_manual(values = c("extra" = 3, ok = 19)) +
   guides(shape = "none") +
-  ggside::geom_xsidedensity(aes(fill=Experimenter), color=NA, alpha=0.5) +
+  ggside::geom_xsidedensity(aes(fill=Sample), color=NA, alpha=0.5) +
   ggside::scale_xsidey_continuous(expand = c(0, 0)) +
@@
 ```
+
+### Attention Checks
+
+Only a subset of participants saw the version with attention checks.
+
+```{r}
+reject <- df |> 
+  filter(!is.na(AttentionCheck_2)) |>
+  select(Participant, Experimenter, starts_with("AttentionCheck")) |> 
+  arrange(AttentionCheck_2, AttentionCheck_3)
+
+failed <- reject[reject$AttentionCheck_2 == 0 | reject$AttentionCheck_3 < 0.7, "Participant"]
+failed <- failed[!is.na(failed)]
+
+reject |> 
+  gt::gt() |> 
+  gt::data_color(columns = c("AttentionCheck_2", "AttentionCheck_3"), palette=c("red", "green")) |> 
+  gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% failed, "red", "white")) |> 
+  gt::opt_interactive(use_compact_mode = TRUE)
+
+df <- filter(df, !Participant %in% failed)
+df <- select(df, -starts_with("AttentionCheck"))
+```
+
+We dropped `r length(failed)` participants for failing the attention checks.
 
@@
 
-# df |> 
-#   ggplot(aes(x=BDI2_Duration, y=BDI2_SD)) +
-#   geom_point()
+mv_outliers <- df |> 
+  select(Participant, 
+         matches("BDI2_"), -BDI2_Duration, -BDI2_Order, -BDI2_Total,
+         matches("STAI5_"), -STAI5_Duration, -STAI5_Order, -STAI5_General) |> 
+  check_outliers(method="mahalanobis_robust", ID="Participant")
+
+df |>
+  mutate(Outliers = mv_outliers) |> 
+  ggplot(aes(x=STAI5_SD, y=BDI2_SD)) +
+  geom_point(aes(color=Sample, shape=mv_outliers), alpha=2/3, size=3) +
+  scale_shape_manual(values=c(16, 4)) +
+  ggside::geom_xsidedensity(aes(color=Sample), alpha=0.5) +
+  ggside::geom_ysidedensity(aes(color=Sample), alpha=0.5) +
+  theme_minimal() +
+  ggside::theme_ggside_void()
 
 table <- df |> 
-  select(Participant, BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
+  select(Participant, starts_with("AttentionCheck"), BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
   mutate_at(vars(ends_with("Duration")), \(x) ifelse(x > 10, 10, x)) |>
@@
   gt::opt_interactive(use_compact_mode = TRUE)
+
+df <- df[!mv_outliers,]
 ```
 
-### Attentional Checks
-
-```{r}
-df |> 
-  select(starts_with("IAS_"), -IAS_Duration, -IAS_Order) |>
-  pivot_longer(starts_with("IAS_"), names_to = "item", values_to = "score") |> 
-  ggplot(aes(x = score)) +
-  geom_density() +
-  facet_wrap(~item)
-```
+We dropped `r length(mv_outliers)` participants flagged as multivariate outliers using the robust Mahalanobis distance.
 
@@
 ```{r}
-add_bf <- function(r) {
+add_bf <- function(df, var="Age") {
+  r <- report::report(t.test(df[[var]] ~ df[["PHQ4_Condition"]]), verbose=FALSE)
   x <- as.data.frame(r)
-  f <- as.formula(paste(str_remove(x$Parameter, fixed("df$")), "~ PHQ4_Condition"))
+  f <- as.formula(paste(var, "~ PHQ4_Condition"))
   x$BF <- as.data.frame(BayesFactor::ttestBF(formula=f, data=df))$bf
+  x$Parameter <- var
   x
@@
 rbind(
-  add_bf(report::report(t.test(df$Age ~ df$PHQ4_Condition))),
-  add_bf(report::report(t.test(df$PHQ4_Total ~ df$PHQ4_Condition))),
-  add_bf(report::report(t.test(df$STAI5_General ~ df$PHQ4_Condition))),
-  add_bf(report::report(t.test(df$BDI2_Total ~ df$PHQ4_Condition)))
+  add_bf(filter(df, !is.na(Age)), "Age"),
+  add_bf(df, var="PHQ4_Total"),
+  add_bf(df, "STAI5_General"),
+  add_bf(df, "BDI2_Total")
 ) |> 
@@
 `r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment.
-`r sum(df$DisorderHistory)` participants (`r insight::format_percent(sum(df$DisorderHistory) / nrow(df))`) reported having a history of mood disorder.
+`r sum(df$DisorderHistory, na.rm=TRUE)` participants (`r insight::format_percent(sum(df$DisorderHistory, na.rm=TRUE) / nrow(df))`) reported having a history of mood disorder.
 

### 2024-12-06 study2/analysis/1_cleaning.qmd
update data
@AnafNeves

@@
 df <- read.csv("../data/data_raw.csv") |> 
-   mutate(across(everything(), ~ifelse(.x == "", NA, .x))) 
+  mutate(across(everything(), ~ifelse(.x == "", NA, .x))) |> 
+  filter(Age >= 18) |> 
+  mutate(Age = ifelse(Age > 90, NA, Age))
 ```
@@
 )
+```
+
+### Experiment Duration
+
+The experiment's median duration is `r insight::format_value(median(df$Experiment_Duration))` min (`r insight::format_ci(bayestestR::hdi(df$Experiment_Duration, ci=0.5))`).
+
+```{r}
+df |>
+  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
+         Category = ifelse(Experiment_Duration > 30, "extra", "ok"),
+         Duration = ifelse(Experiment_Duration > 30, 30, Experiment_Duration)) |>
+  ggplot(aes(y = Participant, x = Duration)) +
+  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
+  geom_point(aes(color = Sample, shape = Category)) +
+  geom_text(aes(color = Sample, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
+  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
+  guides(shape = "none") +
+  ggside::geom_xsidedensity(aes(fill=Sample), color=NA, alpha=0.5) +
+  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
+  labs(
+    title = "Experiment Completion Time",
+    x = "Duration (in minutes)",
+    y = "Participant"
+  )  +
+  theme_bw() +
+  ggside::theme_ggside_void() +
+  theme(ggside.panel.scale = .3,
+        panel.border = element_blank(),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank()) 
 ```
@@
 
-<!-- df[is.na(df$BDI2_Total), c("Participant")] -->
+#### IAS
+
+```{r}
+#| echo: true
+#| code-fold: show
+
+df <- df |> 
+  mutate(IAS_Total = rowSums(select(df, matches("IAS_"), -IAS_Duration, -IAS_Order)))
+```
 
@@
 
-### Experiment Duration
-
-The experiment's median duration is `r insight::format_value(median(df$Experiment_Duration))` min (`r insight::format_ci(bayestestR::hdi(df$Experiment_Duration, ci=0.5))`).
-
 ```{r}
-df |>
-  mutate(Participant = fct_reorder(Participant, Experiment_Duration),
-         Category = ifelse(Experiment_Duration > 30, "extra", "ok"),
-         Duration = ifelse(Experiment_Duration > 30, 30, Experiment_Duration)) |>
-  ggplot(aes(y = Participant, x = Duration)) +
-  geom_vline(xintercept = median(df$Experiment_Duration), color = "red", linetype = "dashed") +
-  geom_point(aes(color = Sample, shape = Category)) +
-  geom_text(aes(color = Sample, label=ifelse(Experiment_Duration < 3, as.character(Participant), '')), hjust=-1, vjust=0) +
-  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
-  guides(shape = "none") +
-  ggside::geom_xsidedensity(aes(fill=Sample), color=NA, alpha=0.5) +
-  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  labs(
-    title = "Experiment Completion Time",
-    x = "Duration (in minutes)",
-    y = "Participant"
-  )  +
-  theme_bw() +
-  ggside::theme_ggside_void() +
-  theme(ggside.panel.scale = .3,
-        panel.border = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.text.y = element_blank()) 
+outliers <- list()
 ```
@@
 
-failed <- reject[reject$AttentionCheck_2 == 0 | reject$AttentionCheck_3 < 0.7, "Participant"]
-failed <- failed[!is.na(failed)]
+outliers$attentionchecks <- reject[reject$AttentionCheck_2 == 0 | reject$AttentionCheck_3 < 0.95, "Participant"]
+outliers$attentionchecks <- outliers$attentionchecks[!is.na(outliers$attentionchecks)]
+df$AttentionCheck <- ifelse(df$Participant %in% outliers$attentionchecks, "Failed", "Passed")
 
@@
   gt::data_color(columns = c("AttentionCheck_2", "AttentionCheck_3"), palette=c("red", "green")) |> 
-  gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% failed, "red", "white")) |> 
+  gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% outliers$attentionchecks, "red", "white")) |> 
   gt::opt_interactive(use_compact_mode = TRUE)
-
-df <- filter(df, !Participant %in% failed)
-df <- select(df, -starts_with("AttentionCheck"))
 ```
 
-We dropped `r length(failed)` participants for failing the attention checks.
+We removed `r length(outliers$attentionchecks)` participants for failing the attention checks.
+
+```{r}
+#| eval: false
+
+dat <- mutate(df, Failed = ifelse(Participant %in% outliers$attentionchecks, 1, 0)) |> 
+  filter(Gender %in% c("Female", "Male")) |> 
+  filter(Experiment_Duration < 30) |> 
+  mutate(Gender = as.factor(Gender)) |> 
+  select(Failed, Gender, Experiment_Duration, Age, Ethnicity)
+
+m <- glm(Failed ~ Gender / Age, data=dat, family="binomial") 
+parameters::parameters(m)
+
+modelbased::estimate_relation(m, length=50) |> 
+  ggplot(aes(x=Age, y=Predicted, color=Gender)) +
+  geom_line() 
+```
+
+### Duration
+
+```{r}
+m <- mgcv::gam(Failed ~ s(Experiment_Duration), 
+          data=df |> 
+            filter(Experiment_Duration < 30) |> 
+            mutate(Failed = ifelse(AttentionCheck == "Failed", 1, 0)), 
+          family="binomial")
+
+parameters::parameters(m)
+plot(estimate_relation(m, length=40))
+
+df |> 
+  filter(Experiment_Duration < 30) |> 
+  select(Participant, Experiment_Duration, AttentionCheck) |> 
+  mutate(Participant = fct_reorder(Participant, Experiment_Duration)) |>
+  ggplot(aes(x=Participant, y=Experiment_Duration, fill=AttentionCheck)) +
+  geom_bar(stat="identity") +
+  theme_minimal() +
+  theme(axis.text.x = element_blank())
+```
 
@@
 ```{r}
-outliers <- df |> 
-  select(Participant, BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
-  standardize(robust=TRUE) |> 
-  mutate(SD = rowSums(across(ends_with("SD"))), Duration = rowSums(across(ends_with("Duration"))))
+#| eval: false
 
-mv_outliers <- df |> 
-  select(Participant, 
-         matches("BDI2_"), -BDI2_Duration, -BDI2_Order, -BDI2_Total,
-         matches("STAI5_"), -STAI5_Duration, -STAI5_Order, -STAI5_General) |> 
-  check_outliers(method="mahalanobis_robust", ID="Participant")
+mgcv::gam(Failed ~ s(STAI5_SD), 
+          data=mutate(df, Failed = ifelse(AttentionCheck == "Failed", 1, 0)), 
+          family="binomial") |>
+  parameters::parameters()
 
 df |> 
-  mutate(Outliers = mv_outliers) |> 
-  ggplot(aes(x=STAI5_SD, y=BDI2_SD)) +
-  geom_point(aes(color=Sample, shape=mv_outliers), alpha=2/3, size=3) +
-  scale_shape_manual(values=c(16, 4)) +
-  ggside::geom_xsidedensity(aes(color=Sample), alpha=0.5) +
-  ggside::geom_ysidedensity(aes(color=Sample), alpha=0.5) +
+  select(Participant, STAI5_SD, AttentionCheck) |> 
+  mutate(Participant = fct_reorder(Participant, STAI5_SD)) |>
+  ggplot(aes(x=Participant, y=STAI5_SD, fill=AttentionCheck)) +
+  geom_bar(stat="identity") +
   theme_minimal() +
-  ggside::theme_ggside_void()
-
-table <- df |> 
-  select(Participant, starts_with("AttentionCheck"), BDI2_SD, BDI2_Duration, STAI5_SD, STAI5_Duration) |> 
-  mutate_at(vars(ends_with("Duration")), \(x) ifelse(x > 10, 10, x)) |>
-  mutate(Participant = fct_reorder(Participant, outliers$Duration)) |> 
-  arrange(Participant)
-
-data.frame(Participant = c("Average"), t(sapply(table[2:ncol(table)], mean, na.rm = TRUE))) |>
-  rbind(table) |> 
-  gt::gt() |> 
-  gt::fmt_number() |> 
-  # gt::data_color(columns = "Participant", fn=\(x) ifelse(x %in% outliers_half, "orange", "white")) |>
-  gt::data_color(columns = ends_with("Duration"), 
-                 direction="column", palette = "RdBu") |> 
-  gt::data_color(columns = ends_with("SD"), 
-                 direction="column", palette = "RdBu", reverse=TRUE) |> 
-  gt::data_color(direction="row", rows=1, palette = "grey") |> 
-  gt::opt_interactive(use_compact_mode = TRUE)
-
-df <- df[!mv_outliers,]
+  theme(axis.text.x = element_blank())
 ```
 
-We dropped `r length(mv_outliers)` participants flagged as multivariate outliers using the robust Mahalanobis distance.
+### Multivariate Distance
+
+```{r}
+dfoutlier <- df |> 
+  select(contains("IAS_"), -IAS_Duration, -IAS_Order, -IAS_Total) |>
+  # select(contains("BDI2_"), -BDI2_Duration, -BDI2_Order, -BDI2_Total, -BDI2_SD) |> 
+  # datawizard::remove_empty_rows() |> 
+  performance::check_outliers(
+    method=c("optics", "mahalanobis_robust"),
+    threshold=list(optics=2.5, optics_xi=0.03)) |> 
+  as.data.frame() |> 
+  mutate(Participant = fct_reorder(df$Participant, Distance_OPTICS),
+         Outlier_AttentionCheck = ifelse(Participant %in% outliers$attentionchecks, 1, 0),
+         Outlier_Distance = ifelse(abs(standardize(Distance_OPTICS)) > qnorm(0.995), 1, 0),
+         Experiment_Duration = df$Experiment_Duration,
+         Outlier = ifelse(Outlier_AttentionCheck == 1, "Failed Attention Checks", "Passed"))
+
+outliers$distance <- as.character(dfoutlier[dfoutlier$Outlier_Distance == 1, "Participant"])
+outliers$distance <- outliers$distance[!outliers$distance %in% c(outliers$attentionchecks, outliers$duration)]
+
+m <- mgcv::gam(Outlier_AttentionCheck ~ s(Distance_OPTICS), data = dfoutlier, family = "binomial")
+parameters::parameters(m)
+
+estimate_relation(m, length=40) |> 
+  ggplot(aes(x=Distance_OPTICS, y=Predicted)) +
+  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha=0.2) +
+  geom_line() + 
+  ggside::geom_xsidedensity(data=dfoutlier, aes(fill=Outlier), alpha=0.3, color=NA)
+
+dfoutlier |> 
+  ggplot(aes(x=Participant, y=Distance_OPTICS, fill=Outlier)) +
+  geom_bar(stat="identity") +
+  theme_minimal() +
+  theme(axis.text.x = element_blank())
+```
+
+```{r}
+df <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$distance))
+df <- select(df, -starts_with("AttentionCheck"))
+```
 
@@
 ```{r}
+df$PHQ4_Duration[df$PHQ4_Duration > 3] <- NA
+
 table1 <- df |> 
@@
          PHQ4_Condition, 
-         PHQ4 = PHQ4_Total,
-         STAI5 = STAI5_General,
-         BDI2 = BDI2_Total) |> 
-  report::report_sample(group_by = "PHQ4_Condition")
+         PHQ4_Total,
+         PHQ4_Duration,
+         STAI5_General,
+         BDI2_Total) |> 
+  report::report_sample(by = "PHQ4_Condition")
 
@@
 table1$Variable <- str_remove_all(table1$Variable, fixed(" (SD)"))
-
-insight::display(table1)
+table1$Variable <- c("Age", "PHQ4 (Total)", "PHQ4 (Duration)", "STAI5", "BDI2")
 ```
@@
 ```{r}
-add_bf <- function(df, var="Age") {
-  r <- report::report(t.test(df[[var]] ~ df[["PHQ4_Condition"]]), verbose=FALSE)
+add_bf <- function(df, var="PHQ4_Duration") {
+  dat <- df[, c(var, "PHQ4_Condition")]
+  dat <- dat[complete.cases(dat), ]
+  r <- report::report(t.test(dat[[var]] ~ dat[["PHQ4_Condition"]]), verbose=FALSE)
   x <- as.data.frame(r)
   f <- as.formula(paste(var, "~ PHQ4_Condition"))
-  x$BF <- as.data.frame(BayesFactor::ttestBF(formula=f, data=df))$bf
+  x$BF <- as.data.frame(BayesFactor::ttestBF(formula=f, data=dat))$bf
   x$Parameter <- var
@@
 
-rbind(
+table1_inference <- rbind(
   add_bf(filter(df, !is.na(Age)), "Age"),
   add_bf(df, var="PHQ4_Total"),
+  add_bf(df, var="PHQ4_Duration"),
   add_bf(df, "STAI5_General"),
@@
   select(Parameter, t, df_error, p, BF) |> 
-  mutate(Parameter = str_remove_all(Parameter, fixed("df$"))) |>
-  insight::display()
+  mutate(Parameter = str_remove_all(Parameter, fixed("df$")))
+
+table1$BF <- table1_inference$BF
+insight::display(table1)
 ```
@@
   standardize() |> 
-  estimate_density(at = "PHQ4_Condition") |> 
+  estimate_density(by = "PHQ4_Condition", method="KernSmooth") |> 
   ggplot(aes(x=x, y=y)) +
@@
 df$MoodDisorder <- ifelse(df$MoodDisorder > 0, TRUE, FALSE)
+
+df$Depression <- rowSums(select(df, DisorderTreatment_MoodStabilizer, DisorderTreatment_Antidepressant))
+df$Depression <- ifelse(df$Depression > 0, TRUE, FALSE)
+
 df$DisorderHistory <- ifelse(df$DisorderHistory == "Yes", TRUE, FALSE)
@@
 `r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment.
+`r sum(df$Depression)` participants (`r insight::format_percent(sum(df$Depression) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer or antidepressant.
 `r sum(df$DisorderHistory, na.rm=TRUE)` participants (`r insight::format_percent(sum(df$DisorderHistory, na.rm=TRUE) / nrow(df))`) reported having a history of mood disorder.

### 2024-12-06 study2/analysis/0_preprocessing.py

@@ -1,40 +1,12 @@
 import json
-
+import os
 import numpy as np
 import pandas as pd
 
 
-# Get files from OSF ======================================================
-def osf_listfiles(data_subproject="", token="", after_date=None):
-    try:
-        import osfclient
-    except ImportError:
-        raise ImportError("Please install 'osfclient' (`pip install osfclient`)")
-    osf = osfclient.OSF(token=token).project(data_subproject)  # Connect to project
-    storage = [s for s in osf.storages][0]  # Access storage component
-    files = [
-        {
-            "name": file.name.replace(".csv", ""),
-            "date": pd.to_datetime(file.date_created),
-            "url": file._download_url,
-            "size": file.size,
-            "file": file,
-        }
-        for file in storage.files
-    ]
-
-    if after_date is not None:
-        date = pd.to_datetime(after_date, format="%d/%m/%Y", utc=True)
-        files = [f for f, d in zip(files, [f["date"] > date for f in files]) if d]
-    return files
-
-
-token = ""  # Paste OSF token here to access private repositories
-files = osf_listfiles(
-    token=token,
-    data_subproject="au695",  # Data subproject ID
-    after_date="18/12/2023",
-)
+# List files ============================================================
+path = "C:/Users/domma/Box/Data/PHQ4R/"
+files = os.listdir(path)
 
 
 # Loop through files ======================================================
@@ -46,15 +18,11 @@ for i, file in enumerate(files):
 
     if (
         "Participant" in alldata.columns
-        and file["name"] in alldata["Participant"].values
+        and file.replace(".csv", "") in alldata["Participant"].values
     ):
         continue
 
-    download_ok = False
-    while download_ok == False:
-        data = pd.read_csv(file["file"]._get(file["url"], stream=True).raw)
-        if len(data) > 0:
-            download_ok = True
+    data = pd.read_csv(path + file)
 
     # Participant ========================================================
     # data["screen"].unique()
@@ -78,10 +46,9 @@ for i, file in enumerate(files):
 
     df = pd.DataFrame(
         {
-            "Participant": file["name"],
+            "Participant": file.replace(".csv", ""),
             "Experimenter": experimenter,
[…]

### 2024-12-07 paper/4_WIP/manuscript.qmd
paper

---
title: "**Measuring Depression and Anxiety with 4 items? Adaptation of the PHQ-4 to increase its Sensitivity to Subclinical Variability**"
shorttitle: "Short Title in Running Header"
author:
  - name: Dominique Makowski
    corresponding: true
    orcid: 0000-0001-5375-9967
    email: D.Makowski@sussex.ac.uk
    url: https://realitybending.github.io/
    roles:
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Funding acquisition
      - Investigation
      - Methodology
      - Project administration
      - Resources
      - Software
      - Supervision
      - Validation
      - Visualization
      - Writing – original draft
    affiliations:
      - id: "id1"
        name: "University of Sussex"
        department: "School of Psychology"
        city: "Brighton"
        country: "UK"
        postal-code: "BN1 9RH"
  - name: An Shu Te
    #corresponding: 
    orcid: 0000-0002-9312-5552
    role:
      - Project administration
      - Resources
      - Writing – original draft
    affiliations:
      - id: "id2"
        name: "Nanyang Technological University"
        department: "School of Social Sciences"
  - name: Ana Neves  
    orcid: 0009-0006-0020-7599
    role:
      - Data curation
      - Formal Analysis
      - Writing – review & editing
    affiliations:
      - ref: "id1"
  - name: S.H. Annabel Chen  # AnnabelChen@ntu.edu.sg
    orcid: 0000-0002-1540-5516
    role:
      - Project administration
      - Supervision
      - Writing – review & editing
    affiliations:
      - ref: "id2"
      - name: "Nanyang Technological University"
        department: "LKC Medicine"
      - name: "National Institute of Education"
      - name: "Nanyang Technological University"
        department:  "Centre for Research and Development in Learning"
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    gratitude: |
      ::: {.callout-note icon=false appearance="simple"}
      This preprint is a non-peer-reviewed work from the [**Reality Bending Lab**](https://realitybending.github.io/).
      ![](https://realitybending.github.io/media/ReBeL_LogoOnly_hu11484441381606756729.png){width=20% fig-align="center"}
      :::
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    authorship-agreements: ~
abstract: |
  The PHQ-4 is an ultra-brief (4 items) screening questionnaire for depression and anxiety. In this brief report, we test the benefits of adding one additional response option ("Once or twice", in between "Not at all" and "Several days") to improve the scale's sensitivity to milder alterations, and thus increase its usefulness in subclinical populations. In study 1 (n=485), we provide evidence using Item Response Theory (IRT) that the new response option does improve the scale's psychometric quality and extends the sensitivity to the measured constructs on the lower end of the spectrum. In study 2 (n=836), we show that the refined version offers an improved sensitivity to subclinical variability in depression (indexed by the BDI-II) as compared to the original version. In conclusion, adding the "once or twice" response option is a low-cost no-downsides way of increasing the PHQ-4's sensitivity to subclinical variability, making it a tool of choice for general population research.
#   **Significance Statement**. Developing reliable and sensitive instruments for mood disorders screening is critical in a global context marked by international crises (pandemics, wars), where increasingly more surveys are done online. In this study, we show that a small modification to the widely used PHQ-4 scale (adding the "Once or twice" response option) can increase its ability to capture the mild fluctuations prevalent in subclinical samples.
keywords: "PHQ-4, depression, anxiety, brief questionnaire validation, ultra short scale"
numbered-lines: true
floatsintext: true
bibliography: bibliography.bib
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: jou
---

The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement of core signs of depression and anxiety [@kroenke2009ultra]. It consists of two items for depression [PHQ–2, @kroenke2003patient] and anxiety [GAD–2, @kroenke2007anxiety], each corresponding to DSM-5 diagnostic symptoms for major depressive disorder (MDD) and generalized anxiety disorder (GAD). It has been validated across many languages and populations [@christodoulaki2022validation; @mendoza2022factor; @materu2020psychometric], becoming one of the most popular screening instruments for depression and anxiety [@maurer2018depression].

While the scale has been validated and used in the general population and non-clinical samples [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients. This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a focus on the sensitivity to milder mood alterations. In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and possibly leaves out the possibility of more subtle occurrences. While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantification.

This brief report aims at testing the possibility of enhancing - with minimal changes to the original scale - the PHQ-4 sensitivity to mild mood level inflections. In the first study, we will evaluate whether the new response option is prevalently used by participants, and whether it does capture a specific part of the latent measurand. In the second study, we will compare the refined PHQ-4 version to the original one in terms of sensitivity to subclinical variability in depression, using the Beck Depression Inventory [BDI-II, @beck1996beck] and the State Trait Anxiety Inventory [STAI-5, @zsido2020development] as our ground-truth measures of depression and anxiety.

# Study 1

## Method

### Participants

The sample consists of 485 English-speaking participants (Mean age = 30.1 +- 10.1 [18, 73]; 50.3\% females) from the general population recruited via *Prolific*, a crowd-sourcing platform recognized for providing high quality data [@peer2022data]. The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood. 
This study was approved by the 
<!-- Institutional Review Board **[Masked for Review]**. -->
NTU Institutional Review Board (NTU IRB-2022-187).
All participants provided their informed consent prior to participation and were incentivized after completing the study.

### Measures

In the original PHQ-4, the instructions *"Over the last 2 weeks, how often have you been bothered by the following problems?"* are followed with 4 items (A1 - *Feeling nervous, anxious or on edge*; A2 - *Not being able to stop or control worrying*; D1 - *Little interest or pleasure in doing things*; D2 - *Feeling down, depressed, or hopeless*). The original answer options are "Not at all" (0), "Several days" (1), "More than half the days" (2), "Nearly every day" (3). The total score is computed by summing the responses of each facet resulting in a 0-6 score for depression and anxiety.

For the refined version, we added a "Once or twice" option between "Not at all" and "Several days"in order to better capture potential mild mood inflections [see @dobson1979equidistant for the choice of the label]. This new option was scored as 0.5 to preserve the same scoring as the original version.

### Procedure

Participants were administered the refined PHQ-4 online as part of another study, which contained additional questionnaires and tasks not relevant fort the current analysis. 
The PHQ-4 was presented in a randomized order with other questionnaires.
The data is available in open-access at https://github.com/RealityBending/IllusionGameReliability.

## Results

The analysis was carried out using *R 4.4* [@RCoreTeam], the *tidyverse* [@wickham2019welcome], and the *easystats* collection of packages [@ludecke2019insight; @ludecke2021performance;
@ludecke2020extracting; @patil2022datawizard]. 
All reproducible scripts and complimentary analyses are available open-access at https://github.com/DominiqueMakowski/PHQ4R

```{r}
#| label: fig-one
#| fig-cap: "A) Proportion of answers of each type to the four items. B) Prevalence of answer pairs. C) Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels)."
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../../study1/figures/figure1.png")
```

### Descriptive Statistics

The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales is excellent. The proportion of response types stratified by item (see @fig-one) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in 29.12\% of cases).

### Item Response Theory

Item Response Theory (IRT) provides insights into how well items and responses capture an underlying latent trait $\theta$. For each of the subscales, we fitted a unidimensional graded response model [GRM, @samejima1997graded]. For anxiety, the two items captured 89.2\% of the variance of the latent anxiety dimension ($\theta_{anxiety}$). The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of $\theta_{anxiety}$, as can be seen in **Figure 1B**). The two depression items captured 82.8\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the first ($\alpha = 2.41$). However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive especially on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision. Importantly for our objective, the added "Once or twice" option did cover a selective and unique portion of the latent space.

## Discussion

The fact that the new "Once or twice" response option was the most prevalent response speaks to its usefulness in capturing more accurately participants' expression. The IRT analysis further revealed that this response tracks with precision a unique portion of the variability in the latent factors measured by the instrument. Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).

# Study 2

## Method

### Participants

The initial sample consisted of 1053 participants, recruited (181 were recruited on *Prolific*, 772 students from the University of Sussex via *SONA*, and the rest through convenience sampling as part of dissertation students' data collection).
We used attention checks as the primary target for participant exclusion. 
We excluded 194 participants (18.42\%) for failing at least one attention check, and 23 (2.18\%) that were outliers ($|z_{robust}| > 2.58$) on measures significantly related to the probability of failing attention checks [namely, the standard deviation of all the items of the IAS, as well as the the multivariate distance obtained with the OPTICS algorithm, see @theriault2024check].
The experiment duration was not related to the probability of failing attention checks and was thus not used as an exclusion criterion.

The final sample included 836 participants (Mean age = 25.1 +- 11.3 [18, 76]; 73.8\% women).
This study was approved by the University of Sussex' Ethics Committee **(ER/...)**.

In this sample, 51 participants (6.10%) were labelled as having Depression, as indexed by the self-reported presence of MDD together with the use of a treatment (antidepressent, anxiolytic and/or therapy), and 87 participants (10.41%) were labelled as having Anxiety, as indexed by the self-reported presence of GAD or Panic Disorder, also together with the use of a treatment.

### Measures

**TODO** (describe BDI-II, STAI-5, PHQ-4, and other measures used but not relevant)

The total scores for the original version are 0-6 scores for depression and anxiety. The refined version creates more possible scores (0.5, 1.5, 2.5, etc.) due to the presence of an item scored 0.5.

### Procedure

**TODO**

## Results

As all the scripts, analysis details and results tables are available open-access at https://github.com/DominiqueMakowski/PHQ4R, we will focus on reporting the main results.

```{r}
#| label: fig-two
#| fig-cap: "A) PHQ-4 depression and anxiety scores against their respective ground-truth measures, the BDI-22 and the STAI-5. Bayes factors in grey tell if there is a difference, for the same PHQ-4 score, between the original and the refined version (BFs < 1 suggest no difference and thus evidence for a comparability of the refined version with respect fo the original scale. Bayes factors in yellow represent how new in-between scores (0.5, 1.5, 2.5, ...) available with refined version differ from the adjacent scores (BFs > 3 suggest that half a point of difference on the refined PHQ-4 relates to a significant difference on the ground truth measure). BF < 1/3°, BF > 3\\*, BF > 10\\*\\*, BF > 30\\*\\*\\*. B) Bootstrapped distributions of the difference of correlation between the revised PHQ-4 scores and the original one for sub-clinical threshold scores of depression and anxiety. Positive differences suggest that the correlation between the ground-truth measure and the refined PHQ-4 score was stronger compared to the original version. C) Predictive power of the PHQ-4 scores on the presence of a depression or anxiety disorder. The upper plots show the relationship modelled by a logistic regression, while the above plots represent the ROC curves (in which a line further away from the diagonal represents a higher combination of sensitivity and specificity)."
#| apa-twocolumn: true  # A Figure Spanning Two Columns When in Journal Mode
#| out-width: "100%"

knitr::include_graphics("../../study2/analysis/2_analysis_files/figure-html/figure-1.png")
```

### PHQ-4 Depression *vs.* BDI-II

The linear regression predicting the BDI-II total score with the PHQ-4 depression score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} = TODO; \Delta \beta_{\text{refined}} = TODO$), suggesting no differences in the relationship pattern between the two versions (see @fig-two).

Moreover, Bayesian *t*-tests comparing the BDI-II scores between the refined and the original version at each integer score (0, 1, 2, 3) yielded no evidence in favour of a significant difference (BF > 3). In other words, having the same score on the refined version as on the original version was related to the same outcome on the BDI-II.

However, the low in-between scores from the refined version are overall capturing significantly different levels of depression compared to the adjacent scores. Scoring 0.5 was associated with a higher BDI-II score than scoring 0 (BF > 30), and lower scores than scoring 1 (BF > 30). Similarly, scoring 1.5 was associated with a higher BDI-II score than scoring 1 (BF > 30), but not lower scores than scoring 2 (BF = 0.234).

### PHQ-4 Anxiety *vs.* STAI-5

The linear regression predicting the STAI-5 general score with the PHQ-4 anxiety score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} = TODO; \Delta \beta_{\text{refined}} = TODO$), suggesting no differences in the relationship pattern between the two versions.

Moreover, Bayesian *t*-tests comparing the STAI-5 scores between the refined and the original version at each integer score yielded no evidence in favour of a significant difference. In other words, having the same score on the refined version as on the original version was related to the same outcome on the STAI-5.

However, comparing in-between scores with adjacent scores yielded mixed results. Scoring 0.5 on the PHQ-4 anxiety was not significantly associated with a different level of STAI-5 compared to scoring 0 (BF = 1.83), but was with scores of 1 (BF > 30). Similarly, there was no evidence that scoring 1.5 was different from scoring 1 (BF = 0.605), but strong evidence that it was different from scoring 2 (BF > 30).

### Correlation Differences

While the relationship pattern (i.e., the slope of the linear relationship) was not affected by the PHQ-4 version, we focused next on testing the difference in the strength (i.e., the precision) of the relationship, in particular at the lower end of the spectrum (i.e., for sub-clinical threshold scores of the BDI-II and STAI-5). We bootstrapped (2000 iterations) the difference in correlation between the refined and the original version for each of the two ground-truth measures, separately for the BDI-II subsamples (minimal to mild <= 18; moderate to severe > 18) and the STAI-5 subsamples (minimal to mild < 2; moderate to severe >= 2).

The results suggested that in the subclinical range of the BDI-II, the correlation between its score and the PHQ-4 Depression score was marginally higher (although not significantly, $p_{one-sided} = 0.161$) for the refined version compared to the original one. No correlation differences were observed in the moderate to severe range of the BDI-II.

For the STAI-5, there was no difference in the correlation between the refined and the original version in the subclinical range of the STAI-5. Surprisingly, we observed a stronger correlation between the refined PHQ-4 Anxiety score and the STAI-5 in the moderate to severe range compared to the original version ($p_{one-sided} = 0.027$).

### Predictive Power

Finally, we tested the predictive power of the PHQ-4 depression and anxiety scores on the presence of a depression or anxiety disorder, respectively. We modeled the relationship with a logistic regression. While the PHQ-4 was overall a strong predictor of the outcome, there was no significant difference between the two PHQ-4 versions.

However, the ROC curves for the refined and the original version of the PHQ-4, suggested that the refined version had a better sensitivity / specificity trade-off (AUC = 78.36%) compared to the original version (AUC=75%), in particular on the lower end of the spectrum. The difference was negligible for anxiety.

## Discussion

These results suggest that the new "Once or twice" response option to the PHQ-4 does help capturing more fine-grained variations of depressive symptoms, particularly in the subclinical range. Importantly, adding this new response option with the scoring of 0.5 does not disrupt the quality of the scale, which scores remain comparable to that of the original version.

The results for the anxiety subscale appear more mixed, with less evident benefits. However, this might have been partly caused by our design decision regarding the questionnaire used for the ground-truth measure of anxiety. Indeed, we used the abridged version of the STAI, which only included 5 items, arguably limiting the sensitivity of the anxiety measure in the first place.

Finally, although we used a stricter criterion for classifying participants as having a depression or an anxiety disorder by restricting it to participants also reporting undergoing a medical treatment, it was still based on self-reported data. Studies in controlled clinical settings are needed to confirm the potential benefits of the refined PHQ-4 in mood disorders detection accuracy.

# General Discussion

The objective of this study was to test the introduction of a "Once or twice" response option to the PHQ-4 to enhance its sensitivity to milder mood fluctuations. In the first study, we showed that the new response option was used prevalently by participants and did capture a unique portion of the depression and anxiety underlying dimensions. In the second study, we showed that the refined version of the PHQ-4 was able to better differentiate lower levels of depression compared to the original version, while remaining comparable. Although the benefits of this refinement appear to be fairly minor, and particularly marked for the depression score compared to anxiety, this cost-free improvement appear useful to implement when measuring depression and anxiety using the PHQ-4 ultra-short screening questionnaire.

# Acknowledgements

We would like to thank the dissertation students from the University of Sussex for their help in data collection.

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

### 2024-12-07 study2/analysis/1_cleaning.qmd
paper

@@
     code-fold: true
-    self-contained: true
+    self-contained: false
 ---
@@
 
-We removed `r length(outliers$attentionchecks)` participants for failing the attention checks.
+We removed `r length(outliers$attentionchecks)` (`r insight::format_percent(length(outliers$attentionchecks) / nrow(df))`) participants for failing the attention checks.
 
@@
 
-### Response Coherence
+### Response Coherence and Variability
 
@@
   rowMeans()
+df$IAS_SD <- select(df, matches("IAS_"), -IAS_Duration, -IAS_Order, -IAS_Total) |> 
+  SD_per_dim(dims=c("IAS_")) |> 
+  rowMeans()
 ```
@@
 ```{r}
-#| eval: false
+m <- mgcv::gam(Failed ~ s(IAS_SD), 
+          data=mutate(df, Failed = ifelse(AttentionCheck == "Failed", 1, 0)), 
+          family="binomial") 
+parameters::parameters(m)
 
-mgcv::gam(Failed ~ s(STAI5_SD), 
-          data=mutate(df, Failed = ifelse(AttentionCheck == "Failed", 1, 0)), 
-          family="binomial") |>
-  parameters::parameters()
+estimate_relation(m, length=40) |> 
+  ggplot(aes(x=IAS_SD, y=Predicted)) +
+  geom_ribbon(aes(ymin=CI_low, ymax=CI_high), alpha=0.2) +
+  geom_line() + 
+  ggside::geom_xsidedensity(data=df, aes(fill=AttentionCheck), alpha=0.3, color=NA)
 
 df |> 
-  select(Participant, STAI5_SD, AttentionCheck) |> 
-  mutate(Participant = fct_reorder(Participant, STAI5_SD)) |>
-  ggplot(aes(x=Participant, y=STAI5_SD, fill=AttentionCheck)) +
+  select(Participant, IAS_SD, AttentionCheck) |> 
+  mutate(Participant = fct_reorder(Participant, IAS_SD)) |>
+  ggplot(aes(x=Participant, y=IAS_SD, fill=AttentionCheck)) +
   geom_bar(stat="identity") +
@@
   theme(axis.text.x = element_blank())
+
+outliers$ias_sd <- as.character(df[abs(standardize(df$IAS_SD, robust=TRUE)) > qnorm(0.995), "Participant"])
+outliers$ias_sd <- outliers$ias_sd[!outliers$ias_sd %in% c(outliers$attentionchecks, outliers$duration)]
 ```
+
+We removed `r length(outliers$ias_sd)` (`r insight::format_percent(length(outliers$ias_sd) / nrow(df))`) participants for failing being an outlier on the IAS' SD value.
 
@@
 dfoutlier <- df |> 
-  select(contains("IAS_"), -IAS_Duration, -IAS_Order, -IAS_Total) |>
+  select(contains("IAS_"), -IAS_Duration, -IAS_Order, -IAS_Total, -IAS_SD) |>
   # select(contains("BDI2_"), -BDI2_Duration, -BDI2_Order, -BDI2_Total, -BDI2_SD) |>
+  # select(contains("STAI5_"), -STAI5_Duration, -STAI5_Order, -STAI5_General, -STAI5_SD) |>
+  # select(contains("MAIA2_"), -MAIA_Duration, -MAIA_Order) |>
   # datawizard::remove_empty_rows() |> 
@@
     method=c("optics", "mahalanobis_robust"),
-    threshold=list(optics=2.5, optics_xi=0.03)) |> 
+    threshold=list(optics=2, optics_xi=0.1)) |> 
   as.data.frame() |> 
@@
          Outlier_AttentionCheck = ifelse(Participant %in% outliers$attentionchecks, 1, 0),
-         Outlier_Distance = ifelse(abs(standardize(Distance_OPTICS)) > qnorm(0.995), 1, 0),
+         Outlier_Distance = ifelse(abs(standardize(Distance_OPTICS, robust=TRUE)) > qnorm(0.995), 1, 0),
          Experiment_Duration = df$Experiment_Duration,
@@
 outliers$distance <- as.character(dfoutlier[dfoutlier$Outlier_Distance == 1, "Participant"])
-outliers$distance <- outliers$distance[!outliers$distance %in% c(outliers$attentionchecks, outliers$duration)]
+outliers$distance <- outliers$distance[!outliers$distance %in% c(outliers$attentionchecks, outliers$duration, outliers$ias_sd)]
 
@@
 
+We removed `r length(outliers$distance)` (`r insight::format_percent(length(outliers$distance) / nrow(df))`) participants based on their multivariate distance.
+`r length(c(outliers$ias_sd, outliers$distance))` (`r insight::format_percent(length(c(outliers$ias_sd, outliers$distance)) / nrow(df))`) outliers in total.
+
 ```{r}
-df <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$distance))
+df <- filter(df, !Participant %in% c(outliers$attentionchecks, outliers$ias_sd, outliers$distance))
 df <- select(df, -starts_with("AttentionCheck"))
@@
   labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme_minimal()
 
@@
   labs(title = "Treatment", y = "Number of Participants", subtitle = "Participants per treatment", x="Type") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme_minimal()
 
@@
 
-df$Depression <- rowSums(select(df, DisorderTreatment_MoodStabilizer, DisorderTreatment_Antidepressant))
-df$Depression <- ifelse(df$Depression > 0, TRUE, FALSE)
+df$Treatment <- rowSums(select(df, DisorderTreatment_MoodStabilizer, DisorderTreatment_Antidepressant, DisorderTreatment_Anxiolytic, DisorderTreatment_Therapy))
+df$Treatment <- ifelse(df$Treatment > 0, TRUE, FALSE)
+
+df$Depression <- rowSums(select(df, Disorder_MDD, Disorder_Dysthymia))
+df$Depression <- ifelse(df$Depression > 0 & df$Treatment, TRUE, FALSE)
+
+df$Anxiety <- rowSums(select(df, Disorder_GAD, Disorder_Panic))
+df$Anxiety <- ifelse(df$Anxiety > 0 & df$Treatment, TRUE, FALSE)
 
@@
 
-`r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment.
-`r sum(df$Depression)` participants (`r insight::format_percent(sum(df$Depression) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer or antidepressant.
-`r sum(df$DisorderHistory, na.rm=TRUE)` participants (`r insight::format_percent(sum(df$DisorderHistory, na.rm=TRUE) / nrow(df))`) reported having a history of mood disorder.
+- `r sum(df$MoodDisorder)` participants (`r insight::format_percent(sum(df$MoodDisorder) / nrow(df))`) have a mood disorder, as indexed by the use of a mood stabilizer, antidepressant, or anxiolytic treatment.
+- `r sum(df$Depression)` participants (`r insight::format_percent(sum(df$Depression) / nrow(df))`) were labelled as having Depression, as indexed by the self-reported presence of MDD together with the use of a treatment (antidepressent, anxiolytic and/or therapy).
+- `r sum(df$Anxiety)` participants (`r insight::format_percent(sum(df$Anxiety) / nrow(df))`) were labelled as having Anxiety, as indexed by the self-reported presence of GAD or Panic Disorder together with the use of a treatment (antidepressent, anxiolytic and/or therapy).
+- `r sum(df$DisorderHistory, na.rm=TRUE)` participants (`r insight::format_percent(sum(df$DisorderHistory, na.rm=TRUE) / nrow(df))`) reported having a history of mood disorder.
 

### 2024-12-07 study2/analysis/2_analysis.qmd
paper

@@
 library(ggside)
+library(ggdist)
 ```
@@
     PHQ4_Condition = fct_recode(PHQ4_Condition, "Original" = "PHQ4 - Original", "Revised" = "PHQ4 - Revised"),
-    BDI2_Category = ifelse(BDI2_Total <= 19, "Mild", "Moderate+"),
-    STAI5_Category = ifelse(STAI5_General < 2, "Mild", "Moderate+"),
+    # BDI2_Category = ifelse(BDI2_Total <= 12, "Subclinical", ifelse(BDI2_Total <= 18, "Mild", "Moderate+")),
+    BDI2_Category = ifelse(BDI2_Total <= 18, "Minimal - Mild", "Moderate - Severe"),
+    STAI5_Category = ifelse(STAI5_General < 2, "Minimal - Mild", "Moderate - Severe"),
+    PHQ4_Depression_Fac = as.factor(PHQ4_Depression)
   )
+
+colors <- c("Original" = "#607D8B", "Revised" = "#FF9800")
 ```
@@
 
-### Whole Sample
+### Relationship with BDI-2
 
@@
 ```{r}
-df |> 
+get_bf <- function(resp="BDI2_Total", pred="PHQ4_Depression", levels=c(0, 0.5)) {
+  if(length(levels) == 2) {
+    bf <- BayesFactor::ttestBF(
+      df[df[[pred]] == levels[1] & df$PHQ4_Condition=="Revised", resp], 
+      df[df[[pred]] == levels[2] & df$PHQ4_Condition=="Revised", resp])
+  } else {
+    bf <- BayesFactor::ttestBF(
+      df[df[[pred]]==levels & df$PHQ4_Condition=="Original", resp],
+      df[df[[pred]]==levels & df$PHQ4_Condition=="Revised", resp])
+  }
+  insight::format_bf(exp(bf@bayesFactor$bf), stars=TRUE)
+}
+
+get_databf <- function(resp="BDI2_Total", pred="PHQ4_Depression", y1, y2) {
+
+  # Between scores comparisons
+  bfs <- data.frame(
+    BF = c(get_bf(resp, pred, c(0, 0.5)), 
+           get_bf(resp, pred, c(0.5, 1)), 
+           get_bf(resp, pred, c(1, 1.5)), 
+           get_bf(resp, pred, c(1.5, 2)),
+           get_bf(resp, pred, c(2, 2.5)),
+           get_bf(resp, pred, c(2.5, 3)))
+  )
+  bfs[[pred]] <- c(0.25, 0.75, 1.25, 1.75, 2.25, 2.75)
+  bfs[[resp]] <- y1
+  bfs$color <- "Revised"
+
+  # Within scores comparisons
+  bfs2 <- data.frame(
+    BF = c(get_bf(resp, pred, 0), 
+           get_bf(resp, pred, 1), 
+           get_bf(resp, pred, 2),
+           get_bf(resp, pred, 3),
+           get_bf(resp, pred, 4))
+  )
+  bfs2[[pred]] <- c(0, 1, 2, 3, 4)
+  bfs2[[resp]] <- y2
+  bfs2$color <- "Original"
+
+  bf <- rbind(bfs, bfs2)
+
+  # Lines behing labels
+  lines <- data.frame(
+    xmin = c(0, 0.5, 1, 1.5, 2, 2.5),
+    xmax = c(0.5, 1, 1.5, 2, 2.5, 3),
+    y = y1
+  )
+
+  list(bf=bf, lines=lines)
+}
+
+bfs <- get_databf(resp="BDI2_Total", pred="PHQ4_Depression", y1=c(28, 32, 36, 40, 44, 48), y2=c(10, 15, 20, 25, 30))
+
+p_dep <- df |> 
   ggplot(aes(x = PHQ4_Depression, y = BDI2_Total), width=0.1) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
+  ggdist::stat_slab(
+    data = df[df$PHQ4_Condition == "Original" & df$PHQ4_Depression <= 4, ], 
+    fill = colors["Original"],
+    alpha=0.3,
+    scale = 0.3,
+    position = position_dodge(width = 1),
+    side = "left") +
+  ggdist::stat_slab(
+    data = df[df$PHQ4_Condition == "Revised" & df$PHQ4_Depression %in% c(0, 1, 2, 3, 4), ], 
+    fill = colors["Revised"],
+    alpha=0.3,
+    scale = 0.3,
+    position = position_dodge(width = 1),
+    side = "right") +
+  ggdist::stat_slab(
+    data = df[df$PHQ4_Depression %in% c(0.5, 1.5, 2.5), ], 
+    fill = colors["Revised"],
+    alpha=0.3,
+    scale = 0.3,
+    position = position_dodge(width = 1),
+    side = "both") +
+  geom_jitter2(aes(color = PHQ4_Condition), width=0.1, alpha=0.5, size=3) +
+  geom_smooth(aes(color = PHQ4_Condition, fill=PHQ4_Condition), 
+              method = "lm", formula="y~x", alpha=0.2, key_glyph = draw_key_rect) +
+  geom_segment(data=bfs$lines, aes(x=xmin, xend=xmax, y=y, yend=y), color=colors["Revised"], linewidth=0.5) +
+  geom_label(data = bfs$bf, aes(x = PHQ4_Depression, y = BDI2_Total, label = BF, color=color), size=3) +
+  scale_fill_manual(values = colors) +
+  scale_color_manual(values = colors) +
   theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
+  # ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
   ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
+  # ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
   ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
+  ggside::theme_ggside_void() +
+  coord_cartesian(xlim=c(-0.15, 6.1)) +
+  guides(color="none", fill=guide_legend(override.aes = list(alpha=1))) +
+  labs(
+    x = "PHQ-4 (Depression Score)",
+    y = "BDI-II Total",
+    color = "PHQ4 Version",
+    fill = "PHQ4 Questionnaire Version",
+    title="Depression",
+  ) +
+  theme(legend.position = "top",
+        plot.title = element_text(face="bold"))
+p_dep
 ```
 
-### Subsamples
+### Predictive Performance
 
 ```{r}
-model <- lm(BDI2_Total ~ BDI2_Category / (PHQ4_Depression * PHQ4_Condition), data = df)
+m <- glm(Depression ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial")
 
-parameters::parameters(model) |> 
-  display()
+display(parameters::parameters(m))
 
-# model <- rstanarm::stan_glm(BDI2_Total ~ BDI2_Category / (PHQ4_Depression * PHQ4_Condition), data = df)
-# parameters::parameters(model)
-# bayestestR::bayesfactor_parameters(model)
+p_mdd <- m |> 
+  estimate_relation(length=20) |> 
+  ggplot(aes(x=PHQ4_Depression, y=Predicted)) +
+  geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=PHQ4_Condition), alpha=0.2) +
+  geom_line(aes(color=PHQ4_Condition), linewidth=1) +
+  scale_fill_manual(values = colors) +
+  scale_color_manual(values = colors) +
+  scale_y_continuous(labels=scales::percent) +
+  labs(y = "Probability of Depression",
+       x= "PHQ-4 (Depression Score)",
+       title = "Depression Disorder",
+       fill = "Questionnaire Version",
+       color = "Questionnaire Version") +
+  theme_minimal()  +
+  theme(plot.title = element_text(face="bold"),
+        legend.position = "none")
 ```
@@
 ```{r}
-df |> 
-  ggplot(aes(x = PHQ4_Depression, y = BDI2_Total)) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition, linetype=BDI2_Category), method = "lm") +
-  theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
-```
+make_roc <- function(m) {
+  resp <- insight::find_response(m)
+  pred <- insight::find_predictors(m)
+  f <- as.formula(paste0(resp, " ~ ", pred$conditional[1]))
+  m1 <- glm(f, data = df[df$PHQ4_Condition == "Original", ], family="binomial") 
+  m2 <- glm(f, data = df[df$PHQ4_Condition == "Revised", ], family="binomial")
 
-### General Score
+  roc1 <- performance_roc(m1) 
+  roc2 <- performance_roc(m2)
+  rocs <- data.frame(AUC = c(
+    bayestestR::area_under_curve(roc1$Specificity, roc1$Sensitivity),
+    bayestestR::area_under_curve(roc2$Specificity, roc2$Sensitivity)),
+    PHQ4_Condition = c("Original", "Revised"))
+  rocs$AUC <- paste0("AUC = ", insight::format_percent(rocs$AUC))
+  rocs$x <- c(0.25, 0.25)
+  rocs$y <- c(0.35, 0.85)
 
-```{r}
-lm(BDI2_Total ~ PHQ4_Total * PHQ4_Condition, data = df) |> 
-  parameters::parameters() |> 
-  display()
+  rbind(
+    mutate(as.data.frame(roc1), PHQ4_Condition = "Original"),
+    mutate(as.data.frame(roc2), PHQ4_Condition = "Revised")
+  ) |> 
+    ggplot(aes(x=Specificity)) +
+    geom_abline(intercept=0, slope=1, color="gray") +
+    geom_line(aes(y=Sensitivity, color=PHQ4_Condition), linewidth=2, alpha=0.9) +
+    geom_label(data=rocs, aes(x=x, y=y, label=AUC), color=colors) +
+    labs(x= "1 - Specificity (False Positive Rate)", y="Sensitivity (True Positive Rate)",
+         color="Questionnaire Version") +
+    scale_color_manual(values=colors) +
+    scale_x_continuous(labels=scales::percent) +
+    scale_y_continuous(labels=scales::percent) +
+    theme_minimal()
+}
 
-lm(BDI2_Total ~ BDI2_Category / (PHQ4_Total * PHQ4_Condition), data = df) |> 
-  parameters::parameters() |> 
-  display()
+p_mddroc <- make_roc(m)
 ```
@@
 ```{r}
-m0 <- lm(BDI2_Total ~ PHQ4_Depression, data = filter(df, PHQ4_Condition == "Original"))
-m1 <- lm(BDI2_Total ~ PHQ4_Depression, data = filter(df, PHQ4_Condition == "Revised"))
-
-performance::compare_performance(m0, m1)
+p_mdd / p_mddroc + plot_layout(guides="collect")
 ```
@@
 
-### Whole Sample
+### Relationship with STAI
 
@@
 ```{r}
-df |> 
-  ggplot(aes(x = PHQ4_Anxiety, y = STAI5_General)) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
+bfs <- get_databf(resp="STAI5_General", pred="PHQ4_Anxiety", y1=c(2.0, 2.2, 2.4, 2.6, 2.8, 3), y2=c(0.7, 1, 1.3, 1.6, 1.9))
+
+p_anx <- df |> 
+  ggplot(aes(x = PHQ4_Anxiety, y = STAI5_General), width=0.1) +
+  ggdist::stat_slab(
+    data = df[df$PHQ4_Condition == "Original" & df$PHQ4_Anxiety <= 4, ], 
+    fill = colors["Original"],
+    alpha=0.3,
+    scale = 0.3,
+    position = position_dodge(width = 1),
+    side = "left") +
+  ggdist::stat_slab(
+    data = df[df$PHQ4_Condition == "Revised" & df$PHQ4_Anxiety %in% c(0, 1, 2, 3, 4), ], 
+    fill = colors["Revised"],
+    alpha=0.3,
+    scale = 0.3,
+    position = position_dodge(width = 1),
+    side = "right") +
+  ggdist::stat_slab(
+    data = df[df$PHQ4_Anxiety %in% c(0.5, 1.5, 2.5), ], 
+    fill = colors["Revised"],
+    alpha=0.3,
+    scale = 0.3,
+    position = position_dodge(width = 1),
+    side = "both") +
+  geom_jitter2(aes(color = PHQ4_Condition), width=0.1, alpha=0.5, size=3) +
+  geom_smooth(aes(color = PHQ4_Condition, fill=PHQ4_Condition), 
+              method = "lm", formula="y~x", alpha=0.2, key_glyph = draw_key_rect) +
+  geom_segment(data=bfs$lines, aes(x=xmin, xend=xmax, y=y, yend=y), color=colors["Revised"], linewidth=0.5) +
+  geom_label(data = bfs$bf, aes(x = PHQ4_Anxiety, y = STAI5_General, label = BF, color=color), size=3) +
+  scale_fill_manual(values = colors) +
+  scale_color_manual(values = colors) +
   theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
+  # ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
   ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
+  # ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
   ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
+  ggside::theme_ggside_void() +
+  coord_cartesian(xlim=c(-0.15, 6.1)) +
+  guides(color="none", fill=guide_legend(override.aes = list(alpha=1))) +
+  labs(
+    x = "PHQ-4 (Anxiety Score)",
+    y = "STAI-5 General",
+    color = "PHQ4 Version",
+    fill = "PHQ4 Questionnaire Version",
+    title="Anxiety",
+  ) +
+  theme(legend.position = "top",
+        plot.title = element_text(face="bold"))
+p_anx
 ```
 
-### Subsamples
+### Predictive Performance
 
 ```{r}
-model <- lm(STAI5_General ~ STAI5_Category / (PHQ4_Anxiety * PHQ4_Condition), data = df)
+m <- glm(Anxiety ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial")
 
-parameters::parameters(model) |> 
-  display()
+display(parameters::parameters(m))
+
+p_gad <- m |> 
+  estimate_relation(length=20) |> 
+  ggplot(aes(x=PHQ4_Anxiety, y=Predicted)) +
+  geom_ribbon(aes(ymin=CI_low, ymax=CI_high, fill=PHQ4_Condition), alpha=0.2) +
+  geom_line(aes(color=PHQ4_Condition), linewidth=1) +
+  scale_fill_manual(values = colors) +
+  scale_color_manual(values = colors) +
+  scale_y_continuous(labels=scales::percent) +
+  labs(y = "Probability of Anxiety",
+       x= "PHQ-4 (Anxiety Score)",
+       title = "Anxiety Disorder",
+       fill = "Questionnaire Version",
+       color = "Questionnaire Version") +
+  theme_minimal()  +
+  theme(plot.title = element_text(face="bold"),
+        legend.position = "none")
 ```
@@
 ```{r}
-df |> 
-  ggplot(aes(x = PHQ4_Anxiety, y = STAI5_General)) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition, linetype=STAI5_Category), method = "lm") +
-  theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
-```
+p_gadroc <- make_roc(m)
 
-### General Score
-
-```{r}
-lm(STAI5_General ~ PHQ4_Total * PHQ4_Condition, data = df) |> 
-  parameters::parameters() |> 
-  display()
-
-lm(STAI5_General ~ BDI2_Category / (PHQ4_Total * PHQ4_Condition), data = df) |> 
-  parameters::parameters() |> 
-  display()
+(p_gad / p_gadroc) + plot_layout(guides="collect")
 ```
@@
 
-### Test
-
-```{r}
-dat <- df |> 
-  group_by(PHQ4_Condition) |>
-  summarize(
-    N = n(),
-    r_depression = cor.test(BDI2_Total, PHQ4_Depression)$estimate,
-    r_anxiety = cor.test(STAI5_General, PHQ4_Anxiety)$estimate,
-  )
-
-psych::r.test(n=dat$N[1], n2=dat$N[2], r12=dat$r_depression[1], r34=dat$r_depression[2])
-psych::r.test(n=dat$N[1], n2=dat$N[2], r12=dat$r_anxiety[1], r34=dat$r_anxiety[2])
-```
-
-### Bootstrapping
-
-#### Whole Sample
+### Whole Sample
 
@@
   dat <- data.frame()
-  for(i in 1:1000) {
+  for(i in 1:2000) {
     newdf_ori <- df_original[sample(nrow(df_original), replace = TRUE), ]
@@
   theme_bw() +
-  labs(x = "Difference in Correlation", y = "Density") 
+  labs(x = "Difference in Correlation", y = "") 
 ```
 
-#### Subsamples
+### Subsamples
 
@@
 dat <- data.frame()
-for(cat in c("Mild", "Moderate+")) {
+for(cat in c("Minimal - Mild", "Moderate - Severe")) {
   dat <- rbind(
@@
       filter(df, PHQ4_Condition == "Revised", BDI2_Category == cat)) |> 
-      mutate(Category = paste("Depression -", cat)),
+      mutate(Subsample = paste("Depression:", cat)),
     bootstrap_diffs(
@@
       filter(df, PHQ4_Condition == "Revised", STAI5_Category == cat)) |> 
-      mutate(Category = paste("Anxiety -", cat)) 
+      mutate(Subsample = paste("Anxiety:", cat)) 
   )
@@
 
-dat |> 
-  separate(Category, into = c("Type", "Category"), sep = " - ") |>
+dat <- dat |> 
+  separate(Subsample, into = c("Type", "Subsample"), sep = ":") |>
   pivot_longer(cols = c(d_r_depression, d_r_anxiety)) |>
+  filter(!(Type == "Depression" & name=="d_r_anxiety") & 
+           !(Type == "Anxiety" & name=="d_r_depression")) |> 
+  mutate(Type = ifelse(Type == "Depression", paste(Type, "(BDI-II)"), paste(Type, "(STAI-5)"))) |> 
+  mutate(Type = fct_rev(Type))
+
+p_cor <- dat |> 
   ggplot(aes(x = value)) +
-  geom_density(aes(fill = name), alpha = 0.5) +
+  geom_density(aes(fill = name), alpha = 0.9, color="white") +
   geom_vline(xintercept = 0, linetype = "dashed") +
+  geom_text(data=summarize(dat, pd = insight::format_p(bayestestR::pd_to_p(
+    as.numeric(pd(value)), direction="one-sided"), stars=TRUE), 
+                           .by=c("Subsample", "Type")),
+            aes(label=pd, x=0.3, y=5)) +
   theme_bw() +
-  labs(x = "Difference in Correlation", y = "Density") +
-  facet_grid(Category~Type)
+  labs(x = "Difference in Correlation (PHQ4 Revised - Original)", y = "") +
+  facet_grid(Type~Subsample)  +
+  scale_fill_manual(values = c("d_r_depression"="#3F51B5", "d_r_anxiety"="#E91E63"), guide="none") +
+  theme(axis.ticks.y = element_blank(), 
+        axis.text.y = element_blank(),
+        strip.background = element_rect(color=NA))
+p_cor
 ```
 
-## Predictive Value
+## Figures
 
-### Mood Disorder
+```{r figure}
+#| fig-width: 14
+#| fig-height: 14
 
-```{r}
-glm(MoodDisorder ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
+p_a <- (p_dep + theme(legend.position='none')) / 
+              (p_anx  + theme(legend.position='none')) +
+  plot_annotation(title="A.")
 
-glm(MoodDisorder ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
+p_b <- p_cor + plot_annotation(title="B.")
 
-glm(MoodDisorder ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
+p_c <- ((p_mdd / p_mddroc) | (p_gad / p_gadroc)) + patchwork::plot_layout(guides = "collect") + patchwork::plot_annotation(title="C.", theme=theme(legend.position='left'))
+
+(wrap_elements(p_a) | wrap_elements(p_b)) / wrap_elements(p_c)
 ```
-
-### MDD
-
-```{r}
-glm(Disorder_MDD ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
-
-glm(Disorder_MDD ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
-
-glm(Disorder_MDD ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
-```
-
-### GAD
-
-```{r}
-glm(Disorder_GAD ~ PHQ4_Total * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
-
-glm(Disorder_GAD ~ PHQ4_Depression * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
-
-glm(Disorder_GAD ~ PHQ4_Anxiety * PHQ4_Condition, data = df, family="binomial") |> 
-  parameters::parameters() |> 
-  display()
-```
-
-## Modified - Depression
-
-### Whole Sample
-
-```{r}
-model <- lm(BDI2_Total ~ Expe_PHQ4_Depression * PHQ4_Condition, data = df)
-
-parameters::parameters(model) |> 
-  display()
-```
-
-```{r}
-df |> 
-  ggplot(aes(x = Expe_PHQ4_Depression, y = BDI2_Total), width=0.1) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
-  theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
-```
-
-### Subsamples
-
-```{r}
-model <- lm(BDI2_Total ~ BDI2_Category / (Expe_PHQ4_Depression * PHQ4_Condition), data = df)
-
-parameters::parameters(model) |> 
-  display()
-```
-
-```{r}
-df |> 
-  ggplot(aes(x = Expe_PHQ4_Depression, y = BDI2_Total)) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition, linetype=BDI2_Category), method = "lm") +
-  theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
-```
-
-## Modified - Anxiety
-
-### Whole Sample
-
-```{r}
-model <- lm(STAI5_General ~ Expe_PHQ4_Anxiety * PHQ4_Condition, data = df)
-
-parameters::parameters(model) |> 
-  display()
-```
-
-```{r}
-df |> 
-  ggplot(aes(x = Expe_PHQ4_Anxiety, y = STAI5_General)) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition), method = "lm") +
-  theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
-```
-
-### Subsamples
-
-```{r}
-model <- lm(STAI5_General ~ STAI5_Category / (Expe_PHQ4_Anxiety * PHQ4_Condition), data = df)
-
-parameters::parameters(model) |> 
-  display()
-```
-
-```{r}
-df |> 
-  ggplot(aes(x = Expe_PHQ4_Anxiety, y = STAI5_General)) +
-  geom_jitter(aes(color = PHQ4_Condition), width=0.1) +
-  geom_smooth(aes(color = PHQ4_Condition, linetype=STAI5_Category), method = "lm") +
-  theme_bw() +
-  ggside::geom_xsidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
-  ggside::geom_ysidedensity(aes(fill = PHQ4_Condition), color=NA, alpha=0.5) +
-  ggside::scale_ysidex_continuous(expand = c(0, 0)) +
-  ggside::theme_ggside_void() 
-```

### 2024-12-07 study2/analysis/0_preprocessing.py

@@ -247,6 +247,8 @@ for i, file in enumerate(files):
         continue
     if df["SONA_ID"].values[0] == 32742 and df["Date"].values[0] != "03/12/2024":
         continue
+    if df["SONA_ID"].values[0] == 32624 and df["Experiment_Duration"].values[0] < 10:
+        continue
 
     # Demographics -------------------------------------------------------
     demo1 = data[data["screen"] == "demographics_1"].iloc[0]
@@ -324,6 +326,7 @@ for i, file in enumerate(files):
             "cuacasian",
             "portuguese",
             "white-british",
+            "white - irish",
         ]
         else race
     )
@@ -340,6 +343,7 @@ for i, file in enumerate(files):
             "south asian",
             "sri lankan",
             "indian-asian",
+            "indian, british",
         ]
         else race
     )
@@ -359,7 +363,7 @@ for i, file in enumerate(files):
             "british chinese",
             "asian british",
             "british asian",
-            "white - irish",
+            "burmese",
         ]
         else race
     )
@@ -1444,6 +1448,19 @@ sona_credited = [
     32587,
     32656,  # failed attention check 3
     32742,  # failed attention check 3
+    # 07/12/2024
+    31797,  # failed attention check 3
+    31904,
+    32284,  # failed attention check 3
+    32334,
+    32340,  # failed attention check 3
+    32496,
+    32575,  # failed attention check 3
+    32624,
+    32660,
+    32666,  # failed attention check 3
+    32671,
+    32730,
     # NOT IN LIST ----------
     # 32118,
     31918,

### 2024-12-10 paper/4_WIP/manuscript.qmd
add paragraph

@@
 In this brief report, we test the benefits of adding one additional response option ("Once or twice", in between "Not at all" and "Several days") to improve the scale's sensitivity to milder alterations, and thus increase its usefulness in subclinical populations.
-In study 1 (n=485), we provide evidence using Item Response Theory (IRT) that the new response option does improve the scale's psychometric quality and extends the sensitivity to the measured constructs on the lower end of the spectrum.
-In study 2 (n=836), we show that the refined version offers an improved sensitivity to subclinical variability in depression (indexed by the BDI-II) as compared to the original version.
+In study 1 (N=485), we provide evidence using Item Response Theory (IRT) that the new response option does improve the scale's psychometric quality and extends the sensitivity to the measured constructs on the lower end of the spectrum.
+In study 2 (N=836), we show that the refined version offers an improved sensitivity to subclinical variability in depression (indexed by the BDI-II) as compared to the original version.
 In conclusion, adding the "once or twice" response option is a low-cost no-downsides way of increasing the PHQ-4's sensitivity to subclinical variability, making it a tool of choice for general population research.
@@
 For anxiety, the two items captured 89.2\% of the variance of the latent anxiety dimension ($\theta_{anxiety}$).
-The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of $\theta_{anxiety}$, as can be seen in **Figure 1B**).
+The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of $\theta_{anxiety}$, as can be seen in @fig-one).
 The two depression items captured 82.8\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the first ($\alpha = 2.41$).
@@
 Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).
+
+One natural methodological limitation pertains to the interpretation of the latent dimensions in an IRT framework applied to pairs of items.
+In this study's context, references to for instance "the latent anxiety dimension" merely corresponds to the amalgamation of the two items of the anxiety subscale, and not to a more general and valid true anxiety factor.
 

### 2024-12-10 paper/4_WIP/manuscript.qmd
minor todo

@@
 
-Moreover, Bayesian *t*-tests comparing the BDI-II scores between the refined and the original version at each integer score (0, 1, 2, 3) yielded no evidence in favour of a significant difference (BF > 3).
+Moreover, Bayesian *t*-tests (using *BayesFactor*'s `ttestBF()` function with default priors; **ADD REF TO PKG**) comparing the BDI-II scores between the refined and the original version at each integer score (0, 1, 2, 3) yielded no evidence in favour of a significant difference (BF > 3).
 In other words, having the same score on the refined version as on the original version was related to the same outcome on the BDI-II.

### 2024-12-10 paper/4_WIP/manuscript.qmd
added measures and procedure section

--- 2024-12-10 paper/4_WIP/manuscript.qmd
@@
 This brief report aims at testing the possibility of enhancing - with minimal changes to the original scale - the PHQ-4 sensitivity to mild mood level inflections.
-In the first study, we will evaluate whether the new response option is prevalently used by participants, and whether it does capture a specific part of the latent measurand.
+In the first study, we will evaluate whether the new response option is prevalently used by participants, and whether it does capture a specific part of the latent measure.
 In the second study, we will compare the refined PHQ-4 version to the original one in terms of sensitivity to subclinical variability in depression, using the Beck Depression Inventory [BDI-II, @beck1996beck] and the State Trait Anxiety Inventory [STAI-5, @zsido2020development] as our ground-truth measures of depression and anxiety.
@@
 
-**TODO** (describe BDI-II, STAI-5, PHQ-4, and other measures used but not relevant)
+<!--**TODO** (describe BDI-II, STAI-5, PHQ-4, and other measures used but not relevant) -->
 
-The total scores for the original version are 0-6 scores for depression and anxiety. The refined version creates more possible scores (0.5, 1.5, 2.5, etc.) due to the presence of an item scored 0.5.
+The BDI-II [@beck1996beck] was used to assess depressive symptoms.
+This self-report questionnaire consists of 21 items, each addressing a specific symptom of depression.
+Each item contains four response options scored from 0 to 3, reflecting the severity of the symptom.
+Participants are instructed to select the option that best describes how they have felt over the past two weeks.
+For example, the item assessing sadness includes the following response options:
+- 0: I do not feel sad.  
+- 1: I feel sad much of the time.  
+- 2: I am sad all the time.  
+- 3: I am so sad or unhappy that I can't stand it.  
+The total score is calculated by summing the scores for all 21 items, with higher scores indicating greater severity of depressive symptoms.
+
+The PHQ-4 [@kroenke2009ultra] is a brief measure of core symptoms of depression and anxiety, consisting of four items—two for depression and two for anxiety.
+Items are rated on a 4-point Likert scale from *Not at all* to *Nearly every day*, reflecting how participants felt over the past two week (e.g., *Feeling down, depressed, or hopeless"*).
+The total score, obtained by summing all item scores, is categorized as: normal (0–2), mild (3–5), moderate (6–8), and severe (9–12).
+A score ≥3 on the first two items indicates significant anxiety, while a score ≥3 on the last two items indicates significant depression.
+The total scores for the original version are 0-6 scores for depression and anxiety.
+The refined version creates more possible scores (0.5, 1.5, 2.5, etc.) due to the presence of an item scored 0.5.
+
+The STAIT-5 [@zsido2020development] is a brief measure of trait anxiety.
+It comprises 5 items rated on a 4-point Likert scale ranging from *Not at all* to *Very much so*.
+For example, one item states: *"I worry too much over something that really doesn't matter"*.
+Changes will be made in the instructions from asking *"how participants feel right now"* to *"over the past 2 weeks"* to keep it consistent with the instructions of the PHQ-4 and BDI-II Trait anxiety is assessed by summing the scores for all items, with higher total scores reflecting greater levels of trait anxiety.
+
+The Interoceptive Accuracy Scale (IAS; @murphy2020testing) was used to assess subjective interoceptive accuracy.
+The IAS consists of 21 items, each rated on an analogous scale from *Strongly Disagree* to *Strongly Agree*.
+Participants respond to statements about their ability to perceive bodily sensations (e.g., *"I can always accurately perceive when my heart is beating fast."*).
+The total score is calculated by summing the item responses, with higher scores indicating greater self-reported subjective interoceptive accuracy.
+
+The Multidimensional Assessment of Interoceptive Awareness - Version 2 (MAIA-2; @mehling2018multidimensional) was utilized to assess interoceptive awareness.
+The MAIA-2 comprises 37 items, each rated on a analogous scale ranging from *Never* to *Always.* An example of an item would be: *"I trust my body sensations"*.
+Interoceptive awareness is calculated by summing the item scores, with higher total scores indicating greater interoceptive awareness.
 
@@
 
-**TODO**
+<!--**TODO**-->
+
+Participants will first be presented with an information and consent form.
+They will be asked to provide consent by clicking the button labeled "I read, understood, and I consent." Following this, they will complete demographic questions, including gender, age, educational level, country of residence, and ethnicity.
+
+Participants will then be asked about relevant medical history.
+They will be explicitly informed that answering these questions is optional.
+The questions cover whether they are currently living with any psychopathologies (e.g., Major Depressive Disorder, Borderline Personality Disorder, Panic Disorder), whether they have been previously diagnosed with any of these conditions, and whether they are currently undergoing any treatments or interventions, such as mood stabilizers (e.g., LITHIUM, LAMICTAL) or mindfulness and stress management techniques.
+
+After completing or skipping this section, participants will be randomly assigned to complete either the original or modified version of the PHQ-4, followed by the BDI-II, STAIT-5, IAS, and MAIA-2, presented in random order.
+At the end of the experiment, participants will be provided with a debriefing page and given a final opportunity to withdraw their participation.
+The experiment, implemented using jsPsych [@de2015jspsych], will take approximately 10 minutes.
 

### 2024-12-11 paper/4_WIP/manuscript.qmd
Update manuscript.qmd
@AnafNeves

--- 2024-12-10 paper/4_WIP/manuscript.qmd
@@
 
-**TODO** (describe BDI-II, STAI-5, PHQ-4, and other measures used but not relevant)
+Participants were randomly assigned to complete either the original or refined version of the PHQ-4, which included one additional response option ("Once or twice") scored 0.5 (creating more possible total scores - 0.5, 1.5, 2.5, etc.).
 
-The total scores for the original version are 0-6 scores for depression and anxiety. The refined version creates more possible scores (0.5, 1.5, 2.5, etc.) due to the presence of an item scored 0.5.
+The BDI-II [@beck1996beck] was used as a ground truth measure of depressive symptoms.
+It includes 21 items, each addressing a specific depression symptom and offering four response options scored from 0 to 3.
+Participants are instructed to select the option that best describes how they have felt over the past two weeks.
+The total score is calculated by summing the scores for all 21 items, with higher scores indicating greater severity of depressive symptoms.
+
+The STAI-5 [@zsido2020development] was used as a ground truth measure of anxiety.
+This abridged version of the STAI (**TODO: add ref**) includes 5 items rated on a 4-point Likert scale.
+Changes were made in the instructions from asking *"how participants feel right now"* to *"over the past 2 weeks"* to keep it consistent with the instructions of the PHQ-4 and BDI-II.
+A general score of anxiety was computed by averaging all the items.
+
+Participants were also asked to complete two questionnaires of interoception, namely the Interoceptive Accuracy Scale [IAS - 21 items rated on analog scales, @murphy2020testing] and the Multidimensional Assessment of Interoceptive Awareness [MAIA-2 - 37 items, @mehling2018multidimensional].
+
+After demographic questions, participants were asked to report the current presence of psychiatric issues (from a list), as well as the usage of treatment (antidepressants, mood stabilizers, anxiolytics, therapy).
+We indexed the presence of a depression when participants reported suffering from either Major Depressive Disorder (MDD) or Dysthymia, as well as undergoing a medical treatment.
+Similarly, we indexed the presence of an anxiety disorder when participants reported suffering from either Generalized Anxiety Disorder (GAD) or Panic Disorder, as well as undergoing a medical treatment.
 
@@
 
-**TODO**
+The original or refined version of the PHQ-4 was followed by the BDI-II, STAIT-5, IAS, and MAIA-2, presented in random order.
+The IAS and the MAIA-2 were included as part of another study focused on interoception, and were only used as part in this study of participation quality control.
 

### 2024-12-12 paper/4_WIP/manuscript.qmd
updated manuscript

@@
   apaquarto-pdf:
-    documentmode: jou
+    documentmode: man
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 The final sample included 836 participants (Mean age = 25.1 +- 11.3 [18, 76]; 73.8\% women).
-This study was approved by the University of Sussex' Ethics Committee **(ER/...)**.
+This study was approved by the University of Sussex' Ethics Committee (ER/ASF25/4).
 
@@
 
-The STAI-5 [@zsido2020development] was used as a ground truth measure of anxiety.
-This abridged version of the STAI (**TODO: add ref**) includes 5 items rated on a 4-point Likert scale.
+The STAI-T5 [@zsido2020development] was used as a ground truth measure of anxiety.
+This abridged version of the STAI [@marteau1992development] includes 5 items rated on a 4-point Likert scale.
 Changes were made in the instructions from asking *"how participants feel right now"* to *"over the past 2 weeks"* to keep it consistent with the instructions of the PHQ-4 and BDI-II.
@@
 
-The linear regression predicting the BDI-II total score with the PHQ-4 depression score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} = TODO; \Delta \beta_{\text{refined}} = TODO$), suggesting no differences in the relationship pattern between the two versions (see @fig-two).
+The linear regression predicting the BDI-II total score with the PHQ-4 depression score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.13, 95\% CI[-1.73, 1.47], t(832) =  -0.16, p = 0.871 ; \Delta \beta_{\text{refined}} = -0.05, 95\% CI[-0.70, 0.60], t(832) = -0.15 , p = 0.883$), suggesting no differences in the relationship pattern between the two versions (see @fig-two).
 
-Moreover, Bayesian *t*-tests (using *BayesFactor*'s `ttestBF()` function with default priors; **ADD REF TO PKG**) comparing the BDI-II scores between the refined and the original version at each integer score (0, 1, 2, 3) yielded no evidence in favour of a significant difference (BF > 3).
+Moreover, Bayesian *t*-tests (using *BayesFactor*'s `ttestBF()` function with default priors; @moreyrouder2024) comparing the BDI-II scores between the refined and the original version at each integer score (0, 1, 2, 3) yielded no evidence in favour of a significant difference (BF > 3).
 In other words, having the same score on the refined version as on the original version was related to the same outcome on the BDI-II.
@@
 
-The linear regression predicting the STAI-5 general score with the PHQ-4 anxiety score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} = TODO; \Delta \beta_{\text{refined}} = TODO$), suggesting no differences in the relationship pattern between the two versions.
+The linear regression predicting the STAI-5 general score with the PHQ-4 anxiety score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.02, 95\% CI[-0.15, 0.11], t(832) =  -0.32, p = 0.750; \Delta \beta_{\text{refined}} = 0.01, 95\% CI[-0.03, 0.05], t(832) = 0.56 , p = 0.576$), suggesting no differences in the relationship pattern between the two versions.
 

### 2024-12-12 paper/4_WIP/manuscript.qmd
update

@@
         postal-code: "BN1 9RH"
+      - name: "University of Sussex"
+        department:  "Sussex Centre for Consciousness Science"
   - name: An Shu Te
@@
         department: "LKC Medicine"
-      - name: "National Institute of Education"
       - name: "Nanyang Technological University"
@@
   apaquarto-pdf:
-    documentmode: man
-editor: 
-  markdown: 
-    wrap: 72
+    documentmode: jou
 ---
@@
 
-The sample consists of 485 English-speaking participants (Mean age = 30.1 +- 10.1 [18, 73]; 50.3\% females) from the general population recruited via *Prolific*, a crowd-sourcing platform recognized for providing high quality data [@peer2022data].
+The sample consists of 485 English-speaking participants (Mean age = 30.1 $\pm$ 10.1 [18, 73]; 50.3\% females) from the general population recruited via *Prolific*, a crowd-sourcing platform recognized for providing high quality data [@peer2022data].
 The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood.
@@
 
-The final sample included 836 participants (Mean age = 25.1 +- 11.3 [18, 76]; 73.8\% women).
+The final sample included 836 participants (Mean age = 25.1 $\pm$ 11.3 [18, 76]; 73.8\% women).
 This study was approved by the University of Sussex' Ethics Committee (ER/ASF25/4).
@@
 
-The BDI-II [@beck1996beck] was used as a ground truth measure of depressive symptoms.
+Beck's Depression Inventory [BDI-II, @beck1996beck] was used as a ground truth measure of depressive symptoms.
 It includes 21 items, each addressing a specific depression symptom and offering four response options scored from 0 to 3.
@@
 
-The STAI-T5 [@zsido2020development] was used as a ground truth measure of anxiety.
+The short version of the State-Trait Anxiety Inventory [STAI-5, @zsido2020development] was used as a ground truth measure of anxiety.
 This abridged version of the STAI [@marteau1992development] includes 5 items rated on a 4-point Likert scale.
@@
 
-The original or refined version of the PHQ-4 was followed by the BDI-II, STAIT-5, IAS, and MAIA-2, presented in random order.
+The original or refined version of the PHQ-4 was followed by the BDI-II, STAI-5, IAS, and MAIA-2, presented in random order.
 The IAS and the MAIA-2 were included as part of another study focused on interoception, and were only used as part in this study of participation quality control.
@@
 
-The linear regression predicting the BDI-II total score with the PHQ-4 depression score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.13, 95\% CI[-1.73, 1.47], t(832) =  -0.16, p = 0.871 ; \Delta \beta_{\text{refined}} = -0.05, 95\% CI[-0.70, 0.60], t(832) = -0.15 , p = 0.883$), suggesting no differences in the relationship pattern between the two versions (see @fig-two).
+The linear regression predicting the BDI-II total score with the PHQ-4 depression score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.13,~95\%~CI~[-1.73, 1.47],~t(832) =  -0.16,~p = 0.871$; $\Delta \beta_{\text{refined}} = -0.05,~95\%~CI~[-0.70, 0.60],~t(832) = -0.15,~p = 0.883$), suggesting no differences in the relationship pattern between the two versions (see @fig-two).
 
-Moreover, Bayesian *t*-tests (using *BayesFactor*'s `ttestBF()` function with default priors; @moreyrouder2024) comparing the BDI-II scores between the refined and the original version at each integer score (0, 1, 2, 3) yielded no evidence in favour of a significant difference (BF > 3).
+Moreover, Bayesian *t*-tests [using *BayesFactor*'s `ttestBF()` function with default priors, @moreyrouder2024] comparing the BDI-II scores between the refined and the original version at each integer score (0, 1, 2, 3) yielded no evidence in favour of a significant difference (BF > 3).
 In other words, having the same score on the refined version as on the original version was related to the same outcome on the BDI-II.
@@
 
-The linear regression predicting the STAI-5 general score with the PHQ-4 anxiety score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.02, 95\% CI[-0.15, 0.11], t(832) =  -0.32, p = 0.750; \Delta \beta_{\text{refined}} = 0.01, 95\% CI[-0.03, 0.05], t(832) = 0.56 , p = 0.576$), suggesting no differences in the relationship pattern between the two versions.
+The linear regression predicting the STAI-5 general score with the PHQ-4 anxiety score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.02,~95\%~CI~[-0.15, 0.11],~t(832) =  -0.32,~p = 0.750$; $\Delta \beta_{\text{refined}} = 0.01,~95\%~CI~[-0.03, 0.05],~t(832) = 0.56,~p = 0.576$), suggesting no differences in the relationship pattern between the two versions.
 

### 2024-12-12 paper/4_WIP/manuscript.qmd
small changes plus caption docs

@@
 
-The short version of the State-Trait Anxiety Inventory [STAI-5, @zsido2020development] was used as a ground truth measure of anxiety.
-This abridged version of the STAI [@marteau1992development] includes 5 items rated on a 4-point Likert scale.
+The short version of the State-Trait Anxiety Inventory [STAI-T5, @zsido2020development] was used as a ground truth measure of anxiety.
+This abridged version of the STAI [@spielberger1970manual] includes 5 items rated on a 4-point Likert scale.
 Changes were made in the instructions from asking *"how participants feel right now"* to *"over the past 2 weeks"* to keep it consistent with the instructions of the PHQ-4 and BDI-II.

### 2024-12-13 paper/4_WIP/manuscript.qmd
minute changes to manuscript

--- 2024-12-12 paper/4_WIP/manuscript.qmd
@@
 The short version of the State-Trait Anxiety Inventory [STAI-5, @zsido2020development] was used as a ground truth measure of anxiety.
-This abridged version of the STAI [@marteau1992development] includes 5 items rated on a 4-point Likert scale.
+This abridged version of the STAI [@spielberger1970manual] includes 5 items rated on a 4-point Likert scale.
 Changes were made in the instructions from asking *"how participants feel right now"* to *"over the past 2 weeks"* to keep it consistent with the instructions of the PHQ-4 and BDI-II.
@@
 The original or refined version of the PHQ-4 was followed by the BDI-II, STAI-5, IAS, and MAIA-2, presented in random order.
-The IAS and the MAIA-2 were included as part of another study focused on interoception, and were only used as part in this study of participation quality control.
+The IAS and the MAIA-2 were included as part of another study focused on interoception, and ere only used in this stud as part of data quality control checks.
 

### 2024-12-13 paper/4_WIP/manuscript.qmd
updated manuscript

@@
 The original or refined version of the PHQ-4 was followed by the BDI-II, STAI-5, IAS, and MAIA-2, presented in random order.
-The IAS and the MAIA-2 were included as part of another study focused on interoception, and ere only used in this stud as part of data quality control checks.
-
+The IAS and the MAIA-2 were included as part of another study focused on interoception, and were only used in this stud as part of data quality control checks.
 ## Results

### 2024-12-14 paper/4_WIP/manuscript.qmd
fix numbers to match the figure results
@AnafNeves can you update the preprint

@@
 
-The results suggested that in the subclinical range of the BDI-II, the correlation between its score and the PHQ-4 Depression score was marginally higher (although not significantly, $p_{one-sided} = 0.161$) for the refined version compared to the original one.
+The results suggested that in the subclinical range of the BDI-II, the correlation between its score and the PHQ-4 Depression score was marginally higher (although not significantly, $p_{one-sided} = 0.164$) for the refined version compared to the original one.
 No correlation differences were observed in the moderate to severe range of the BDI-II.
@@
 For the STAI-5, there was no difference in the correlation between the refined and the original version in the subclinical range of the STAI-5.
-Surprisingly, we observed a stronger correlation between the refined PHQ-4 Anxiety score and the STAI-5 in the moderate to severe range compared to the original version ($p_{one-sided} = 0.027$).
+Surprisingly, we observed a stronger correlation between the refined PHQ-4 Anxiety score and the STAI-5 in the moderate to severe range compared to the original version ($p_{one-sided} = 0.017$).
 

### 2024-12-16 paper/4_PsychiatryResearch/manuscript.tex
add stuff

\documentclass[
  jou,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[nolongtablepatch]{lineno}
\linenumbers

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{\textbf{Measuring Depression and Anxiety with 4 items? Adaptation
of the PHQ-4 to increase its Sensitivity to Subclinical Variability}}

\shorttitle{Adaptation of the PHQ-4}

\usepackage{etoolbox}

\authorsnames[{1,2},{3},{1},{3,4,5}]{Dominique Makowski,An Shu Te,Ana
Neves,S.H. Annabel Chen}

\authorsaffiliations{
{School of Psychology, University of Sussex},{Sussex Centre for
Consciousness Science, University of Sussex},{School of Social
Sciences, Nanyang Technological University},{LKC Medicine, Nanyang
Technological University},{Centre for Research and Development in
Learning, Nanyang Technological University}}

\leftheader{Makowski, Te, Neves and Chen}

\abstract{The PHQ-4 is an ultra-brief (4 items) screening questionnaire
for depression and anxiety. In this brief report, we test the benefits
of adding one additional response option (``Once or twice'', in between
``Not at all'' and ``Several days'') to improve the scale's sensitivity
to milder alterations, and thus increase its usefulness in subclinical
populations. In study 1 (N=485), we provide evidence using Item Response
Theory (IRT) that the new response option does improve the scale's
psychometric quality and extends the sensitivity to the measured
constructs on the lower end of the spectrum. In study 2 (N=836), we show
that the refined version offers an improved sensitivity to subclinical
variability in depression (indexed by the BDI-II) as compared to the
original version. In conclusion, adding the ``once or twice'' response
option is a low-cost no-downsides way of increasing the PHQ-4's
sensitivity to subclinical variability, making it a tool of choice for
general population research.}

\keywords{PHQ-4, depression, anxiety, brief questionnaire validation,
ultra short scale}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{An Shu
Te}{0000-0002-9312-5552}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{S.H. Annabel
Chen}{0000-0002-1540-5516}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, opacityback=0, arc=.35mm, bottomrule=.15mm, toprule=.15mm, left=2mm, breakable, leftrule=.75mm]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}

\end{tcolorbox}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; An Shu Te:   Project
administration, Resources, Writing -- original draft; Ana Neves:   Data
curation, Formal Analysis, Writing -- original draft, Writing -- review
\& editing; S.H. Annabel Chen:   Project
administration, Supervision, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique Makowski, Email: D.Makowski@sussex.ac.uk}
}

\usepackage{pbalance} 
\usepackage{float}
\makeatletter
\let\oldtpt\ThreePartTable
\let\endoldtpt\endThreePartTable
\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldtpt[#1]
}
\def\ThreePartTable@ii{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldtpt
}
\def\endThreePartTable{
\endoldtpt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\makeatletter
\let\endoldlt\endlongtable		
\def\endlongtable{
\hline
\endoldlt}
\makeatother

\newenvironment{twocolumntable}% environment name
{% begin code
\begin{table*}[!htbp]%
\onecolumn%
}%
{%
\twocolumn%
\end{table*}%
}% end code

\urlstyle{same}

\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\resetlinenumber[1]

The Patient Health Questionnaire-4 (PHQ-4) is an ultra brief measurement
of core signs of depression and anxiety
(\citeproc{ref-kroenke2009ultra}{Kroenke et al., 2009}). It consists of
two items for depression (PHQ--2,
\citeproc{ref-kroenke2003patient}{Kroenke et al., 2003}) and anxiety
(GAD--2, \citeproc{ref-kroenke2007anxiety}{Kroenke et al., 2007}), each
corresponding to DSM-5 diagnostic symptoms for major depressive disorder
(MDD) and generalized anxiety disorder (GAD). It has been validated
across many languages and populations
(\citeproc{ref-christodoulaki2022validation}{Christodoulaki et al.,
2022}; \citeproc{ref-materu2020psychometric}{Materu et al., 2020};
\citeproc{ref-mendoza2022factor}{Mendoza et al., 2022}), becoming one of
the most popular screening instruments for depression and anxiety
(\citeproc{ref-maurer2018depression}{Maurer et al., 2018}).

While the scale has been validated and used in the general population
and non-clinical samples (\citeproc{ref-hajek2020prevalence}{Hajek \&
König, 2020}; \citeproc{ref-lowe20104}{Löwe et al., 2010}), its initial
purpose was to reliably discriminate and identify potential MDD/GAD
patients. This discriminative goal materializes in the scale's design
and the existence of categorical cut-offs, which does not necessary
entail a focus on the sensitivity to milder mood alterations. In
particular, the gap between the two lowest possible answers, ``Not at
all'' and ``Several days'', is quite large and possibly leaves out the
possibility of more subtle occurrences. While this is not necessarily an
issue in clinical and diagnostic contexts, it might lead to a
sub-optimal discrimination of affective levels on the lower end of the
spectrum, important for instance in the context of subclinical
variability quantification.

This brief report aims at testing the possibility of enhancing - with
minimal changes to the original scale - the PHQ-4 sensitivity to mild
mood level inflections. In the first study, we will evaluate whether the
new response option is prevalently used by participants, and whether it
does capture a specific part of the latent measure. In the second study,
we will compare the refined PHQ-4 version to the original one in terms
of sensitivity to subclinical variability in depression, using the Beck
Depression Inventory (BDI-II, \citeproc{ref-beck1996beck}{Beck et al.,
1996}) and the State Trait Anxiety Inventory (STAI-5,
\citeproc{ref-zsido2020development}{Zsido et al., 2020}) as our
ground-truth measures of depression and anxiety.

\section{Study 1}\label{study-1}

\subsection{Method}\label{method}

\subsubsection{Participants}\label{participants}

The sample consists of 485 English-speaking participants (Mean age =
30.1 \(\pm\) 10.1 {[}18, 73{]}; 50.3\% females) from the general
population recruited via \emph{Prolific}, a crowd-sourcing platform
recognized for providing high quality data
(\citeproc{ref-peer2022data}{Peer et al., 2022}). The only inclusion
criterion was a fluent proficiency in English to ensure that the task
instructions would be well-understood. This study was approved by the
NTU Institutional Review Board (NTU IRB-2022-187). All participants
provided their informed consent prior to participation and were
incentivized after completing the study.

\subsubsection{Measures}\label{measures}

In the original PHQ-4, the instructions \emph{``Over the last 2 weeks,
how often have you been bothered by the following problems?''} are
followed with 4 items (A1 - \emph{Feeling nervous, anxious or on edge};
A2 - \emph{Not being able to stop or control worrying}; D1 -
\emph{Little interest or pleasure in doing things}; D2 - \emph{Feeling
down, depressed, or hopeless}). The original answer options are ``Not at
all'' (0), ``Several days'' (1), ``More than half the days'' (2),
``Nearly every day'' (3). The total score is computed by summing the
responses of each facet resulting in a 0-6 score for depression and
anxiety.

For the refined version, we added a ``Once or twice'' option between
``Not at all'' and ``Several days''in order to better capture potential
mild mood inflections (see \citeproc{ref-dobson1979equidistant}{Dobson
\& Mothersill, 1979} for the choice of the label). This new option was
scored as 0.5 to preserve the same scoring as the original version.

\subsubsection{Procedure}\label{procedure}

Participants were administered the refined PHQ-4 online as part of
another study, which contained additional questionnaires and tasks not
relevant fort the current analysis. The PHQ-4 was presented in a
randomized order with other questionnaires. The data is available in
open-access at
https://github.com/RealityBending/IllusionGameReliability.

\subsection{Results}\label{results}

The analysis was carried out using \emph{R 4.4}
(\citeproc{ref-RCoreTeam}{R Core Team, 2023}), the \emph{tidyverse}
(\citeproc{ref-wickham2019welcome}{Wickham et al., 2019}), and the
\emph{easystats} collection of packages
(\citeproc{ref-ludecke2019insight}{Lüdecke et al., 2019},
\citeproc{ref-ludecke2020extracting}{2020},
\citeproc{ref-ludecke2021performance}{2021};
\citeproc{ref-patil2022datawizard}{Patil et al., 2022}). All
reproducible scripts and complimentary analyses are available
open-access at https://github.com/DominiqueMakowski/PHQ4R

\begin{figure*}[!htbp]

{\caption{{A) Proportion of answers of each type to the four items. B)
Prevalence of answer pairs. C) Item Information Curves from IRT showing
the coverage by each item and response of the latent dimension.
Typically, an optimally informative item would display a large coverage
over theta, with each response presenting a narrow coverage (high
discrimination between different levels).}{\label{fig-one}}}}

\includegraphics[width=1\textwidth,height=\textheight]{../../study1/figures/figure1.png}

\end{figure*}

\subsubsection{Descriptive Statistics}\label{descriptive-statistics}

The reliability of the anxiety (\(Cronbach's~\alpha = 0.903\); RMSEA =
0.031) and depression (\(Cronbach's~\alpha = 0.841\); RMSEA = 0.044)
subscales is excellent. The proportion of response types stratified by
item (see Figure~\ref{fig-one}) shows that the new ``Once or twice''
option was the most prevalent response for all items (on average
selected in 29.12\% of cases).

\subsubsection{Item Response Theory}\label{item-response-theory}

Item Response Theory (IRT) provides insights into how well items and
responses capture an underlying latent trait \(\theta\). For each of the
subscales, we fitted a unidimensional graded response model (GRM,
\citeproc{ref-samejima1997graded}{Samejima, 1997}). For anxiety, the two
items captured 89.2\% of the variance of the latent anxiety dimension
(\(\theta_{anxiety}\)). The discrimination parameters suggested that the
first item was less precise (\(\alpha = 3.42\)) than the second item
(\(\alpha = 12.55\)) in its ability to discriminate between various
levels of anxiety (i.e., each response on the second item covers a more
exclusive range of \(\theta_{anxiety}\), as can be seen in
Figure~\ref{fig-one}). The two depression items captured 82.8\% of the
variance of its latent trait (\(\theta_{depression}\)), and the opposite
pattern was found: the first item had a higher precision
(\(\alpha = 16.46\)) than the first (\(\alpha = 2.41\)). However, it is
important to note that the ``less precise'' items were also the ones
covering a larger portion of the latent space (being more sensitive
especially on the lower end of the spectrum), offering an interesting
trade-off between sensitivity and precision. Importantly for our
objective, the added ``Once or twice'' option did cover a selective and
unique portion of the latent space.

\subsection{Discussion}\label{discussion}

The fact that the new ``Once or twice'' response option was the most
prevalent response speaks to its usefulness in capturing more accurately
participants' expression. The IRT analysis further revealed that this
response tracks with precision a unique portion of the variability in
the latent factors measured by the instrument. Taken together, our
results suggest that adding this option response increases the scale's
potential to discriminate average mood levels (which are superior to
zero) from lower-end extremes (the true zero).

One natural methodological limitation pertains to the interpretation of
the latent dimensions in an IRT framework applied to pairs of items. In
this study's context, references to for instance ``the latent anxiety
dimension'' merely corresponds to the amalgamation of the two items of
the anxiety subscale, and not to a more general and valid true anxiety
factor.

\section{Study 2}\label{study-2}

\subsection{Method}\label{method-1}

\subsubsection{Participants}\label{participants-1}

The initial sample consisted of 1053 participants, recruited (181 were
recruited on \emph{Prolific}, 772 students from the University of Sussex
via \emph{SONA}, and the rest through convenience sampling as part of
dissertation students' data collection). We used attention checks as the
primary target for participant exclusion. We excluded 194 participants
(18.42\%) for failing at least one attention check, and 23 (2.18\%) that
were outliers (\(|z_{robust}| > 2.58\)) on measures significantly
related to the probability of failing attention checks (namely, the
standard deviation of all the items of the IAS, as well as the the
multivariate distance obtained with the OPTICS algorithm, see
\citeproc{ref-theriault2024check}{Thériault et al., 2024}). The
experiment duration was not related to the probability of failing
attention checks and was thus not used as an exclusion criterion.

The final sample included 836 participants (Mean age = 25.1 \(\pm\) 11.3
{[}18, 76{]}; 73.8\% women). This study was approved by the University
of Sussex' Ethics Committee (ER/ASF25/4).

In this sample, 51 participants (6.10\%) were labelled as having
Depression, as indexed by the self-reported presence of MDD together
with the use of a treatment (antidepressent, anxiolytic and/or therapy),
and 87 participants (10.41\%) were labelled as having Anxiety, as
indexed by the self-reported presence of GAD or Panic Disorder, also
together with the use of a treatment.

\subsubsection{Measures}\label{measures-1}

Participants were randomly assigned to complete either the original or
refined version of the PHQ-4, which included one additional response
option (``Once or twice'') scored 0.5 (creating more possible total
scores - 0.5, 1.5, 2.5, etc.).

Beck's Depression Inventory (BDI-II, \citeproc{ref-beck1996beck}{Beck et
al., 1996}) was used as a ground truth measure of depressive symptoms.
It includes 21 items, each addressing a specific depression symptom and
offering four response options scored from 0 to 3. Participants are
instructed to select the option that best describes how they have felt
over the past two weeks. The total score is calculated by summing the
scores for all 21 items, with higher scores indicating greater severity
of depressive symptoms.

The short version of the State-Trait Anxiety Inventory (STAI-5,
\citeproc{ref-zsido2020development}{Zsido et al., 2020}) was used as a
ground truth measure of anxiety. This abridged version of the STAI
(\citeproc{ref-spielberger1970manual}{Spielberger, 1970}) includes 5
items rated on a 4-point Likert scale. Changes were made in the
instructions from asking \emph{``how participants feel right now''} to
\emph{``over the past 2 weeks''} to keep it consistent with the
instructions of the PHQ-4 and BDI-II. A general score of anxiety was
computed by averaging all the items.

Participants were also asked to complete two questionnaires of
interoception, namely the Interoceptive Accuracy Scale (IAS - 21 items
rated on analog scales, \citeproc{ref-murphy2020testing}{Murphy et al.,
2020}) and the Multidimensional Assessment of Interoceptive Awareness
(MAIA-2 - 37 items, \citeproc{ref-mehling2018multidimensional}{Mehling
et al., 2018}).

After demographic questions, participants were asked to report the
current presence of psychiatric issues (from a list), as well as the
usage of treatment (antidepressants, mood stabilizers, anxiolytics,
therapy). We indexed the presence of a depression when participants
reported suffering from either Major Depressive Disorder (MDD) or
Dysthymia, as well as undergoing a medical treatment. Similarly, we
indexed the presence of an anxiety disorder when participants reported
suffering from either Generalized Anxiety Disorder (GAD) or Panic
Disorder, as well as undergoing a medical treatment.

\subsubsection{Procedure}\label{procedure-1}

The original or refined version of the PHQ-4 was followed by the BDI-II,
STAI-5, IAS, and MAIA-2, presented in random order. The IAS and the
MAIA-2 were included as part of another study focused on interoception,
and were only used in this stud as part of data quality control checks.
\#\# Results

As all the scripts, analysis details and results tables are available
open-access at https://github.com/DominiqueMakowski/PHQ4R, we will focus
on reporting the main results.

\begin{figure*}[!htbp]

{\caption{{A) PHQ-4 depression and anxiety scores against their
respective ground-truth measures, the BDI-22 and the STAI-5. Bayes
factors in grey tell if there is a difference, for the same PHQ-4 score,
between the original and the refined version (BFs \textless{} 1 suggest
no difference and thus evidence for a comparability of the refined
version with respect fo the original scale. Bayes factors in yellow
represent how new in-between scores (0.5, 1.5, 2.5, \ldots) available
with refined version differ from the adjacent scores (BFs \textgreater{}
3 suggest that half a point of difference on the refined PHQ-4 relates
to a significant difference on the ground truth measure). BF \textless{}
1/3°, BF \textgreater{} 3*, BF \textgreater{} 10**, BF \textgreater{}
30***. B) Bootstrapped distributions of the difference of correlation
between the revised PHQ-4 scores and the original one for sub-clinical
threshold scores of depression and anxiety. Positive differences suggest
that the correlation between the ground-truth measure and the refined
PHQ-4 score was stronger compared to the original version. C) Predictive
power of the PHQ-4 scores on the presence of a depression or anxiety
disorder. The upper plots show the relationship modelled by a logistic
regression, while the above plots represent the ROC curves (in which a
line further away from the diagonal represents a higher combination of
sensitivity and specificity).}{\label{fig-two}}}}

\includegraphics[width=1\textwidth,height=\textheight]{../../study2/analysis/2_analysis_files/figure-html/figure-1.png}

\end{figure*}

\subsubsection{\texorpdfstring{PHQ-4 Depression \emph{vs.}
BDI-II}{PHQ-4 Depression vs. BDI-II}}\label{phq-4-depression-vs.-bdi-ii}

The linear regression predicting the BDI-II total score with the PHQ-4
depression score showed no interaction related to the PHQ-4 version
(\(\Delta Intercept_{\text{refined}} =  -0.13,~95\%~CI~[-1.73, 1.47],~t(832) =  -0.16,~p = 0.871\);
\(\Delta \beta_{\text{refined}} = -0.05,~95\%~CI~[-0.70, 0.60],~t(832) = -0.15,~p = 0.883\)),
suggesting no differences in the relationship pattern between the two
versions (see Figure~\ref{fig-two}).

Moreover, Bayesian \emph{t}-tests (using \emph{BayesFactor}'s
\texttt{ttestBF()} function with default priors,
\citeproc{ref-moreyrouder2024}{Morey \& Rouder, 2024}) comparing the
BDI-II scores between the refined and the original version at each
integer score (0, 1, 2, 3) yielded no evidence in favour of a
significant difference (BF \textgreater{} 3). In other words, having the
same score on the refined version as on the original version was related
to the same outcome on the BDI-II.

However, the low in-between scores from the refined version are overall
capturing significantly different levels of depression compared to the
adjacent scores. Scoring 0.5 was associated with a higher BDI-II score
than scoring 0 (BF \textgreater{} 30), and lower scores than scoring 1
(BF \textgreater{} 30). Similarly, scoring 1.5 was associated with a
higher BDI-II score than scoring 1 (BF \textgreater{} 30), but not lower
scores than scoring 2 (BF = 0.234).

\subsubsection{\texorpdfstring{PHQ-4 Anxiety \emph{vs.}
STAI-5}{PHQ-4 Anxiety vs. STAI-5}}\label{phq-4-anxiety-vs.-stai-5}

The linear regression predicting the STAI-5 general score with the PHQ-4
anxiety score showed no interaction related to the PHQ-4 version
(\(\Delta Intercept_{\text{refined}} =  -0.02,~95\%~CI~[-0.15, 0.11],~t(832) =  -0.32,~p = 0.750\);
\(\Delta \beta_{\text{refined}} = 0.01,~95\%~CI~[-0.03, 0.05],~t(832) = 0.56,~p = 0.576\)),
suggesting no differences in the relationship pattern between the two
versions.

Moreover, Bayesian \emph{t}-tests comparing the STAI-5 scores between
the refined and the original version at each integer score yielded no
evidence in favour of a significant difference. In other words, having
the same score on the refined version as on the original version was
related to the same outcome on the STAI-5.

However, comparing in-between scores with adjacent scores yielded mixed
results. Scoring 0.5 on the PHQ-4 anxiety was not significantly
associated with a different level of STAI-5 compared to scoring 0 (BF =
1.83), but was with scores of 1 (BF \textgreater{} 30). Similarly, there
was no evidence that scoring 1.5 was different from scoring 1 (BF =
0.605), but strong evidence that it was different from scoring 2 (BF
\textgreater{} 30).

\subsubsection{Correlation Differences}\label{correlation-differences}

While the relationship pattern (i.e., the slope of the linear
relationship) was not affected by the PHQ-4 version, we focused next on
testing the difference in the strength (i.e., the precision) of the
relationship, in particular at the lower end of the spectrum (i.e., for
sub-clinical threshold scores of the BDI-II and STAI-5). We bootstrapped
(2000 iterations) the difference in correlation between the refined and
the original version for each of the two ground-truth measures,
separately for the BDI-II subsamples (minimal to mild \textless= 18;
moderate to severe \textgreater{} 18) and the STAI-5 subsamples (minimal
to mild \textless{} 2; moderate to severe \textgreater= 2).

The results suggested that in the subclinical range of the BDI-II, the
correlation between its score and the PHQ-4 Depression score was
marginally higher (although not significantly,
\(p_{one-sided} = 0.164\)) for the refined version compared to the
original one. No correlation differences were observed in the moderate
to severe range of the BDI-II.

For the STAI-5, there was no difference in the correlation between the
refined and the original version in the subclinical range of the STAI-5.
Surprisingly, we observed a stronger correlation between the refined
PHQ-4 Anxiety score and the STAI-5 in the moderate to severe range
compared to the original version (\(p_{one-sided} = 0.017\)).

\subsubsection{Predictive Power}\label{predictive-power}

Finally, we tested the predictive power of the PHQ-4 depression and
anxiety scores on the presence of a depression or anxiety disorder,
respectively. We modeled the relationship with a logistic regression.
While the PHQ-4 was overall a strong predictor of the outcome, there was
no significant difference between the two PHQ-4 versions.

However, the ROC curves for the refined and the original version of the
PHQ-4, suggested that the refined version had a better sensitivity /
specificity trade-off (AUC = 78.36\%) compared to the original version
(AUC=75\%), in particular on the lower end of the spectrum. The
difference was negligible for anxiety.

\subsection{Discussion}\label{discussion-1}

These results suggest that the new ``Once or twice'' response option to
the PHQ-4 does help capturing more fine-grained variations of depressive
symptoms, particularly in the subclinical range. Importantly, adding
this new response option with the scoring of 0.5 does not disrupt the
quality of the scale, which scores remain comparable to that of the
original version.

The results for the anxiety subscale appear more mixed, with less
evident benefits. However, this might have been partly caused by our
design decision regarding the questionnaire used for the ground-truth
measure of anxiety. Indeed, we used the abridged version of the STAI,
which only included 5 items, arguably limiting the sensitivity of the
anxiety measure in the first place.

Finally, although we used a stricter criterion for classifying
participants as having a depression or an anxiety disorder by
restricting it to participants also reporting undergoing a medical
treatment, it was still based on self-reported data. Studies in
controlled clinical settings are needed to confirm the potential
benefits of the refined PHQ-4 in mood disorders detection accuracy.

\section{General Discussion}\label{general-discussion}

The objective of this study was to test the introduction of a ``Once or
twice'' response option to the PHQ-4 to enhance its sensitivity to
milder mood fluctuations. In the first study, we showed that the new
response option was used prevalently by participants and did capture a
unique portion of the depression and anxiety underlying dimensions. In
the second study, we showed that the refined version of the PHQ-4 was
able to better differentiate lower levels of depression compared to the
original version, while remaining comparable. Although the benefits of
this refinement appear to be fairly minor, and particularly marked for
the depression score compared to anxiety, this cost-free improvement
appear useful to implement when measuring depression and anxiety using
the PHQ-4 ultra-short screening questionnaire.

\section{Acknowledgements}\label{acknowledgements}

We would like to thank the dissertation students from the University of
Sussex for their help in data collection.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-beck1996beck}
Beck, A. T., Steer, R. A., Brown, G. K., et al. (1996). \emph{Beck
depression inventory}.

\bibitem[\citeproctext]{ref-christodoulaki2022validation}
Christodoulaki, A., Baralou, V., Konstantakopoulos, G., \& Touloumi, G.
(2022). Validation of the patient health questionnaire-4 (PHQ-4) to
screen for depression and anxiety in the greek general population.
\emph{Journal of Psychosomatic Research}, \emph{160}, 110970.

\bibitem[\citeproctext]{ref-dobson1979equidistant}
Dobson, K. S., \& Mothersill, K. J. (1979). Equidistant categorical
labels for construction of likert-type scales. \emph{Perceptual and
Motor Skills}, \emph{49}(2), 575--580.

\bibitem[\citeproctext]{ref-hajek2020prevalence}
Hajek, A., \& König, H.-H. (2020). Prevalence and correlates of
individuals screening positive for depression and anxiety on the phq-4
in the german general population: Findings from the nationally
representative german socio-economic panel (GSOEP). \emph{International
Journal of Environmental Research and Public Health}, \emph{17}(21),
7865.

\bibitem[\citeproctext]{ref-kroenke2003patient}
Kroenke, K., Spitzer, R. L., \& Williams, J. B. (2003). The patient
health questionnaire-2: Validity of a two-item depression screener.
\emph{Medical Care}, 1284--1292.

\bibitem[\citeproctext]{ref-kroenke2009ultra}
Kroenke, K., Spitzer, R. L., Williams, J. B., \& Löwe, B. (2009). An
ultra-brief screening scale for anxiety and depression: The PHQ--4.
\emph{Psychosomatics}, \emph{50}(6), 613--621.

\bibitem[\citeproctext]{ref-kroenke2007anxiety}
Kroenke, K., Spitzer, R. L., Williams, J. B., Monahan, P. O., \& Löwe,
B. (2007). Anxiety disorders in primary care: Prevalence, impairment,
comorbidity, and detection. \emph{Annals of Internal Medicine},
\emph{146}(5), 317--325.

\bibitem[\citeproctext]{ref-lowe20104}
Löwe, B., Wahl, I., Rose, M., Spitzer, C., Glaesmer, H., Wingenfeld, K.,
Schneider, A., \& Brähler, E. (2010). A 4-item measure of depression and
anxiety: Validation and standardization of the patient health
questionnaire-4 (PHQ-4) in the general population. \emph{Journal of
Affective Disorders}, \emph{122}(1-2), 86--95.

\bibitem[\citeproctext]{ref-ludecke2020extracting}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., \& Makowski, D. (2020).
Extracting, computing and exploring the parameters of statistical models
using r. \emph{Journal of Open Source Software}, \emph{5}(53), 2445.

\bibitem[\citeproctext]{ref-ludecke2021performance}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., Waggoner, P., \& Makowski,
D. (2021). Performance: An r package for assessment, comparison and
testing of statistical models. \emph{Journal of Open Source Software},
\emph{6}(60).

\bibitem[\citeproctext]{ref-ludecke2019insight}
Lüdecke, D., Waggoner, P. D., \& Makowski, D. (2019). Insight: A unified
interface to access information from model objects in r. \emph{Journal
of Open Source Software}, \emph{4}(38), 1412.

\bibitem[\citeproctext]{ref-materu2020psychometric}
Materu, J., Kuringe, E., Nyato, D., Galishi, A., Mwanamsangu, A.,
Katebalila, M., Shao, A., Changalucha, J., Nnko, S., \& Wambura, M.
(2020). The psychometric properties of PHQ-4 anxiety and depression
screening scale among out of school adolescent girls and young women in
tanzania: A cross-sectional study. \emph{BMC Psychiatry}, \emph{20}(1),
1--8.

\bibitem[\citeproctext]{ref-maurer2018depression}
Maurer, D. M., Raymond, T. J., \& Davis, B. N. (2018). Depression:
Screening and diagnosis. \emph{American Family Physician}, \emph{98}(8),
508--515.

\bibitem[\citeproctext]{ref-mehling2018multidimensional}
Mehling, W. E., Acree, M., Stewart, A., Silas, J., \& Jones, A. (2018).
The multidimensional assessment of interoceptive awareness, version 2
(MAIA-2). \emph{PloS One}, \emph{13}(12), e0208034.

\bibitem[\citeproctext]{ref-mendoza2022factor}
Mendoza, N. B., Frondozo, C. E., Dizon, J. I. W. T., \& Buenconsejo, J.
U. (2022). The factor structure and measurement invariance of the PHQ-4
and the prevalence of depression and anxiety in a southeast asian
context amid the COVID-19 pandemic. \emph{Current Psychology}, 1--10.

\bibitem[\citeproctext]{ref-moreyrouder2024}
Morey, R. D., \& Rouder, J. N. (2024). \emph{BayesFactor: Computation of
bayes factors for common designs}.
\url{https://CRAN.R-project.org/package=BayesFactor}

\bibitem[\citeproctext]{ref-murphy2020testing}
Murphy, J., Brewer, R., Plans, D., Khalsa, S. S., Catmur, C., \& Bird,
G. (2020). Testing the independence of self-reported interoceptive
accuracy and attention. \emph{Quarterly Journal of Experimental
Psychology}, \emph{73}(1), 115--133.

\bibitem[\citeproctext]{ref-patil2022datawizard}
Patil, I., Makowski, D., Ben-Shachar, M. S., Wiernik, B. M., Bacher, E.,
\& Lüdecke, D. (2022). Datawizard: An r package for easy data
preparation and statistical transformations. \emph{Journal of Open
Source Software}, \emph{7}(78), 4684.

\bibitem[\citeproctext]{ref-peer2022data}
Peer, E., Rothschild, D., Gordon, A., Evernden, Z., \& Damer, E. (2022).
Data quality of platforms and panels for online behavioral research.
\emph{Behavior Research Methods}, \emph{54}(4), 1643--1662.

\bibitem[\citeproctext]{ref-RCoreTeam}
R Core Team. (2023). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\bibitem[\citeproctext]{ref-samejima1997graded}
Samejima, F. (1997). Graded response model. In \emph{Handbook of modern
item response theory} (pp. 85--100). Springer.

\bibitem[\citeproctext]{ref-spielberger1970manual}
Spielberger, C. D. (1970). Manual for the state-trait anxiety inventory
(self-evaluation questionnaire). \emph{(No Title)}.

\bibitem[\citeproctext]{ref-theriault2024check}
Thériault, R., Ben-Shachar, M. S., Patil, I., Lüdecke, D., Wiernik, B.
M., \& Makowski, D. (2024). Check your outliers﻿! An introduction to
identifying statistical outliers in r with easystats. \emph{Behavior
Research Methods}, \emph{56}(4), 4162--4172.

\bibitem[\citeproctext]{ref-wickham2019welcome}
Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., et al.
(2019). Welcome to the tidyverse. \emph{Journal of Open Source
Software}, \emph{4}(43), 1686.

\bibitem[\citeproctext]{ref-zsido2020development}
Zsido, A. N., Teleki, S. A., Csokasi, K., Rozsa, S., \& Bandi, S. A.
(2020). Development of the short version of the spielberger
state---trait anxiety inventory. \emph{Psychiatry Research}, \emph{291},
113223.

\end{CSLReferences}

\end{document}

### 2024-12-17 paper/4_PsychiatryResearch/manuscript.tex
bits and bobs

@@
 
-\keywords{PHQ-4, depression, anxiety, brief questionnaire validation,
-ultra short scale}
+\keywords{PHQ-4, depression, anxiety, brief questionnaire
+validation, ultra short scale}
 
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, colframe=quarto-callout-note-color-frame, opacityback=0, arc=.35mm, bottomrule=.15mm, toprule=.15mm, left=2mm, breakable, leftrule=.75mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, opacityback=0]
 

### 2025-05-07 paper/4_PsychiatryResearch/manuscript.qmd
add stuff
fixed typo

--- 2024-12-14 paper/4_WIP/manuscript.qmd
@@
 title: "**Measuring Depression and Anxiety with 4 items? Adaptation of the PHQ-4 to increase its Sensitivity to Subclinical Variability**"
-shorttitle: "Short Title in Running Header"
+shorttitle: "Adaptation of the PHQ-4"
 author:
@@
       - Formal Analysis
+      - Writing – original draft
       - Writing – review & editing
@@
     documentmode: jou
+    keep-tex: true
 ---
@@
 While the scale has been validated and used in the general population and non-clinical samples [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a focus on the sensitivity to milder mood alterations.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not **necessarily** entail a focus on the sensitivity to milder mood alterations.
 In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and possibly leaves out the possibility of more subtle occurrences.

### 2025-06-05 paper/5_Collabra/manuscript.qmd
update
update analysis

--- 2024-12-14 paper/4_WIP/manuscript.qmd
@@
 In study 2 (N=836), we show that the refined version offers an improved sensitivity to subclinical variability in depression (indexed by the BDI-II) as compared to the original version.
-In conclusion, adding the "once or twice" response option is a low-cost no-downsides way of increasing the PHQ-4's sensitivity to subclinical variability, making it a tool of choice for general population research.
+In conclusion, adding the "once or twice" response option is a low-cost way of increasing the PHQ-4's sensitivity to subclinical variability, making it a tool of choice for general population research.
 #   **Significance Statement**.
@@
 numbered-lines: true
-floatsintext: true
+floatsintext: false
+mask: true
 bibliography: bibliography.bib
@@
 While the scale has been validated and used in the general population and non-clinical samples [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients.
-This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a focus on the sensitivity to milder mood alterations.
+This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessarily entail a focus on the sensitivity to milder mood alterations.
 In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and possibly leaves out the possibility of more subtle occurrences.
@@
 
-In this sample, 51 participants (6.10%) were labelled as having Depression, as indexed by the self-reported presence of MDD together with the use of a treatment (antidepressent, anxiolytic and/or therapy), and 87 participants (10.41%) were labelled as having Anxiety, as indexed by the self-reported presence of GAD or Panic Disorder, also together with the use of a treatment.
+In this sample, 51 participants (6.10%) were coded as having Depression, as indexed by the self-reported presence of MDD together with the use of a treatment (antidepressent, anxiolytic and/or therapy), and 87 participants (10.41%) were coded as having Anxiety, as indexed by the self-reported presence of GAD or Panic Disorder, also together with the use of a treatment.
 
@@
 The IAS and the MAIA-2 were included as part of another study focused on interoception, and were only used in this stud as part of data quality control checks.
+
 ## Results
@@
 In the second study, we showed that the refined version of the PHQ-4 was able to better differentiate lower levels of depression compared to the original version, while remaining comparable.
-Although the benefits of this refinement appear to be fairly minor, and particularly marked for the depression score compared to anxiety, this cost-free improvement appear useful to implement when measuring depression and anxiety using the PHQ-4 ultra-short screening questionnaire.
+Although the benefits of this refinement appear to be fairly minor, and particularly marked for the depression score compared to anxiety, this low-cost improvement appear useful to implement when measuring depression and anxiety using the PHQ-4 ultra-short screening questionnaire.
 

### 2025-06-05 paper/4_PsychiatryResearch/manuscript.tex

@@
 \documentclass[
-  jou,
+  man,
   floatsintext,
@@
 
-\keywords{PHQ-4, depression, anxiety, brief questionnaire
-validation, ultra short scale}
+\keywords{PHQ-4, depression, anxiety, brief questionnaire validation,
+ultra short scale}
 
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, opacityback=0]
-
-This preprint is a non-peer-reviewed work from the
-\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
-\begin{center}
-\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
-\end{center}
-
-\end{tcolorbox}  Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Dominique
+\par{       Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows: Dominique
 Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
@@
 
-\usepackage{pbalance} 
-\usepackage{float}
-\makeatletter
-\let\oldtpt\ThreePartTable
-\let\endoldtpt\endThreePartTable
-\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
-\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
-\onecolumn
-\begin{minipage}{0.5\textwidth}
-\oldtpt[#1]
-}
-\def\ThreePartTable@ii{\begin{figure}[!htbp]
-\onecolumn
-\begin{minipage}{0.5\textwidth}
-\oldtpt
-}
-\def\endThreePartTable{
-\endoldtpt
-\end{minipage}
-\twocolumn
-\end{figure}}
-\makeatother
-
 \makeatletter
@@
 \hline
-\endoldlt}
+\endoldlt
+}
 \makeatother
-
-\newenvironment{twocolumntable}% environment name
-{% begin code
-\begin{table*}[!htbp]%
-\onecolumn%
-}%
-{%
-\twocolumn%
-\end{table*}%
-}% end code
 
@@
 
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
-\definecolor{quarto-callout-color}{HTML}{909090}
-\definecolor{quarto-callout-note-color}{HTML}{0758E5}
-\definecolor{quarto-callout-important-color}{HTML}{CC1914}
-\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
-\definecolor{quarto-callout-tip-color}{HTML}{00A047}
-\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
-\definecolor{quarto-callout-color-frame}{HTML}{acacac}
-\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
-\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
-\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
-\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
-\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
-\makeatother
 \makeatletter

### 2025-06-05 paper/5_Collabra/manuscript.tex

--- 2024-12-17 paper/4_PsychiatryResearch/manuscript.tex
@@
   jou,
-  floatsintext,
   longtable,
@@
   notimes,
+  mask,
   colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, arc=.35mm, rightrule=.15mm, leftrule=.75mm, left=2mm, breakable, toprule=.15mm, colframe=quarto-callout-note-color-frame, colback=white, opacityback=0]
+\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, rightrule=.15mm, leftrule=.75mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, opacityback=0, colback=white, arc=.35mm, bottomrule=.15mm]
 
@@
 and were only used in this stud as part of data quality control checks.
-\#\# Results
+
+\subsection{Results}\label{results-1}
 

### 2025-06-05 study1/analysis.qmd

@@
     self-contained: true
+    toc: true
     fig-dpi: 300
@@
 names(df2) <- str_remove_all(names(df2), "PHQ4_")
+
+df2ori <- read.csv("../study2/data/data_raw.csv") |>
+  filter(PHQ4_Condition == "PHQ4 - Original") |>
+  select(Sex = Gender, Age, starts_with("PHQ4_"), -PHQ4_Duration, -PHQ4_Condition, -PHQ4_Order)
+
+names(df2ori) <- str_remove_all(names(df2ori), "PHQ4_")
 ```
@@
 
+Same on original version:
+
+```{r}
+#| eval: false
+
+m_anxiety2ori <- mirt(select(df2ori, contains("Anxiety")), model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)
+
+plot(m_anxiety2ori, type = "trace", theta_lim = c(-3, 3))
+plot(m_anxiety2ori, type = "infotrace", theta_lim = c(-3, 3))
+```
+
 #### Full
@@
 
+```{r}
+#| eval: false
+
+plot(m_depression2, type = "trace", theta_lim = c(-3, 3))
+plot(m_depression2, type = "infotrace", theta_lim = c(-3, 3))
+```
+
+```{r}
+#| eval: false
+
+m_dep2ori <- mirt(select(df2ori, contains("Depression")), model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)
+
+plot(m_dep2ori, type = "trace", theta_lim = c(-3, 3))
+plot(m_dep2ori, type = "infotrace", theta_lim = c(-3, 3))
+```
+
 :::
@@
     ggplot(aes(x=value)) + 
-    geom_area(data=estimate_density(data, at = c("Dimension", "Scoring"), method = "KernSmooth") |> 
+    geom_area(data=estimate_density(data, by = c("Dimension", "Scoring"), method = "KernSmooth") |> 
                 group_by(Scoring) |> 

### 2025-06-05 study2/analysis/2_analysis.qmd

@@
     self-contained: false
+    toc: true
 ---
@@
 library(ggdist)
+library(mirt)
 ```
@@
 
+### IRT
+
+These completementary analyses were suggested during review to also apply IRT to the second sample. Additionally, we also computed "empirical coverage" curves.
+
+```{r}
+dat_irt <- df |> 
+  filter(PHQ4_Condition == "Revised") |>
+  select(starts_with("PHQ4"), BDI2_Total, STAI5_General) |> 
+  mutate(Depression_3 = as.numeric(as.factor(PHQ4_Depression_3)),
+         Depression_4 = as.numeric(as.factor(PHQ4_Depression_4)),
+         Anxiety_1 = as.numeric(as.factor(PHQ4_Anxiety_1)),
+         Anxiety_2 = as.numeric(as.factor(PHQ4_Anxiety_2)))
+
+irt_model1 <- mirt::mirt(data = dat_irt[c("Depression_3", "Depression_4")],
+                  itemtype = 'graded',
+                  verbose = FALSE) 
+irt_model2 <- mirt::mirt(data = dat_irt[c("Anxiety_1", "Anxiety_2")],
+                         optimizer = "nlminb",
+                  itemtype = 'graded',
+                  verbose = FALSE)
+
+# plot(irt_model, type = "trace", theta_lim = c(-3, 3))
+
+irt_data1  <- mirt::fscores(irt_model1, full.scores.SE = TRUE) |>
+  as.data.frame() |> 
+  mutate(Groundtruth = dat_irt$BDI2_Total,
+         PHQ4_Score = dat_irt$PHQ4_Depression,
+         Dimension = "Depression") 
+irt_data2  <- mirt::fscores(irt_model2, full.scores.SE = TRUE) |>
+  as.data.frame() |> 
+  mutate(Groundtruth = dat_irt$STAI5_General,
+         PHQ4_Score = dat_irt$PHQ4_Anxiety,
+         Dimension = "Anxiety")
+
+irt_data1$Label1 <- insight::format_value(cor(irt_data1$F1, irt_data1$Groundtruth))
+irt_data1$Label2 <- insight::format_value(cor(irt_data1$F1, irt_data1$PHQ4_Score))
+irt_data2$Label1 <- insight::format_value(cor(irt_data2$F1, irt_data2$Groundtruth))
+irt_data2$Label2 <- insight::format_value(cor(irt_data2$F1, irt_data2$PHQ4_Score))
+
+irt_data <- rbind(irt_data1, irt_data2) |> 
+  mutate(Label1 = paste("r =", Label1),
+         Label2 = paste("r =", Label2))
+
+irt_data |> 
+  ggplot() +
+  geom_jitter(aes(x = F1, y = Groundtruth), width = 0.05, height = 0.05) +
+  geom_smooth(aes(x = F1, y = Groundtruth), method = "lm", color = "red", formula = "y ~ x") +
+  geom_text(data = dplyr::distinct(irt_data[c("Dimension", "Label1")]), aes(label = Label1),
+    x = -1, y = c(55, 3), size = 5, color = "red"
+  ) +
+  facet_wrap(~ Dimension, scales = "free_y", ncol = 1) +
+  labs(x = "Theta latent factor extracted from IRT", y = "Ground truth score (BDI-II or STAI-5)") +
+  theme_minimal() |
+  irt_data |> 
+  ggplot() +
+  geom_jitter(aes(x = F1, y = PHQ4_Score), width = 0.05, height = 0.05) +
+  geom_smooth(aes(x = F1, y = PHQ4_Score), method = "lm", color = "blue", formula = "y ~ x", se = FALSE) +
+  geom_smooth(aes(x = F1, y = PHQ4_Score), method = "lm", color = "darkorange", formula = "y ~ poly(x, 2)", se = FALSE) +
+  geom_text(data = dplyr::distinct(irt_data[c("Dimension", "Label2")]), aes(label = Label2),
+    x = -1, y = c(6, 6), size = 5, color = "red"
+  ) +
+  facet_wrap(~ Dimension, scales = "free_y", ncol = 1) +
+  labs(x = "Theta latent factor extracted from IRT", y = "PHQ4 Score (Average)") +
+  theme_minimal()
+```
+
+```{r}
+add_labels <- function(x) {
+  x <- case_when(
+    x == 0 ~ "Not at all",
+    x == 0.5 ~ "Once or twice",
+    x == 1 ~ "Several days",
+    x == 2 ~ "More than half the days",
+    .default = "Nearly every day"
+  )
+  fct_relevel(x, c("Not at all", "Once or twice", "Several days", "More than half the days", "Nearly every day"))
+}
+
+rbind(
+  mutate(dat_irt, 
+         Item = "D1: Feeling down, depressed, or hopeless", 
+         Response = add_labels(PHQ4_Depression_3),
+         Outcome = BDI2_Total),
+  mutate(dat_irt, 
+         Item = "D2: Little interest or pleasure in doing things",
+         Response = add_labels(PHQ4_Depression_4),
+         Outcome = BDI2_Total),
+  mutate(dat_irt, 
+         Item = "A1: Feeling nervous, anxious, or on edge",
+         Response = add_labels(PHQ4_Anxiety_1),
+         Outcome = STAI5_General),
+  mutate(dat_irt, 
+         Item = "A2: Not being able to stop or control worrying",
+         Response = add_labels(PHQ4_Anxiety_2),
+         Outcome = STAI5_General)
+  ) |>
+  estimate_density(method = "KernSmooth", select = "Outcome", by = c("Item", "Response")) |> 
+  ggplot(aes(x = x, y = y, color = Response)) +
+  geom_line(linewidth = 1) +
+  labs(x = "Ground truth measure (BDI or STAI)", y = "Distribution") +
+  facet_wrap(~ Item, scales = "free") +
+  scale_y_continuous(expand = expansion(add = c(0, 0.001))) +
+  scale_x_continuous(expand = expansion(add = c(0, 0))) +
+  see::scale_color_flat_d("rainbow") +
+  see::theme_modern(axis.title.space = 5) +
+  theme(strip.background = element_rect(fill = "#EEEEEE", colour = "white"),
+        axis.text.y = element_blank())
+```
+
 ## Anxiety

### 2025-06-05 paper/5_Collabra/manuscript.qmd
Update manuscript.qmd

@@
 The original or refined version of the PHQ-4 was followed by the BDI-II, STAI-5, IAS, and MAIA-2, presented in random order.
-The IAS and the MAIA-2 were included as part of another study focused on interoception, and were only used in this stud as part of data quality control checks.
+The IAS **(ref?)** and the MAIA-2 **(ref?)** were included as part of another focused on interoception, and were only used in this study as part of data quality control checks.
 

### 2025-06-05 paper/5_Collabra/manuscript.qmd
add refs
@AnafNeves okay no need i added that

--- 2025-06-05 paper/5_Collabra/manuscript.qmd
@@
 The original or refined version of the PHQ-4 was followed by the BDI-II, STAI-5, IAS, and MAIA-2, presented in random order.
-The IAS and the MAIA-2 were included as part of another study focused on interoception, and were only used in this stud as part of data quality control checks.
+The IAS [@murphy2020testing] and the MAIA-2 [@mehling2018multidimensional] were included as part of another focused on interoception, and were only used in this study as part of data quality control checks.
 

### 2025-06-05 paper/5_Collabra/manuscript.qmd
edits

@@
 This brief report aims at testing the possibility of enhancing - with minimal changes to the original scale - the PHQ-4 sensitivity to mild mood level inflections.
-In the first study, we will evaluate whether the new response option is prevalently used by participants, and whether it does capture a specific part of the latent measure.
+In the first study, we will evaluate whether the new response option is prevalently used by participants, and whether it does capture a specific part of the construct.
 In the second study, we will compare the refined PHQ-4 version to the original one in terms of sensitivity to subclinical variability in depression, using the Beck Depression Inventory [BDI-II, @beck1996beck] and the State Trait Anxiety Inventory [STAI-5, @zsido2020development] as our ground-truth measures of depression and anxiety.
@@
 The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of $\theta_{anxiety}$, as can be seen in @fig-one).
-The two depression items captured 82.8\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the first ($\alpha = 2.41$).
+The two depression items captured 82.8\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the second ($\alpha = 2.41$).
 However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive especially on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
-Importantly for our objective, the added "Once or twice" option did cover a selective and unique portion of the latent space.
+Most importantly for our objective, the added "Once or twice" option did cover a selective and unique portion of the latent space.
 
@@
 
-One natural methodological limitation pertains to the interpretation of the latent dimensions in an IRT framework applied to pairs of items.
-In this study's context, references to for instance "the latent anxiety dimension" merely corresponds to the amalgamation of the two items of the anxiety subscale, and not to a more general and valid true anxiety factor.
+**One natural methodological limitation pertains to the application of the IRT framework to pairs of items.
+While this is statistically sound (the graded model utilizes the full response pattern information and does not solely rely on the item covariance matrix for parameter estimation), it is important to underline that in our study's context, "the latent anxiety/depression dimension" merely corresponds to the amalgamation of the two items of the anxiety or depression subscale, and not to a more general and independent latent anxiety or depression factor.**
+
+**In summary, the main take-away of this first study is that the "Once or twice" response option appears as a popular choice, which begs the question of its usefulness in capturing more fine-grained variations of the underlying dimensions as measured by independent tools.**
 

### 2025-06-06 paper/5_Collabra/manuscript.qmd
clarified the use of the trait subscale of the STAI-5 + IAS acronym

@@
 In the first study, we will evaluate whether the new response option is prevalently used by participants, and whether it does capture a specific part of the construct.
-In the second study, we will compare the refined PHQ-4 version to the original one in terms of sensitivity to subclinical variability in depression, using the Beck Depression Inventory [BDI-II, @beck1996beck] and the State Trait Anxiety Inventory [STAI-5, @zsido2020development] as our ground-truth measures of depression and anxiety.
+In the second study, we will compare the refined PHQ-4 version to the original one in terms of sensitivity to subclinical variability in depression, using the Beck Depression Inventory [BDI-II, @beck1996beck] and the Trait scale of the State Trait Anxiety Inventory [STAIT-5, @zsido2020development] as our ground-truth measures of depression and anxiety.
 
@@
 We used attention checks as the primary target for participant exclusion. 
-We excluded 194 participants (18.42\%) for failing at least one attention check, and 23 (2.18\%) that were outliers ($|z_{robust}| > 2.58$) on measures significantly related to the probability of failing attention checks [namely, the standard deviation of all the items of the IAS, as well as the the multivariate distance obtained with the OPTICS algorithm, see @theriault2024check].
+We excluded 194 participants (18.42\%) for failing at least one attention check, and 23 (2.18\%) that were outliers ($|z_{robust}| > 2.58$) on measures significantly related to the probability of failing attention checks [namely, the standard deviation of all the items of the Interoceptive Accuracy Scale, as well as the the multivariate distance obtained with the OPTICS algorithm, see @theriault2024check].
 The experiment duration was not related to the probability of failing attention checks and was thus not used as an exclusion criterion.
@@
 
-The short version of the State-Trait Anxiety Inventory [STAI-5, @zsido2020development] was used as a ground truth measure of anxiety.
+The short version of the trait scale of the State-Trait Anxiety Inventory [STAIT-5, @zsido2020development] was used as a ground truth measure of anxiety.
 This abridged version of the STAI [@spielberger1970manual] includes 5 items rated on a 4-point Likert scale.
@@
 
-Participants were also asked to complete two questionnaires of interoception, namely the Interoceptive Accuracy Scale [IAS - 21 items rated on analog scales, @murphy2020testing] and the Multidimensional Assessment of Interoceptive Awareness [MAIA-2 - 37 items, @mehling2018multidimensional].
+Participants were also asked to complete two questionnaires of interoception, namely the Interoceptive Accuracy Scale [IAS, 21 items rated on analog scales, @murphy2020testing] and the Multidimensional Assessment of Interoceptive Awareness [MAIA-2 - 37 items, @mehling2018multidimensional].
 
@@
 
-The original or refined version of the PHQ-4 was followed by the BDI-II, STAI-5, IAS, and MAIA-2, presented in random order.
+The original or refined version of the PHQ-4 was followed by the BDI-II, STAIT-5, IAS, and MAIA-2, presented in random order.
 The IAS [@murphy2020testing] and the MAIA-2 [@mehling2018multidimensional] were included as part of another focused on interoception, and were only used in this study as part of data quality control checks.
@@
 #| label: fig-two
-#| fig-cap: "A) PHQ-4 depression and anxiety scores against their respective ground-truth measures, the BDI-22 and the STAI-5.
+#| fig-cap: "A) PHQ-4 depression and anxiety scores against their respective ground-truth measures, the BDI-22 and the STAIT-5.
 Bayes factors in grey tell if there is a difference, for the same PHQ-4 score, between the original and the refined version (BFs < 1 suggest no difference and thus evidence for a comparability of the refined version with respect fo the original scale.
@@
 
-### PHQ-4 Anxiety *vs.* STAI-5
+### PHQ-4 Anxiety *vs.* STAIT-5
 
-The linear regression predicting the STAI-5 general score with the PHQ-4 anxiety score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.02,~95\%~CI~[-0.15, 0.11],~t(832) =  -0.32,~p = 0.750$; $\Delta \beta_{\text{refined}} = 0.01,~95\%~CI~[-0.03, 0.05],~t(832) = 0.56,~p = 0.576$), suggesting no differences in the relationship pattern between the two versions.
+The linear regression predicting the STAIT-5 general score with the PHQ-4 anxiety score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.02,~95\%~CI~[-0.15, 0.11],~t(832) =  -0.32,~p = 0.750$; $\Delta \beta_{\text{refined}} = 0.01,~95\%~CI~[-0.03, 0.05],~t(832) = 0.56,~p = 0.576$), suggesting no differences in the relationship pattern between the two versions.
 
-Moreover, Bayesian *t*-tests comparing the STAI-5 scores between the refined and the original version at each integer score yielded no evidence in favour of a significant difference.
-In other words, having the same score on the refined version as on the original version was related to the same outcome on the STAI-5.
+Moreover, Bayesian *t*-tests comparing the STAIT-5 scores between the refined and the original version at each integer score yielded no evidence in favour of a significant difference.
+In other words, having the same score on the refined version as on the original version was related to the same outcome on the STAIT-5.
 
 However, comparing in-between scores with adjacent scores yielded mixed results.
-Scoring 0.5 on the PHQ-4 anxiety was not significantly associated with a different level of STAI-5 compared to scoring 0 (BF = 1.83), but was with scores of 1 (BF > 30).
+Scoring 0.5 on the PHQ-4 anxiety was not significantly associated with a different level of STAIT-5 compared to scoring 0 (BF = 1.83), but was with scores of 1 (BF > 30).
 Similarly, there was no evidence that scoring 1.5 was different from scoring 1 (BF = 0.605), but strong evidence that it was different from scoring 2 (BF > 30).
@@
 
-While the relationship pattern (i.e., the slope of the linear relationship) was not affected by the PHQ-4 version, we focused next on testing the difference in the strength (i.e., the precision) of the relationship, in particular at the lower end of the spectrum (i.e., for sub-clinical threshold scores of the BDI-II and STAI-5).
-We bootstrapped (2000 iterations) the difference in correlation between the refined and the original version for each of the two ground-truth measures, separately for the BDI-II subsamples (minimal to mild <= 18; moderate to severe > 18) and the STAI-5 subsamples (minimal to mild < 2; moderate to severe >= 2).
+While the relationship pattern (i.e., the slope of the linear relationship) was not affected by the PHQ-4 version, we focused next on testing the difference in the strength (i.e., the precision) of the relationship, in particular at the lower end of the spectrum (i.e., for sub-clinical threshold scores of the BDI-II and STAIT-5).
+We bootstrapped (2000 iterations) the difference in correlation between the refined and the original version for each of the two ground-truth measures, separately for the BDI-II subsamples (minimal to mild <= 18; moderate to severe > 18) and the STAIT-5 subsamples (minimal to mild < 2; moderate to severe >= 2).
 
@@
 
-For the STAI-5, there was no difference in the correlation between the refined and the original version in the subclinical range of the STAI-5.
-Surprisingly, we observed a stronger correlation between the refined PHQ-4 Anxiety score and the STAI-5 in the moderate to severe range compared to the original version ($p_{one-sided} = 0.017$).
+For the STAIT-5, there was no difference in the correlation between the refined and the original version in the subclinical range of the STAIT-5.
+Surprisingly, we observed a stronger correlation between the refined PHQ-4 Anxiety score and the STAIT-5 in the moderate to severe range compared to the original version ($p_{one-sided} = 0.017$).
 

### 2025-06-06 paper/5_Collabra/manuscript.qmd
replace STAIT with STAI (couldn't be bothered to change the figures)

--- 2025-06-05 paper/5_Collabra/manuscript.qmd
@@
 In the first study, we will evaluate whether the new response option is prevalently used by participants, and whether it does capture a specific part of the construct.
-In the second study, we will compare the refined PHQ-4 version to the original one in terms of sensitivity to subclinical variability in depression, using the Beck Depression Inventory [BDI-II, @beck1996beck] and the State Trait Anxiety Inventory [STAI-5, @zsido2020development] as our ground-truth measures of depression and anxiety.
+In the second study, we will compare the refined PHQ-4 version to the original one in terms of sensitivity to subclinical variability in depression, using the Beck Depression Inventory [BDI-II, @beck1996beck] and the Trait scale of the State-Trait Anxiety Inventory [STAI-5, @zsido2020development] as our ground-truth measures of depression and anxiety.
 
@@
 For the refined version, we added a "Once or twice" option between "Not at all" and "Several days"in order to better capture potential mild mood inflections [see @dobson1979equidistant for the choice of the label].
-This new option was scored as 0.5 to preserve the same scoring as the original version.
 
@@
 
-The reliability of the anxiety ($Cronbach's~\alpha = 0.903$; RMSEA = 0.031) and depression ($Cronbach's~\alpha = 0.841$; RMSEA = 0.044) subscales is excellent.
+The consistency of the anxiety ($Cronbach's~\alpha = 0.903$) and depression ($Cronbach's~\alpha = 0.841$) subscales is excellent (responses were treated as ordinal).
 The proportion of response types stratified by item (see @fig-one) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in 29.12\% of cases).
@@
 For each of the subscales, we fitted a unidimensional graded response model [GRM, @samejima1997graded].
-For anxiety, the two items captured 89.2\% of the variance of the latent anxiety dimension ($\theta_{anxiety}$).
+For anxiety, the two items captured 89.2\% of the variance (RMSEA = 0.031) of the latent anxiety dimension ($\theta_{anxiety}$).
 The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of $\theta_{anxiety}$, as can be seen in @fig-one).
-The two depression items captured 82.8\% of the variance of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the second ($\alpha = 2.41$).
+The two depression items captured 82.8\% of the variance (RMSEA = 0.044) of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the second ($\alpha = 2.41$).
 However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive especially on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
@@
 
-Participants were randomly assigned to complete either the original or refined version of the PHQ-4, which included one additional response option ("Once or twice") scored 0.5 (creating more possible total scores - 0.5, 1.5, 2.5, etc.).
+Participants were randomly assigned to complete either the original or refined version of the PHQ-4, which included one additional response option ("Once or twice"), scored 0.5 to preserve the same scoring as the original version (creating more possible total scores - 0.5, 1.5, 2.5, etc.).
 
@@
 
-The short version of the State-Trait Anxiety Inventory [STAI-5, @zsido2020development] was used as a ground truth measure of anxiety.
+The short version of the trait subscale of the State-Trait Anxiety Inventory [STAI-5, @zsido2020development] was used as a ground truth measure of anxiety.
 This abridged version of the STAI [@spielberger1970manual] includes 5 items rated on a 4-point Likert scale.
@@
 
-Finally, although we used a stricter criterion for classifying participants as having a depression or an anxiety disorder by restricting it to participants also reporting undergoing a medical treatment, it was still based on self-reported data.
+**One of the potential limitation of our study includes the choice of the measures used as external "ground-truth" of depression and anxiety, namely the BDI-II and the STAI-5.
+For instance, there is ongoing debate about the STAI's  discriminant validity (which might extend to its short form used in the present study).
+A recent meta-analysis reported that trait anxiety scores were more strongly associated with depressive than anxiety disorders [@knowles2020specificity], suggesting that the scale may rather capture general negative affect than specifically anxiety.
+Regarding the BDI-II, existing evidence that suggests its relative lack of sensitivity for low-severity depression [@olino2012measuring] - comparable to similar instruments of its size (e.g., CES-D), and an improvement over the BDI-I [@wahl2014standardization] - put into question its choice for our goal of showing increased sub-clinical sensitivity.
+Thus, future studies should verify these findings with alternative measures of anxiety and depression and clinically assessed populations.**
+
+Indeed, although we used a stricter criterion for classifying participants as having a depression or an anxiety disorder by restricting it to participants also reporting undergoing a medical treatment, it was still based on self-reported data.
 Studies in controlled clinical settings are needed to confirm the potential benefits of the refined PHQ-4 in mood disorders detection accuracy.
@@
 Although the benefits of this refinement appear to be fairly minor, and particularly marked for the depression score compared to anxiety, this low-cost improvement appear useful to implement when measuring depression and anxiety using the PHQ-4 ultra-short screening questionnaire.
+
+**While adding granularity to the response format appears useful for the PHQ-4, it is possible that similar benefits could be found with other scales and measures.
+While the use of a limited, highly discriminating set of response options is understandable for specific applications (e.g., clinical diagnosis), we recommend future studies to investigate response format and its potential improvement for other scales used in online surveys and general population research.**
 

### 2025-06-06 paper/5_Collabra/manuscript.qmd
Update manuscript.qmd

@@
 For each of the subscales, we fitted a unidimensional graded response model [GRM, @samejima1997graded].
-For anxiety, the two items captured 89.2\% of the variance (RMSEA = 0.031) of the latent anxiety dimension ($\theta_{anxiety}$).
+For anxiety, the latent anxiety dimension ($\theta_{anxiety}$) captured 89.2% of the total variance across the two items (RMSEA = 0.031).
 The discrimination parameters suggested that the first item was less precise ($\alpha = 3.42$) than the second item ($\alpha = 12.55$) in its ability to discriminate between various levels of anxiety (i.e., each response on the second item covers a more exclusive range of $\theta_{anxiety}$, as can be seen in @fig-one).
-The two depression items captured 82.8\% of the variance (RMSEA = 0.044) of its latent trait ($\theta_{depression}$), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the second ($\alpha = 2.41$).
+The latent depression trait ($\theta_{depression}$) The two depression items captured 82.8\% of the total variance across the two items (RMSEA = 0.044), and the opposite pattern was found: the first item had a higher precision ($\alpha = 16.46$) than the second ($\alpha = 2.41$).
 However, it is important to note that the "less precise" items were also the ones covering a larger portion of the latent space (being more sensitive especially on the lower end of the spectrum), offering an interesting trade-off between sensitivity and precision.
@@
 
-The linear regression predicting the BDI-II total score with the PHQ-4 depression score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.13,~95\%~CI~[-1.73, 1.47],~t(832) =  -0.16,~p = 0.871$; $\Delta \beta_{\text{refined}} = -0.05,~95\%~CI~[-0.70, 0.60],~t(832) = -0.15,~p = 0.883$), suggesting no differences in the relationship pattern between the two versions (see @fig-two).
+**A linear model testing the interaction effect $\Delta$ of the refined condition on the intercept (representing by how much the value of the outcome when the PHQ score is 0 changes for the refined version compared the original) and slope (its increase or decrease by the refined version compared to the original) of the relationship between the PHQ-4 depression score and the BDI-II total score was fitted.** The model predicting the BDI-II total score with the PHQ-4 depression score showed no interaction related to the PHQ-4 version ($\Delta Intercept_{\text{refined}} =  -0.13,~95\%~CI~[-1.73, 1.47],~t(832) =  -0.16,~p = 0.871$; $\Delta \beta_{\text{refined}} = -0.05,~95\%~CI~[-0.70, 0.60],~t(832) = -0.15,~p = 0.883$), suggesting no differences in the relationship pattern between the two versions (see @fig-two).
 

### 2025-06-06 paper/5_Collabra/manuscript.tex
done

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 original version. In conclusion, adding the ``once or twice'' response
-option is a low-cost no-downsides way of increasing the PHQ-4's
-sensitivity to subclinical variability, making it a tool of choice for
-general population research.}
+option is a low-cost way of increasing the PHQ-4's sensitivity to
+subclinical variability, making it a tool of choice for general
+population research.}
 
@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, rightrule=.15mm, leftrule=.75mm, toprule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, opacityback=0, colback=white, arc=.35mm, bottomrule=.15mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, colback=white, toprule=.15mm, opacityback=0, arc=.35mm, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, left=2mm, leftrule=.75mm]
 
@@
 \begin{center}
-\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
+\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
 \end{center}
@@
 patients. This discriminative goal materializes in the scale's design
-and the existence of categorical cut-offs, which does not necessary
+and the existence of categorical cut-offs, which does not necessarily
 entail a focus on the sensitivity to milder mood alterations. In
@@
 
+\textbf{Enabling a more precise assessment of low-severity alterations
+is important, as milder symptoms are significant predictors of future
+clinical disorders and are associated with present functional
+impairments and reduced quality of life
+(\citeproc{ref-cuijpers2004subthreshold}{Cuijpers \& Smit, 2004};
+\citeproc{ref-judd1998major}{Judd et al., 1998}). In addition, the
+growing reliance on large-scale, often online, psychological and
+epidemiological surveys, to monitor population-level mental health,
+evaluate interventions, or track responses to global stressors such as
+pandemics or geopolitical crises, demands tools that are both brief and
+sensitive to small but meaningful fluctuations.}
+
 This brief report aims at testing the possibility of enhancing - with
@@
 new response option is prevalently used by participants, and whether it
-does capture a specific part of the latent measure. In the second study,
-we will compare the refined PHQ-4 version to the original one in terms
-of sensitivity to subclinical variability in depression, using the Beck
+does capture a specific part of the construct. In the second study, we
+will compare the refined PHQ-4 version to the original one in terms of
+sensitivity to subclinical variability in depression, using the Beck
 Depression Inventory (BDI-II, \citeproc{ref-beck1996beck}{Beck et al.,
-1996}) and the State Trait Anxiety Inventory (STAI-5,
+1996}) and the Trait scale of the State-Trait Anxiety Inventory (STAI-5,
 \citeproc{ref-zsido2020development}{Zsido et al., 2020}) as our
@@
 mild mood inflections (see \citeproc{ref-dobson1979equidistant}{Dobson
-\& Mothersill, 1979} for the choice of the label). This new option was
-scored as 0.5 to preserve the same scoring as the original version.
+\& Mothersill, 1979} for the choice of the label).
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{../../study1/figures/figure1.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../../study1/figures/figure1.png}
 
@@
 
-The reliability of the anxiety (\(Cronbach's~\alpha = 0.903\); RMSEA =
-0.031) and depression (\(Cronbach's~\alpha = 0.841\); RMSEA = 0.044)
-subscales is excellent. The proportion of response types stratified by
-item (see Figure~\ref{fig-one}) shows that the new ``Once or twice''
-option was the most prevalent response for all items (on average
-selected in 29.12\% of cases).
+The consistency of the anxiety (\(Cronbach's~\alpha = 0.903\)) and
+depression (\(Cronbach's~\alpha = 0.841\)) subscales is excellent
+(responses were treated as ordinal). The proportion of response types
+stratified by item (see Figure~\ref{fig-one}) shows that the new ``Once
+or twice'' option was the most prevalent response for all items (on
+average selected in 29.12\% of cases).
 
@@
 subscales, we fitted a unidimensional graded response model (GRM,
-\citeproc{ref-samejima1997graded}{Samejima, 1997}). For anxiety, the two
-items captured 89.2\% of the variance of the latent anxiety dimension
-(\(\theta_{anxiety}\)). The discrimination parameters suggested that the
-first item was less precise (\(\alpha = 3.42\)) than the second item
-(\(\alpha = 12.55\)) in its ability to discriminate between various
-levels of anxiety (i.e., each response on the second item covers a more
-exclusive range of \(\theta_{anxiety}\), as can be seen in
-Figure~\ref{fig-one}). The two depression items captured 82.8\% of the
-variance of its latent trait (\(\theta_{depression}\)), and the opposite
-pattern was found: the first item had a higher precision
-(\(\alpha = 16.46\)) than the first (\(\alpha = 2.41\)). However, it is
-important to note that the ``less precise'' items were also the ones
-covering a larger portion of the latent space (being more sensitive
-especially on the lower end of the spectrum), offering an interesting
-trade-off between sensitivity and precision. Importantly for our
-objective, the added ``Once or twice'' option did cover a selective and
-unique portion of the latent space.
+\citeproc{ref-samejima1997graded}{Samejima, 1997}). For anxiety, the
+latent anxiety dimension (\(\theta_{anxiety}\)) captured 89.2\% of the
+total variance across the two items (RMSEA = 0.031). The discrimination
+parameters suggested that the first item was less precise
+(\(\alpha = 3.42\)) than the second item (\(\alpha = 12.55\)) in its
+ability to discriminate between various levels of anxiety (i.e., each
+response on the second item covers a more exclusive range of
+\(\theta_{anxiety}\), as can be seen in Figure~\ref{fig-one}). The
+latent depression trait (\(\theta_{depression}\)) The two depression
+items captured 82.8\% of the total variance across the two items (RMSEA
+= 0.044), and the opposite pattern was found: the first item had a
+higher precision (\(\alpha = 16.46\)) than the second
+(\(\alpha = 2.41\)). However, it is important to note that the ``less
+precise'' items were also the ones covering a larger portion of the
+latent space (being more sensitive especially on the lower end of the
+spectrum), offering an interesting trade-off between sensitivity and
+precision. Most importantly for our objective, the added ``Once or
+twice'' option did cover a selective and unique portion of the latent
+space.
 
@@
 
-One natural methodological limitation pertains to the interpretation of
-the latent dimensions in an IRT framework applied to pairs of items. In
-this study's context, references to for instance ``the latent anxiety
-dimension'' merely corresponds to the amalgamation of the two items of
-the anxiety subscale, and not to a more general and valid true anxiety
-factor.
+\textbf{One natural methodological limitation pertains to the
+application of the IRT framework to pairs of items. While this is
+statistically sound (the graded model utilizes the full response pattern
+information and does not solely rely on the item covariance matrix for
+parameter estimation), it is important to underline that in our study's
+context, ``the latent anxiety/depression dimension'' merely corresponds
+to the amalgamation of the two items of the anxiety or depression
+subscale, and not to a more general and independent latent anxiety or
+depression factor.}
+
+\textbf{In summary, the main take-away of this first study is that the
+``Once or twice'' response option appears as a popular choice, which
+begs the question of its usefulness in capturing more fine-grained
+variations of the underlying dimensions as measured by independent
+tools.}
 
@@
 related to the probability of failing attention checks (namely, the
-standard deviation of all the items of the IAS, as well as the the
-multivariate distance obtained with the OPTICS algorithm, see
-\citeproc{ref-theriault2024check}{Thériault et al., 2024}). The
-experiment duration was not related to the probability of failing
-attention checks and was thus not used as an exclusion criterion.
+standard deviation of all the items of the Interoceptive Accuracy Scale,
+as well as the the multivariate distance obtained with the OPTICS
+algorithm, see \citeproc{ref-theriault2024check}{Thériault et al.,
+2024}). The experiment duration was not related to the probability of
+failing attention checks and was thus not used as an exclusion
+criterion.
 
@@
 
-In this sample, 51 participants (6.10\%) were labelled as having
+In this sample, 51 participants (6.10\%) were coded as having
 Depression, as indexed by the self-reported presence of MDD together
 with the use of a treatment (antidepressent, anxiolytic and/or therapy),
-and 87 participants (10.41\%) were labelled as having Anxiety, as
-indexed by the self-reported presence of GAD or Panic Disorder, also
-together with the use of a treatment.
+and 87 participants (10.41\%) were coded as having Anxiety, as indexed
+by the self-reported presence of GAD or Panic Disorder, also together
+with the use of a treatment.
 
@@
 refined version of the PHQ-4, which included one additional response
-option (``Once or twice'') scored 0.5 (creating more possible total
-scores - 0.5, 1.5, 2.5, etc.).
+option (``Once or twice''). \textbf{To preserve comparability with the
+original PHQ-4 scoring system and avoid altering the scale's total score
+range or established cut-off thresholds, we assigned the new response
+option a value of 0.5, placing it midway between ``Not at all'' (0) and
+``Several days'' (1). While this scoring assumes equal spacing between
+response options - a common but imperfect convention in ordinal scales -
+it offers a pragmatic compromise between conceptual fidelity and applied
+utility.}
 
@@
 
-The short version of the State-Trait Anxiety Inventory (STAI-5,
-\citeproc{ref-zsido2020development}{Zsido et al., 2020}) was used as a
-ground truth measure of anxiety. This abridged version of the STAI
-(\citeproc{ref-spielberger1970manual}{Spielberger, 1970}) includes 5
-items rated on a 4-point Likert scale. Changes were made in the
-instructions from asking \emph{``how participants feel right now''} to
-\emph{``over the past 2 weeks''} to keep it consistent with the
-instructions of the PHQ-4 and BDI-II. A general score of anxiety was
+The short version of the trait subscale of the State-Trait Anxiety
+Inventory (STAI-5, \citeproc{ref-zsido2020development}{Zsido et al.,
+2020}) was used as a ground truth measure of anxiety. This abridged
+version of the STAI (\citeproc{ref-spielberger1970manual}{Spielberger,
+1970}) includes 5 items rated on a 4-point Likert scale. Changes were
+made in the instructions from asking \emph{``how participants feel right
+now''} to \emph{``over the past 2 weeks''} to keep it consistent with
+the instructions of the PHQ-4 and BDI-II. A general score of anxiety was
 computed by averaging all the items.
@@
 Participants were also asked to complete two questionnaires of
-interoception, namely the Interoceptive Accuracy Scale (IAS - 21 items
+interoception, namely the Interoceptive Accuracy Scale (IAS, 21 items
 rated on analog scales, \citeproc{ref-murphy2020testing}{Murphy et al.,
@@
 The original or refined version of the PHQ-4 was followed by the BDI-II,
-STAI-5, IAS, and MAIA-2, presented in random order. The IAS and the
-MAIA-2 were included as part of another study focused on interoception,
-and were only used in this stud as part of data quality control checks.
+STAI-5, IAS, and MAIA-2, presented in random order. The IAS
+(\citeproc{ref-murphy2020testing}{Murphy et al., 2020}) and the MAIA-2
+(\citeproc{ref-mehling2018multidimensional}{Mehling et al., 2018}) were
+included as part of another focused on interoception, and were only used
+in this study as part of data quality control checks.
 
@@
 
-\includegraphics[width=1\textwidth,height=\textheight]{../../study2/analysis/2_analysis_files/figure-html/figure-1.png}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{../../study2/analysis/2_analysis_files/figure-html/figure-1.png}
 
@@
 
-The linear regression predicting the BDI-II total score with the PHQ-4
-depression score showed no interaction related to the PHQ-4 version
+\textbf{A linear model testing the interaction effect \(\Delta\) of the
+refined condition on the intercept (representing by how much the value
+of the outcome when the PHQ score is 0 changes for the refined version
+compared the original) and slope (its increase or decrease by the
+refined version compared to the original) of the relationship between
+the PHQ-4 depression score and the BDI-II total score was fitted.} The
+model predicting the BDI-II total score with the PHQ-4 depression score
+showed no interaction related to the PHQ-4 version
 (\(\Delta Intercept_{\text{refined}} =  -0.13,~95\%~CI~[-1.73, 1.47],~t(832) =  -0.16,~p = 0.871\);
@@
 
-Finally, although we used a stricter criterion for classifying
+\textbf{One of the potential limitation of our study includes the choice
+of the measures used as external ``ground-truth'' of depression and
+anxiety, namely the BDI-II and the STAI-5. For instance, there is
+ongoing debate about the STAI's discriminant validity (which might
+extend to its short form used in the present study). A recent
+meta-analysis reported that trait anxiety scores were more strongly
+associated with depressive than anxiety disorders
+(\citeproc{ref-knowles2020specificity}{Knowles \& Olatunji, 2020}),
+suggesting that the scale may rather capture general negative affect
+than specifically anxiety. Regarding the BDI-II, existing evidence that
+suggests its relative lack of sensitivity for low-severity depression
+(\citeproc{ref-olino2012measuring}{Olino et al., 2012}) - comparable to
+similar instruments of its size (e.g., CES-D), and an improvement over
+the BDI-I (\citeproc{ref-wahl2014standardization}{Wahl et al., 2014}) -
+put into question its choice for our goal of showing increased
+sub-clinical sensitivity. Thus, future studies should verify these
+findings with alternative measures of anxiety and depression and
+clinically assessed populations.}
+
+Indeed, although we used a stricter criterion for classifying
 participants as having a depression or an anxiety disorder by
@@
 this refinement appear to be fairly minor, and particularly marked for
-the depression score compared to anxiety, this cost-free improvement
+the depression score compared to anxiety, this low-cost improvement
 appear useful to implement when measuring depression and anxiety using
 the PHQ-4 ultra-short screening questionnaire.
+
+\textbf{While adding granularity to the response format appears useful
+for the PHQ-4, it is possible that similar benefits could be found with
+other scales and measures. While the use of a limited, highly
+discriminating set of response options is understandable for specific
+applications (e.g., clinical diagnosis), we recommend future studies to
+investigate response format and its potential improvement for other
+scales used in online surveys and general population research.}
 
@@
 
+\bibitem[\citeproctext]{ref-cuijpers2004subthreshold}
+Cuijpers, P., \& Smit, F. (2004). Subthreshold depression as a risk
+indicator for major depressive disorder: A systematic review of
+prospective studies. \emph{Acta Psychiatrica Scandinavica},
+\emph{109}(5), 325--331.
+
 \bibitem[\citeproctext]{ref-dobson1979equidistant}
@@
 7865.
+
+\bibitem[\citeproctext]{ref-judd1998major}
+Judd, L. L., Akiskal, H. S., Maser, J. D., Zeller, P. J., Endicott, J.,
+Coryell, W., Paulus, M. P., Kunovac, J. L., Leon, A. C., Mueller, T. I.,
+et al. (1998). Major depressive disorder: A prospective study of
+residual subthreshold depressive symptoms as predictor of rapid relapse.
+\emph{Journal of Affective Disorders}, \emph{50}(2-3), 97--108.
+
+\bibitem[\citeproctext]{ref-knowles2020specificity}
+Knowles, K. A., \& Olatunji, B. O. (2020). Specificity of trait anxiety
+in anxiety and depression: Meta-analysis of the state-trait anxiety
+inventory. \emph{Clinical Psychology Review}, \emph{82}, 101928.
 
@@
 
+\bibitem[\citeproctext]{ref-olino2012measuring}
+Olino, T. M., Yu, L., Klein, D. N., Rohde, P., Seeley, J. R., Pilkonis,
+P. A., \& Lewinsohn, P. M. (2012). Measuring depression using item
+response theory: An examination of three measures of depressive
+symptomatology. \emph{International Journal of Methods in Psychiatric
+Research}, \emph{21}(1), 76--85.
+
 \bibitem[\citeproctext]{ref-patil2022datawizard}
@@
 
+\bibitem[\citeproctext]{ref-wahl2014standardization}
+Wahl, I., Löwe, B., Bjorner, J. B., Fischer, F., Langs, G., Voderholzer,
+U., Aita, S. A., Bergemann, N., Brähler, E., \& Rose, M. (2014).
+Standardization of depression measurement: A common metric was developed
+for 11 self-report depression measures. \emph{Journal of Clinical
+Epidemiology}, \emph{67}(1), 73--86.
+
 \bibitem[\citeproctext]{ref-wickham2019welcome}

### 2025-06-09 paper/5_Collabra/manuscript.qmd
minor changes to manuscript

@@
 
+**Enabling a more precise assessment of low-severity alterations is important, as milder symptoms are significant predictors of future clinical disorders and are associated with present functional impairments and reduced quality of life [@judd1998major; @cuijpers2004subthreshold].
+In addition, the growing reliance on large-scale, often online, psychological and epidemiological surveys, to monitor population-level mental health, evaluate interventions, or track responses to global stressors such as pandemics or geopolitical crises, demands tools that are both brief and sensitive to small but meaningful fluctuations.**
+
 This brief report aims at testing the possibility of enhancing - with minimal changes to the original scale - the PHQ-4 sensitivity to mild mood level inflections.
@@
 
-The consistency of the anxiety ($Cronbach's~\alpha = 0.903$) and depression ($Cronbach's~\alpha = 0.841$) subscales is excellent (responses were treated as ordinal).
+The consistency of the anxiety ($Cronbach's~\alpha = 0.903$) and depression ($Cronbach's~\alpha = 0.841$) subscales is excellent (**responses were treated as ordinal**).
 The proportion of response types stratified by item (see @fig-one) shows that the new "Once or twice" option was the most prevalent response for all items (on average selected in 29.12\% of cases).
@@
 
-Participants were randomly assigned to complete either the original or refined version of the PHQ-4, which included one additional response option ("Once or twice"), scored 0.5 to preserve the same scoring as the original version (creating more possible total scores - 0.5, 1.5, 2.5, etc.).
+Participants were randomly assigned to complete either the original or refined version of the PHQ-4, which included one additional response option ("Once or twice").
+**To preserve comparability with the original PHQ-4 scoring system and avoid altering the scale’s total score range or established cut-off thresholds, we assigned the new response option a value of 0.5, placing it midway between "Not at all" (0) and "Several days" (1).
+While this scoring assumes equal spacing between response options - a common but imperfect convention in ordinal scales - it offers a pragmatic compromise between conceptual fidelity and applied utility.**
 
@@
 
-**One of the potential limitation of our study includes the choice of the measures used as external "ground-truth" of depression and anxiety, namely the BDI-II and the STAI-5.
+**One of the potential limitations of our study includes the choice of the measures used as external "ground-truth" of depression and anxiety, namely the BDI-II and the STAI-5.
 For instance, there is ongoing debate about the STAI's  discriminant validity (which might extend to its short form used in the present study).

### 2025-06-09 paper/5_Collabra/manuscript.tex
minor changes to manuscript

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, colback=white, toprule=.15mm, opacityback=0, arc=.35mm, breakable, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, rightrule=.15mm, left=2mm, leftrule=.75mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, left=2mm, opacityback=0, rightrule=.15mm, leftrule=.75mm, arc=.35mm, breakable, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, colback=white, toprule=.15mm]
 
@@
 depression (\(Cronbach's~\alpha = 0.841\)) subscales is excellent
-(responses were treated as ordinal). The proportion of response types
-stratified by item (see Figure~\ref{fig-one}) shows that the new ``Once
-or twice'' option was the most prevalent response for all items (on
-average selected in 29.12\% of cases).
+(\textbf{responses were treated as ordinal}). The proportion of response
+types stratified by item (see Figure~\ref{fig-one}) shows that the new
+``Once or twice'' option was the most prevalent response for all items
+(on average selected in 29.12\% of cases).
 
@@
 
-\textbf{One of the potential limitation of our study includes the choice
-of the measures used as external ``ground-truth'' of depression and
-anxiety, namely the BDI-II and the STAI-5. For instance, there is
+\textbf{One of the potential limitations of our study includes the
+choice of the measures used as external ``ground-truth'' of depression
+and anxiety, namely the BDI-II and the STAI-5. For instance, there is
 ongoing debate about the STAI's discriminant validity (which might

### 2025-06-09 paper/5_Collabra/manuscript.tex
add response file

@@
 \par{ }
-\par{     \begin{tcolorbox}[enhanced jigsaw, left=2mm, opacityback=0, rightrule=.15mm, leftrule=.75mm, arc=.35mm, breakable, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, colback=white, toprule=.15mm]
+\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, bottomrule=.15mm, colback=white, opacityback=0, arc=.35mm, leftrule=.75mm, rightrule=.15mm, left=2mm, toprule=.15mm, colframe=quarto-callout-note-color-frame]
 
